### 2021-07-16 analysis/preliminaryAnalysis.Rmd
initial commit

---
title: "An assessment of statistical guidance for authors at influential academic journals across scientific disciplines: Preliminary analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = T)
```

```{r load_packages}
library(knitr) # for literate programming
library(papaja) # for article template
library(tidyverse) # for data munging
library(tidylog) # inline analysis feedback
library(here) # for finding files
```

```{r perform_preprocessing}
# loads raw data, performs preprocessing, saves preprocessed files
source(here('analysis','preprocessing.R'))
```

```{r load_processed_data}
# loads the processed data files output by the preprocessing performed above
load(here('data','processed','d_coding.R'))
```

```{r load_functions}
# load custom functions
source(here('analysis','functions.R'))
```

### 2021-07-16 analysis/preprocessing.R

@@ -1,0 +1,4 @@
+# pre-processing code goes here...
+# load data from 'primary' folder
+# do any necessary munging
+# save processed data (in csv format and R format)

### 2021-08-03 analysis/preprocessing.R
Clean coding, waiting on manual edits

@@ -2,3 +2,159 @@
 # load data from 'primary' folder
 # do any necessary munging
 # save processed data (in csv format and R format)
+library(tidyverse)
+library(assertr)
+
+# Load data
+d_coding <- read_csv(here::here("data", "primary", "data_coding.csv"))
+d_journals <- read_csv(here::here("data", "primary", "data_journals.csv"))
+
+d_coding <-
+  d_coding %>% 
+  
+  # Rename variables
+  rename(
+    coder = "Choose coder initials",
+    journal = "Enter journal name:",
+    has_guidance = "Does the journal provide any guidance related to statistics or data analysis? (including referring to external guidelines)",
+    has_external_guidance = "Are any external statistical guidelines referred to? (this includes external reporting guidelines, such as CONSORT; publisher guidelines shared by many journals, such as Nature life sciences reporting checklist; and references to specific papers covering statistical issues)",
+    external_guidance = "If \"Yes\", which external statistical guidelines are referred to? (NOTE that these external guidelines should not be included in your coding of mentioned topics for this journal)",
+    has_internal_guidance = "Other than referring to external guidelines, does the journal have its own statistical guidelines?",
+    has_p_value = "Is any guidance provided about p-values?",
+    p_value = "If YES, please copy and paste all guidance related to p-values here:",
+    has_significance = "Is any guidance provided about statistical significance?",
+    significance = "If YES, please copy and paste all guidance related to statistical significance here:",
+    has_null_hypo = "Is any guidance provided about null hypotheses?",
+    null_hypo = "If YES, please copy and paste all guidance related to null hypotheses here:",
+    has_sample_size = "Is any guidance provided about sample size justification (including statistical power)?",
+    sample_size = "If YES, please copy and paste all guidance related to sample size justification (including statistical power) here:",
+    has_conf_int = "Is any guidance provided about confidence intervals?",
+    conf_int = "If YES, please copy and paste all guidance related to confidence intervals here:",
+    has_effect_size = "Is any guidance provided about effect sizes?",
+    effect_size = "If YES, please copy and paste all guidance related to effect sizes here:",
[…]

### 2021-08-03 analysis/preprocessing.R
Update cleaning after manual changes, save cleaned coding

@@ -69,7 +69,7 @@ d_coding <-
   assert(in_set("DS/TH", "MM/TH", "MSH/TH", "TB/TH"), coder) %>% 
   
   # Check that there is one row per journal
-  # verify(nrow(.) == nrow(d_journals)) %>% 
+  verify(nrow(.) == nrow(d_journals)) %>%
   
   # Check that journals are unique
   assert(is_uniq, journal) %>% 
@@ -94,15 +94,6 @@ d_coding <-
     has_internal_guidance = 
       if_else(str_detect(has_internal_guidance, "Yes"), "Yes", has_internal_guidance),
     .after = has_internal_guidance
-  ) %>% 
-    
-  # Manual entry errors in "Lancet Respiratory Medicine"
-  # For now, correct here but asking Tom to check in manual entry
-  mutate(
-    subgroup = if_else(journal == "Lancet Respiratory Medicine", has_subgroup, subgroup),
-    has_subgroup = if_else(journal == "Lancet Respiratory Medicine", "YES", has_subgroup),
-    prespecify = if_else(journal == "Lancet Respiratory Medicine", has_prespecify, prespecify),
-    has_prespecify = if_else(journal == "Lancet Respiratory Medicine", "YES", has_prespecify)
   ) %>% 
   
   # Check that all booleans "Yes"/"No"/NA and convert to TRUE/FALSE/NA
@@ -136,8 +127,7 @@ for (var_bool in vars_bool) {
 }
 
 # Text should NOT be provided if boolean FALSE
-# TODO: Manual data entry error for external_guidelines; asked Tom; currently skipping this check, so change once resolved
-for (var_bool in vars_bool[2:length(vars_bool)]) {
+for (var_bool in vars_bool) {
   
   var_text <- str_remove(var_bool, "has_")
   
@@ -146,15 +136,12 @@ for (var_bool in vars_bool[2:length(vars_bool)]) {
   )
 }
 
-# d_coding <-
-#   d_coding %>%
-# 
-#   # Check that all journals with internal guidance have TRUE/FALSE (not NA) for booleans
-#   # TODO: fix 2 manual entry errors in BRITISH JOURNAL OF PHARMACOLOGY; asked Tom; for now commented out
-#   # filter(journal != "BRITISH JOURNAL OF PHARMACOLOGY")
-#   pointblank::col_vals_not_null(
-#     d_coding, vars(!!!vars_bool), ~. %>% filter(has_internal_guidance)
-#   )
[…]

### 2021-08-04 analysis/preprocessing.R
Tidy external guidance

@@ -143,5 +143,184 @@ d_coding <-
     vars(!!!vars_bool), ~. %>% filter(has_internal_guidance)
   )
 
+
+# EXTERNAL GUIDANCE -------------------------------------------------------
+
+# 89; too many guidelines for boolean columns; how to organize? for now semi-colon separated column, but could nest
+# make a separate table of guidelines
+
+
+
+# Tidy external guidance
+
+# For manual tidying, get all unique external guidance lines for visual inspection
+unique_ext_guidance_groups <-
+  d_coding  %>% 
+  select(external_guidance) %>%
+  filter(!is.na(external_guidance)) %>% 
+  distinct()
+
+# Manually edit some external guidance via lookup table
+cleaned_external_guidances <- tribble(
+  ~external_guidance, ~external_guidance_clean,
+  "We encourage all authors to utilize the EQUATOR network to identify appropriate reporting guidelines based upon study type. Authors should refer to the BRISQ reporting guidelines for any study in which human biospecimens are used. For authors reporting animal research, we encourage use of the ARRIVE guidelines. For studies using cell lines, authors should report the source of the line, whether the line has been authenticated and how, and the mycoplasma contamination status.",
+  "EQUATOR; BRISQ; ARRIVE",
+  
+  "CONSORT; STROBE; ARRIVE; The statistical guidelines advocated by the International Committee of Medical Journal Editors (Ann Intern Med 1988; 108: 266-73) should be followed.",
+  "CONSORT; STROBE; ARRIVE; ICMJE",
+  
+  "\"Design and Analysis Transparency: Authors should follow field standards for disclosing key aspects of research design and data analysis, and should report the standards used in their study. See the Equator Network for information about standards across disciplines.\";  International Union of Crystallography (IUCr) Guidelines; Standards for Reporting Enzymology Data (STRENDA) Guidelines",
+  "EQUATOR; IUCr; STRENDA",
+  
[…]

### 2021-08-04 analysis/preliminaryAnalysis.Rmd

@@
 output: html_document
+editor_options: 
+  chunk_output_type: console
 ---
@@
 library(here) # for finding files
+library(gtsummary) # for summary tables
 ```
@@
 # loads the processed data files output by the preprocessing performed above
-load(here('data','processed','d_coding.R'))
+load(here('data','processed','d_coding.rds'))
+load(here('data','processed','lookup_external_guidance.rds'))
 ```
@@
 ```
+Look at counts and proportions mentioning each topic. Use all journals (regardless of whether internal guidance is available) for denominator.
+```{r tbl_counts_props}
+d_coding %>%
+
+  # Recode NAs to FALSE for internal guidance elements
+  mutate(across(starts_with("has_") & (!contains("guidance")), ~replace_na(., FALSE) )) %>%
+
+  select(starts_with("has_")) %>% 
+  gtsummary::tbl_summary() %>%
+  add_n() %>% 
+  modify_footnote(everything() ~ NA)
+```
+
+Try a mosaic plot for internal guidance...nope!
+```{r plot_internal_guidance_mosaic}
+my_palette <- c("lightcoral", "lightgrey", "lightblue")
+
+d_internal_guidance <-
+  d_coding %>% 
+  select(journal, starts_with("has_") & (!contains("guidance"))) %>% 
+  pivot_longer(
+    -journal,
+    names_to = "topic",
+    values_to = "outcome",
+    values_drop_na = FALSE
+  ) %>% 
+  mutate(outcome = case_when(
+    outcome ~ "Mentioned",
+    !outcome ~ "Not mentioned",
+    TRUE ~ "Not applicable (No internal guidance)"
+  ))
+
+# plt <- 
+  d_internal_guidance %>%
+  ggplot(aes(x = topic, y = factor(journal))) +
+  geom_tile(aes(fill = outcome), colour = 'black', size = .25) +
+  scale_fill_manual(
+    values = my_palette,
+    # labels = c('Not mentioned', 'Unclear', 'Mentioned')
+  ) +
+  scale_x_discrete(position = 'top') +
+  papaja::theme_apa() +
+  xlab("Topics mentioned") +
+  theme(plot.margin = margin(0, 50, 0, 0),
+        axis.ticks = element_blank(),
+        axis.title.y =element_blank(),
+        panel.grid.major=element_line(colour=NA), 
+        panel.grid.minor=element_line(colour=NA),
+        legend.position = 'bottom',
+        axis.text.x  = element_text(angle = 45, hjust = 0),
+        strip.background = element_rect(
+          color="black", fill="white", size=0, linetype="solid"
+        ),
+        strip.text.y = element_text(
+          size = 8, color = "black", face = 'bold'
+        ))
+
+# ggsave('internal-guidance.png', width = 11, height = 10)
+```
+
+Try bar chart
+```{r plot_internal_guidance_bar}
+d_internal_guidance %>% 
+  ggplot(aes(y = topic, fill = outcome)) +
+  geom_bar() +
+  scale_fill_manual(values = my_palette)
+```
+
+Look at external guidance
+```{r tbl_external_guidance}
+lookup_external_guidance %>% 
+  knitr::kable()
+```

### 2021-08-04 analysis/preprocessing.R
Update cleaning based on responses to issue #2

@@ -253,7 +253,7 @@ d_coding <-
   
   # Manually tidy guidance
   mutate(external_guidance = case_when(
-    str_detect(external_guidance, "some papers related to statistics are mentioned") ~ NA_character_, 
+    # str_detect(external_guidance, "some papers related to statistics are mentioned") ~ NA_character_, 
     
     # Guidelines
     str_detect(external_guidance, "star methods") ~ "cell star methods",
@@ -271,8 +271,8 @@ d_coding <-
     str_detect(external_guidance, "p?risma(?!-p)") ~ "prisma",
     str_detect(external_guidance, "st?rega") ~ "strega",
     str_detect(external_guidance, "strobe(?!-me)") ~ "strobe",
-    str_detect(external_guidance, "nih preclinical") ~ "nih principles and guidelines for reporting preclinical research",
     str_detect(external_guidance, "https://www.nlm.nih.gov/services/research_report_guide.html") ~ "nlm research reporting guidelines and initiatives (https://www.nlm.nih.gov/services/research_report_guide.html)",
+    str_detect(external_guidance, "nih preclinical|^nih$") ~ "nih principles and guidelines for reporting preclinical research",
     str_detect(external_guidance, "https://www.amstat.org/asa/your-career/ethical-guidelines-for-statistical-practice.aspx") ~ "asa (https://www.amstat.org/asa/your-career/ethical-guidelines-for-statistical-practice.aspx)",
     
     # Publications

### 2021-08-04 analysis/preprocessing.R
Fix external guidance counts, add type, clean journals

@@ -4,10 +4,17 @@
 # save processed data (in csv format and R format)
 library(tidyverse)
 library(assertr)
-
+library(janitor)
 # Load data
 d_coding <- read_csv(here::here("data", "primary", "data_coding.csv"))
 d_journals <- read_csv(here::here("data", "primary", "data_journals.csv"))
+
+d_journals <-
+  d_journals %>% 
+  
+  # Rename variables
+  janitor::clean_names() %>% 
+  rename(journal = "full_journal_title")
 
 d_coding <-
   d_coding %>% 
@@ -253,7 +260,6 @@ d_coding <-
   
   # Manually tidy guidance
   mutate(external_guidance = case_when(
-    # str_detect(external_guidance, "some papers related to statistics are mentioned") ~ NA_character_, 
     
     # Guidelines
     str_detect(external_guidance, "star methods") ~ "cell star methods",
@@ -271,25 +277,47 @@ d_coding <-
     str_detect(external_guidance, "p?risma(?!-p)") ~ "prisma",
     str_detect(external_guidance, "st?rega") ~ "strega",
     str_detect(external_guidance, "strobe(?!-me)") ~ "strobe",
+    str_detect(external_guidance, "nature life sciences") ~ "nature life sciences reporting",
     str_detect(external_guidance, "https://www.nlm.nih.gov/services/research_report_guide.html") ~ "nlm research reporting guidelines and initiatives (https://www.nlm.nih.gov/services/research_report_guide.html)",
     str_detect(external_guidance, "nih preclinical|^nih$") ~ "nih principles and guidelines for reporting preclinical research",
     str_detect(external_guidance, "https://www.amstat.org/asa/your-career/ethical-guidelines-for-statistical-practice.aspx") ~ "asa (https://www.amstat.org/asa/your-career/ethical-guidelines-for-statistical-practice.aspx)",
     
     # Publications
+    str_detect(external_guidance, "altman") ~ "altman et al. 1983 (10.1136/bmj.286.6376.1489)",
+    str_detect(external_guidance, "boushey harris bruemmer and archer 2008") ~ "boushey et al. 2008 (10.1016/j.jada.2008.01.002)",
[…]

### 2021-08-04 analysis/preliminaryAnalysis.Rmd
Add n and prob by field

@@
 load(here('data','processed','d_coding.rds'))
+load(here('data','processed','d_journals.rds'))
 load(here('data','processed','lookup_external_guidance.rds'))
@@
 ```
-Look at counts and proportions mentioning each topic. Use all journals (regardless of whether internal guidance is available) for denominator.
-```{r tbl_counts_props}
-d_coding %>%
 
-  # Recode NAs to FALSE for internal guidance elements
-  mutate(across(starts_with("has_") & (!contains("guidance")), ~replace_na(., FALSE) )) %>%
+```{r set_style_defaults}
+theme_gtsummary_compact()
+```
+Look at counts and proportions mentioning each topic. Use all journals for denominator. Overall and by field.
+``````{r tbl_counts_props}
+d_journals %>% 
+  select(journal, esi_field) %>% 
+  left_join(d_coding, ., by = "journal") %>%
 
-  select(starts_with("has_")) %>% 
-  gtsummary::tbl_summary() %>%
-  add_n() %>% 
-  modify_footnote(everything() ~ NA)
+  # Recode NAs to FALSE
+  mutate(across(
+    everything()
+    # starts_with("has_") & (!contains("guidance")) # for internal guidance elements
+    , ~replace_na(., FALSE) )) %>%
+
+  select(esi_field, starts_with("has_")) %>% 
+  gtsummary::tbl_summary(
+    by = esi_field,
+      # statistic = list(everything() ~ "{p}% ({n} / {N})")
+  ) %>%
+  add_overall %>% 
+  modify_footnote(everything() ~ NA) %>% 
+  modify_caption("**Number and proportion of journals overall and by field providing internal statistical guidance by topic.** Denominator for all proportions is the number of journals in the field, regardless of whether the journal has internal guidance.")
 ```
+Some initial observations
+- Math and computer science are indeed at 0% across the board.
+- Clinical medicine has 100% guidance. Also high in related field: biology & biochemistry, immunology, molecular bio & genetics. Also psych and neuro fields.
 
@@
 ```
+
+External guidance types
+```{r tbl_external_guidance_types}
+lookup_external_guidance %>% 
+  count(guidance_type) %>% 
+  kable()
+```

### 2021-08-04 analysis/preprocessing.R
Fix typo

@@ -363,8 +363,6 @@ lookup_external_guidance <-
     TRUE ~ NA_character_
   ))
 
-# by field
-
 readr::write_csv(d_coding, file =  here::here("data", "processed", "d_coding.csv"))
 save(d_coding, file =  here::here("data", "processed", "d_coding.rds"))
 

### 2021-08-10 analysis/rawToPrimary.R
fix for direct download of journals data

@@ -9,7 +9,7 @@ d_coding <- read_csv(here("data","raw","Statistical guidance for authors coding 
 d_coding <- d_coding %>% select(-Timestamp, -`If relevant, use this space to note anything unusual about this journal that the team needs to discuss.`) # remove timestamps, remove notes
 write_csv(d_coding,here("data","primary","data_coding.csv")) # save file to primary data folder
 
-d_journals <- read_csv(here("data","raw","data_journals.csv")) # load raw data
+d_journals <- read_csv(here("data","raw","statGuide_journalList - PPPR_journalList.csv")) # load raw data
 # no changes necessary
 write_csv(d_journals,here("data","primary","data_journals.csv")) # save file to primary data folder
 

### 2021-08-10 analysis/rawToPrimary.R
removing column during raw to primary conversion

@@ -6,11 +6,12 @@ library(here)
 # Load raw (Google drive) data
 
 d_coding <- read_csv(here("data","raw","Statistical guidance for authors coding form 2.0 (Responses) - Form Responses.csv")) # load raw data
-d_coding <- d_coding %>% select(-Timestamp, -`If relevant, use this space to note anything unusual about this journal that the team needs to discuss.`) # remove timestamps, remove notes
+d_coding <- d_coding %>% 
+  select(-Timestamp, -`If relevant, use this space to note anything unusual about this journal that the team needs to discuss.`) # remove timestamps, remove notes
 write_csv(d_coding,here("data","primary","data_coding.csv")) # save file to primary data folder
 
-d_journals <- read_csv(here("data","raw","statGuide_journalList - PPPR_journalList.csv")) # load raw data
-# no changes necessary
+d_journals <- read_csv(here("data","raw","statGuide_journalList - PPPR_journalList.csv")) %>% # load raw data
+  select(-`TH check`) # remove column
 write_csv(d_journals,here("data","primary","data_journals.csv")) # save file to primary data folder
 
 

### 2021-08-12 analysis/rawToPrimary.R
note added for manual fix for parsing issue

@@ -1,5 +1,8 @@
 # convert raw data to primary data
 
+###!!! NB - the data file created by the code below in the primary folder results in a parsing warning that we've been unable to resolve. 
+# The problem is fixed by opening the csv file in the primary folder in Excel and saving it. Hopefully we can find a programmatic solution, but currently this manual step needs to be performed whenever the raw data are converted to primary.
+
 library(tidyverse)
 library(here)
 

### 2021-08-12 analysis/preprocessing.R
preparing to integrate shared publisher guidance

@@ -5,6 +5,7 @@
 library(tidyverse)
 library(assertr)
 library(janitor)
+
 # Load data
 d_coding <- read_csv(here::here("data", "primary", "data_coding.csv"))
 d_journals <- read_csv(here::here("data", "primary", "data_journals.csv"))
@@ -277,7 +278,7 @@ d_coding <-
     str_detect(external_guidance, "p?risma(?!-p)") ~ "prisma",
     str_detect(external_guidance, "st?rega") ~ "strega",
     str_detect(external_guidance, "strobe(?!-me)") ~ "strobe",
-    str_detect(external_guidance, "nature life sciences") ~ "nature life sciences reporting",
+    #str_detect(external_guidance, "nature life sciences") ~ "nature life sciences reporting",
     str_detect(external_guidance, "https://www.nlm.nih.gov/services/research_report_guide.html") ~ "nlm research reporting guidelines and initiatives (https://www.nlm.nih.gov/services/research_report_guide.html)",
     str_detect(external_guidance, "nih preclinical|^nih$") ~ "nih principles and guidelines for reporting preclinical research",
     str_detect(external_guidance, "https://www.amstat.org/asa/your-career/ethical-guidelines-for-statistical-practice.aspx") ~ "asa (https://www.amstat.org/asa/your-career/ethical-guidelines-for-statistical-practice.aspx)",
@@ -362,6 +363,17 @@ lookup_external_guidance <-
     str_detect(external_guidance, "apa|asa|cell star|frontiersin") ~ "publisher",
     TRUE ~ NA_character_
   ))
+
+# Tom addition: identify journals that inherit publisher guidance and insert that publisher guidance into the journal row
+# NB - there is one exception - the journal Scientific Data inherits the publisher guidance, but also has its own journal specific guidance, so we need to handle this one separately
+natureJournals <- d_coding %>%
+  filter(str_detect(external_guidance_list, "nature"))
+
+cellJournals <- d_coding %>%
+  filter(str_detect(external_guidance_list, "cell"))
+
+frontiersJournals <- d_coding %>%
+  filter(str_detect(external_guidance_list, "frontiers"))
 
 readr::write_csv(d_coding, file =  here::here("data", "processed", "d_coding.csv"))
 save(d_coding, file =  here::here("data", "processed", "d_coding.rds"))

### 2021-08-13 analysis/preliminaryAnalysis.Rmd
changes to integrate publisher guidance

@@
 title: "An assessment of statistical guidance for authors at influential academic journals across scientific disciplines: Preliminary analysis"
+author: "Maia & Tom"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
 output: html_document
@@
 
+```{r load_functions}
+# load custom functions
+source(here('analysis','functions.R'))
+```
+
 ```{r perform_preprocessing}
@@
 
-```{r load_functions}
-# load custom functions
-source(here('analysis','functions.R'))
+```{r set_style_defaults}
+theme_gtsummary_compact() # set table theme
 ```
 
-```{r set_style_defaults}
-theme_gtsummary_compact()
-```
+<br>
+We have now completed data collection for the first phase of this project.
+The preregistered study protocol is available [here](https://osf.io/cz9g3/).
+Below we present a preliminary analysis of the phase one data.
+
+There are three types of guidance:
+* journal specific i.e., guidance written specifically for a particular journal and including in author-facing instructions/guidance documentation
+* publisher i.e., guidance written for a group of journals that share the same publisher, for example "Cell STAR Methods"
+* external i.e., academic papers or non-publisher reporting guidelines (usually present on EQUATOR)
+
 Look at counts and proportions mentioning each topic. Use all journals for denominator. Overall and by field.
+
 ``````{r tbl_counts_props}
@@
 ```
+
 Some initial observations
@@
 Try a mosaic plot for internal guidance...nope!
+
 ```{r plot_internal_guidance_mosaic}
@@
 Try bar chart
+
 ```{r plot_internal_guidance_bar}
@@
 Look at external guidance
+
 ```{r tbl_external_guidance}
@@
 External guidance types
+
 ```{r tbl_external_guidance_types}

### 2021-08-13 analysis/functions.R

@@ -1,1 +1,3 @@
 # functions
+
+`%notin%` <- function(x,y) !(x %in% y)

### 2021-08-13 analysis/preprocessing.R

@@ -10,6 +10,7 @@ library(janitor)
 d_coding <- read_csv(here::here("data", "primary", "data_coding.csv"))
 d_journals <- read_csv(here::here("data", "primary", "data_journals.csv"))
 
+# Clean up column names
 d_journals <-
   d_journals %>% 
   
@@ -75,9 +76,6 @@ d_coding <-
   
   # Check for expected coder combinations
   assert(in_set("DS/TH", "MM/TH", "MSH/TH", "TB/TH"), coder) %>% 
-  
-  # Check that there is one row per journal
-  verify(nrow(.) == nrow(d_journals)) %>%
   
   # Check that journals are unique
   assert(is_uniq, journal) %>% 
@@ -151,7 +149,69 @@ d_coding <-
     vars(!!!vars_bool), ~. %>% filter(has_internal_guidance)
   )
 
+# SHARED PUBLISHER-LEVEL GUIDANCE -------------------------------------------------------
 
+# remove shared guidance from d_coding and retain in separate dataframe
+d_coding_shared <- d_coding %>% 
+  filter(str_detect(journal, "SHARED")) %>%
+  filter(journal %in% c("SHARED: Nature consolidated", "SHARED: Frontiers", "SHARED: STAR Methods")) # retain only the shared guidance we need (i.e., drop individual Nature guidelines which are covered by Nature consolidated)
+
+d_coding <- d_coding %>% 
+  filter(!str_detect(journal, "SHARED"))
+
+# Identify journals that inherit publisher guidance 
+# NB - there is one exception - the journal Scientific Data inherits the publisher guidance, but also has its own journal specific guidance
+# to address this, we have created a special row for the journal Scientific Data 
+natureJournals <- d_coding %>%
+  filter(str_detect(external_guidance, regex("nature life", ignore_case = T))) %>%
+  filter(journal != "Scientific Data") %>%
+  pull(journal)
+
+cellJournals <- d_coding %>%
+  filter(str_detect(external_guidance, regex("\\bstar\\b", ignore_case = T))) %>%
+  pull(journal)
+
+frontiersJournals <- d_coding %>%
+  filter(str_detect(external_guidance, regex("frontiers", ignore_case = T))) %>%
+  pull(journal)
+
+# For each publisher (nature, cell, frontiers) create a new dataframe from d_coding that identifies coders and journals
+# create a new dataframe from d_coding shared that represents the publisher guidance
+# bind these two dataframes
+
[…]

### 2021-08-13 analysis/preprocessing.R
homogenize arrive and arrive guidelines

@@ -338,7 +338,7 @@ d_coding <-
     str_detect(external_guidance, "p?risma(?!-p)") ~ "prisma",
     str_detect(external_guidance, "st?rega") ~ "strega",
     str_detect(external_guidance, "strobe(?!-me)") ~ "strobe",
-    #str_detect(external_guidance, "nature life sciences") ~ "nature life sciences reporting",
+    str_detect(external_guidance, "arrive guidelines") ~ "arrive",
     str_detect(external_guidance, "https://www.nlm.nih.gov/services/research_report_guide.html") ~ "nlm research reporting guidelines and initiatives (https://www.nlm.nih.gov/services/research_report_guide.html)",
     str_detect(external_guidance, "nih preclinical|^nih$") ~ "nih principles and guidelines for reporting preclinical research",
     str_detect(external_guidance, "https://www.amstat.org/asa/your-career/ethical-guidelines-for-statistical-practice.aspx") ~ "asa (https://www.amstat.org/asa/your-career/ethical-guidelines-for-statistical-practice.aspx)",

### 2021-08-13 analysis/preprocessing.R
handling Scientific Data Consolidated

@@ -154,7 +154,7 @@ d_coding <-
 # remove shared guidance from d_coding and retain in separate dataframe
 d_coding_shared <- d_coding %>% 
   filter(str_detect(journal, "SHARED")) %>%
-  filter(journal %in% c("SHARED: Nature consolidated", "SHARED: Frontiers", "SHARED: STAR Methods")) # retain only the shared guidance we need (i.e., drop individual Nature guidelines which are covered by Nature consolidated)
+  filter(journal %in% c("SHARED: Nature consolidated", "SHARED: Frontiers", "SHARED: STAR Methods", "SHARED: Scientific Data Consolidated")) # retain only the shared guidance we need (i.e., drop individual Nature guidelines which are covered by Nature consolidated)
 
 d_coding <- d_coding %>% 
   filter(!str_detect(journal, "SHARED"))
@@ -203,7 +203,10 @@ frontiers_df2 <- d_coding_shared %>% filter(journal == "SHARED: Frontiers") %>%
 
 frontiers_shared <- bind_cols(frontiers_df1, frontiers_df2)
 
-allPublishers_shared <- bind_rows(nature_shared, cell_shared, frontiers_shared)
+scientific_data_consolidated <- d_coding_shared %>% filter(journal == "SHARED: Scientific Data Consolidated") %>% # extract the special row for Scientific Data Consolidated
+  mutate(journal = "Scientific Data")
+
+allPublishers_shared <- bind_rows(nature_shared, cell_shared, frontiers_shared, scientific_data_consolidated)
 
 # remove journals from d_coding and replace with the dataframes where they are attached to publisher guidance
 d_coding <- d_coding %>% filter(journal %notin% allPublishers_shared$journal)
@@ -418,8 +421,6 @@ lookup_external_guidance <-
     str_detect(external_guidance, "consort|apa jars") ~ "reporting guideline",
     str_detect(external_guidance, "^nih|^nlm|^fda") ~ "reporting guideline",
     str_detect(external_guidance, "biosharing|hugenet") ~ "reporting guideline", #unsure
-    
-    str_detect(external_guidance, "^nature") ~ "publisher",
     str_detect(external_guidance, "apa|asa|cell star|frontiersin") ~ "publisher",
     TRUE ~ NA_character_
   ))

### 2021-08-13 analysis/preliminaryAnalysis.Rmd
added some explanatory notes

@@
 <br>
-We have now completed data collection for the first phase of this project.
+We have now completed data collection for the first phase of this project in which we extracted author-facing statistical guidance provided by the top-15 journals (by Impact Factor) in each of 22 scientific disciplines (i.e., total N = 330 journals).
 The preregistered study protocol is available [here](https://osf.io/cz9g3/).
+
+> *Operational definition of statistical guidance*: guidance related to the appropriate selection, implementation, reporting, and interpretation of statistical analyses
+
+We recorded whether each journal:
+
+* Provided any statistical guidance at all
+* Had a dedicated statistical guidance section
+* Referred to statistical guidance in external sources (e.g., reporting guidelines or articles)
+
+*Note:* When journals referred to guidance in external sources, we have recorded the names of the sources, but *we have currently not extracted guidance from those sources*.
+We return to this point in the *issues to address* section at the end of this document.
+
+* If any statistical guidance was provided by the journal, we coded (YES/NO) whether it mentioned each of 20 statistical topics (e.g., "p-values"), and extracted the verbatim text.
+
+All data extraction was performed by a first coder (DS, MM, TB, MSH) and a second coder (TEH) with differences resolved through discussion.
+
+*Note* Some publishers provided statistical guidance that was shared across their portfolio journals (specifically, 32 Nature journals, 12 Cell journals, 2 Frontiers journals) .
+Publisher guidance was coded by two team members (as above), but is represented in our data and analysis for each individual journal the guidance applies to.
+For example, the Nature guidance is inherited by 32 journals so appears 32 times in the dataset.
+
 Below we present a preliminary analysis of the phase one data.
-
-There are three types of guidance:
-* journal specific i.e., guidance written specifically for a particular journal and including in author-facing instructions/guidance documentation
-* publisher i.e., guidance written for a group of journals that share the same publisher, for example "Cell STAR Methods"
-* external i.e., academic papers or non-publisher reporting guidelines (usually present on EQUATOR)
 

### 2021-08-13 analysis/preliminaryAnalysis.Rmd
d_coding and d_journals combined to d_all

--- 2021-08-13 analysis/preliminaryAnalysis.Rmd
@@
 title: "An assessment of statistical guidance for authors at influential academic journals across scientific disciplines: Preliminary analysis"
-author: "Maia & Tom"
-date: "`r format(Sys.time(), '%d %B, %Y')`"
+author: "Analysis performed by: Maia Salholz-Hillel & Tom Hardwicke"
+date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
 output: html_document
@@
 
+## Introduction
 <br>
 We have now completed data collection for the first phase of this project.
+We extracted author-facing statistical guidance provided by the top-15 journals (by Impact Factor) in each of 22 scientific disciplines (i.e., total N = 330 journals).
 The preregistered study protocol is available [here](https://osf.io/cz9g3/).
-Below we present a preliminary analysis of the phase one data.
+Below we present a preliminary analysis of the phase one data and highlight some issues to discuss as a team.
 
-There are three types of guidance:
-* journal specific i.e., guidance written specifically for a particular journal and including in author-facing instructions/guidance documentation
-* publisher i.e., guidance written for a group of journals that share the same publisher, for example "Cell STAR Methods"
-* external i.e., academic papers or non-publisher reporting guidelines (usually present on EQUATOR)
+## Method summary
+
+> **Operational definition of statistical guidance**: any advice or instruction related to the appropriate selection, implementation, reporting, or interpretation of statistical analyses.
+
+We recorded whether each journal:
+
+* Provided any statistical guidance at all
+* Had a dedicated statistical guidance section
+* Referred to statistical guidance in external sources (e.g., reporting guidelines or articles)
+* If any statistical guidance was provided by the journal, we coded (YES/NO) whether it mentioned each of 20 statistical topics (e.g., "p-values"), and extracted the verbatim text.
+
+**Note:** When journals referred to guidance in external sources, we have recorded the names of the sources, but *we have currently not extracted guidance from those sources*.
+We return to this point in the *issues to address* section at the end of this document.
+
+All data extraction was performed by a first coder (DS, MM, TB, MSH) and a second coder (TEH) with any coding differences resolved through discussion.
+
+**Note** Some publishers provided statistical guidance that was shared across their portfolio of journals (specifically, 32 Nature journals, 12 Cell journals, and 2 Frontiers journals) .
+Publisher guidance was coded by two team members (as above), but is represented in our data and analysis for each individual journal the guidance applies to.
+For example, the Nature guidance is inherited by 32 journals so appears 32 times in the dataset.
+
+## Results
 
@@
 ```
+
+## Issues to address
+
+* Pre-registration including general clinical trials advice
+* Coding of external guidance

### 2021-08-13 analysis/preprocessing.R

@@ -425,11 +425,11 @@ lookup_external_guidance <-
     TRUE ~ NA_character_
   ))
 
-readr::write_csv(d_coding, file =  here::here("data", "processed", "d_coding.csv"))
-save(d_coding, file =  here::here("data", "processed", "d_coding.rds"))
+# bind d_coding and d_journals
+d_all <- inner_join(d_coding,d_journals, by = 'journal')
 
-readr::write_csv(d_journals, file =  here::here("data", "processed", "d_journals.csv"))
-save(d_journals, file =  here::here("data", "processed", "d_journals.rds"))
+readr::write_csv(d_all, file =  here::here("data", "processed", "d_all.csv"))
+save(d_coding, file =  here::here("data", "processed", "d_all.rds"))
 
 readr::write_csv(lookup_external_guidance, file =  here::here("data", "processed", "lookup_external_guidance.csv"))
 save(lookup_external_guidance, file =  here::here("data", "processed", "lookup_external_guidance.rds"))

### 2021-08-13 analysis/rawToPrimary.R

@@ -14,7 +14,7 @@ d_coding <- d_coding %>%
 write_csv(d_coding,here("data","primary","data_coding.csv")) # save file to primary data folder
 
 d_journals <- read_csv(here("data","raw","statGuide_journalList - PPPR_journalList.csv")) %>% # load raw data
-  select(-`TH check`) # remove column
+  select(-`TH check`,-coder) # remove columns
 write_csv(d_journals,here("data","primary","data_journals.csv")) # save file to primary data folder
 
 

### 2021-08-16 analysis/preprocessing.R
identifying shared publisher guidance

@@ -185,7 +185,8 @@ nature_df1 <- d_coding %>% filter(journal %in% natureJournals) %>%
 nature_df2 <- d_coding_shared %>% filter(journal == "SHARED: Nature consolidated") %>%
   select(-coder, -journal)
 
-nature_shared <- bind_cols(nature_df1, nature_df2)
+nature_shared <- bind_cols(nature_df1, nature_df2) %>%
+  mutate(publisher_guidance = "Nature") # add a marker to say these journals share Nature publisher guidance
 
 cell_df1 <- d_coding %>% filter(journal %in% cellJournals) %>%
   select(coder, journal)
@@ -193,7 +194,8 @@ cell_df1 <- d_coding %>% filter(journal %in% cellJournals) %>%
 cell_df2 <- d_coding_shared %>% filter(journal == "SHARED: STAR Methods") %>%
   select(-coder, -journal)
 
-cell_shared <- bind_cols(cell_df1, cell_df2)
+cell_shared <- bind_cols(cell_df1, cell_df2) %>%
+  mutate(publisher_guidance = "Cell") # add a marker to say these journals share Cell publisher guidance
 
 frontiers_df1 <- d_coding %>% filter(journal %in% frontiersJournals) %>%
   select(coder, journal)
@@ -201,20 +203,25 @@ frontiers_df1 <- d_coding %>% filter(journal %in% frontiersJournals) %>%
 frontiers_df2 <- d_coding_shared %>% filter(journal == "SHARED: Frontiers") %>%
   select(-coder, -journal)
 
-frontiers_shared <- bind_cols(frontiers_df1, frontiers_df2)
+frontiers_shared <- bind_cols(frontiers_df1, frontiers_df2) %>%
+  mutate(publisher_guidance = "Frontiers") # add a marker to say these journals share Frontiers publisher guidance
 
 scientific_data_consolidated <- d_coding_shared %>% filter(journal == "SHARED: Scientific Data Consolidated") %>% # extract the special row for Scientific Data Consolidated
   mutate(journal = "Scientific Data")
 
-allPublishers_shared <- bind_rows(nature_shared, cell_shared, frontiers_shared, scientific_data_consolidated)
+allPublishers_shared <- bind_rows(nature_shared, cell_shared, frontiers_shared, scientific_data_consolidated) %>%
+  mutate(has_publisher_guidance = T) # add a marker to say these journals have shared publisher guidance
 
 # remove journals from d_coding and replace with the dataframes where they are attached to publisher guidance
[…]

### 2021-08-16 analysis/preliminaryAnalysis.Rmd
Update preliminaryAnalysis.Rmd
added new graphs and other assorted changes

@@
 date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
-output: html_document
+output:
+  bookdown::html_document2
 editor_options: 
@@
 library(gtsummary) # for summary tables
+library(lemon) # to have axes across all graph facets
+library(ggthemr) # for ggplot theme
+library(ggrepel) # for ggplot text labels
 ```
@@
 # loads the processed data files output by the preprocessing performed above
-load(here('data','processed','d_coding.rds'))
-load(here('data','processed','d_journals.rds'))
+load(here('data','processed','d_all.rds'))
 load(here('data','processed','lookup_external_guidance.rds'))
@@
 theme_gtsummary_compact() # set table theme
+ggthemr('fresh') # set ggplot theme
 ```
 
-## Introduction
+# Introduction
 <br>
@@
 
-## Method summary
+# Methods summary
 
@@
 
-**Note** Some publishers provided statistical guidance that was shared across their portfolio of journals (specifically, 32 Nature journals, 12 Cell journals, and 2 Frontiers journals) .
+```{r shared_publisher_journals}
+# count number of journals that have shared publisher guidance
+publisher_nature_n <- d_all %>% filter(publisher_guidance == "Nature") %>% nrow()
+publisher_cell_n <- d_all %>% filter(publisher_guidance == "Cell") %>% nrow()
+publisher_frontiers_n <- d_all %>% filter(publisher_guidance == "Frontiers") %>% nrow()
+```
+
+**Note:** Some publishers provided statistical guidance that was shared across their portfolio of journals (specifically, `r publisher_nature_n` Nature journals, `r publisher_cell_n` Cell journals, and `r publisher_frontiers_n` Frontiers journals) .
 Publisher guidance was coded by two team members (as above), but is represented in our data and analysis for each individual journal the guidance applies to.
@@
 
-## Results
+# Results
 
-Look at counts and proportions mentioning each topic. Use all journals for denominator. Overall and by field.
+```{r summarise_data}
 
-``````{r tbl_counts_props}
-d_journals %>% 
-  select(journal, esi_field) %>% 
-  left_join(d_coding, ., by = "journal") %>%
+# couldn't figure out to extract summary data from gtsummary::tblsummary so making my own summary dataframe
+sum_tab <- d_all %>%
 
@@
   select(esi_field, starts_with("has_")) %>% 
-  gtsummary::tbl_summary(
-    by = esi_field,
-      # statistic = list(everything() ~ "{p}% ({n} / {N})")
+  group_by(esi_field) %>%
+  summarise(across(starts_with("has_"), list(n = sum, N = ~n()), .names = "{col}__{fn}")) %>%
+  adorn_totals(name = 'OVERALL') %>%
+  pivot_longer(cols = !esi_field, names_to = c('variable', '.value'), names_sep = '__', names_prefix = 'has_') %>%
+  mutate(prop = n/N, percent = round(prop*100, 0)) %>%
+  mutate(esi_abbr = case_when( # creating abbreivated field labels
+    esi_field == "OVERALL" ~ "OVERALL",
+    esi_field == "AGRICULTURAL SCIENCES" ~ "AGRI",
+    esi_field == "BIOLOGY & BIOCHEMISTRY" ~ "BIO",       
+    esi_field == "CHEMISTRY" ~ "CHEM" ,                   
+    esi_field == "CLINICAL MEDICINE" ~ "MED",           
+    esi_field == "COMPUTER SCIENCE" ~ "COMSCI",             
+    esi_field == "ECONOMICS & BUSINESS" ~ "ECON",         
+    esi_field == "ENGINEERING" ~ "ENGIN",                  
+    esi_field == "ENVIRONMENT_ECOLOGY" ~ "ECO",         
+    esi_field == "GEOSCIENCES" ~ "GEO",                  
+    esi_field == "IMMUNOLOGY" ~ "IMMUN",                   
+    esi_field == "MATERIALS SCIENCE" ~ "MATSCI",            
+    esi_field == "MATHEMATICS" ~ "MATH",                
+    esi_field == "MICROBIOLOGY" ~ "MICBIO",                 
+    esi_field == "MOLECULAR BIOLOGY & GENETICS" ~ "MOLBIO", 
+    esi_field == "Multidisciplinary" ~ "MULTI",            
+    esi_field == "NEUROSCIENCE & BEHAVIOR" ~ "NEURO",     
+    esi_field == "PHARMACOLOGY & TOXICOLOGY" ~ "PHARM",    
+    esi_field == "PHYSICS" ~ "PHYS",                     
+    esi_field == "PLANT & ANIMAL SCIENCE" ~ "PLANT",       
+    esi_field == "PSYCHIATRY_PSYCHOLOGY" ~ "PSY",       
+    esi_field == "SOCIAL SCIENCES, GENERAL" ~ "SOCSCI",     
+    esi_field == "SPACE SCIENCE" ~ "SPACE"          
+  )) %>%
+  mutate(esi_field_with_abbr = paste0(esi_field,' (', esi_abbr,')'))
+
+# get fields with no statistical guidance at all
+null_fields <- sum_tab %>% 
+  filter(
+    variable == "guidance",
+    prop == 0
   ) %>%
-  add_overall %>% 
-  modify_footnote(everything() ~ NA) %>% 
-  modify_caption("**Number and proportion of journals overall and by field providing internal statistical guidance by topic.** Denominator for all proportions is the number of journals in the field, regardless of whether the journal has internal guidance.")
+  pull(esi_field)
+
+# select only data on general any/internal/external statistical guidance classifications 
+sum_tab_general <- sum_tab %>%
+  filter(variable %in% c('guidance', 'internal_guidance', 'internal_guidance_section', 'external_guidance')) %>% # select only data on general classifications
+  mutate(variable = factor(variable, levels = c('guidance', 'internal_guidance', 'internal_guidance_section', 'external_guidance'))) # set variable order
+
+# select only data on topics
+sum_tab_topics <- sum_tab %>%
+  filter(esi_field %notin% null_fields) %>% # remove fields that have no statistical guidance at all
+  filter(variable %notin% c('guidance', 'internal_guidance', 'internal_guidance_section', 'external_guidance')) %>% # select only data on topics
+  rename(topic = variable)
+
+# get list of statistical topics in order of overall proportions
+ordered_topics <- sum_tab_topics %>%
+  filter(esi_field == "OVERALL") %>%
+  arrange(desc(prop)) %>% # order from highest to lowest
+  pull(topic)
+
+# set order of topics
+sum_tab_topics <- sum_tab_topics %>%
+  mutate(topic = factor(topic, levels = ordered_topics))
 ```
+
+Figure 1 shows the percentage of journals in each scientific discipline (N = 15) and overall (N = 330) that offered (a) any statistical guidance (`r sum_tab %>% filter(esi_field == "OVERALL", variable == "guidance") %>% pull(percent)`% overall); (b) journal-specific statistical guidance (`r sum_tab %>% filter(esi_field == "OVERALL", variable == "internal_guidance") %>% pull(percent)`% overall); (c) a dedicated statistical guidance section (`r sum_tab %>% filter(esi_field == "OVERALL", variable == "internal_guidance_section") %>% pull(percent)`% overall); or referred to (d) external guidelines, such as reporting guidelines or academic articles (`r sum_tab %>% filter(esi_field == "OVERALL", variable == "external_guidance") %>% pull(percent)`% overall).
+
+How much guidance was provided? Some journals mention just one thing, some mention several.
 
@@
 
-Try a mosaic plot for internal guidance...nope!
+<br>
+
+```{r fig.cap = "Percentage of journals offering statistical guidance by scientific discipline.
+Overall proportion shown by black diamond.
+Graphs ordered from left to right and top to bottom in order of highest proportion overall across scientific disciplines.", fig.topcaption = TRUE, fig.width = 10, fig.height = 8}
+# extract overall data
+sum_tab_general_overall <- sum_tab_general %>%
+  filter(esi_field == "OVERALL")
+
+sum_tab_general %>%
+  filter(esi_field != "OVERALL") %>%
+  ggplot(aes(x=percent, y = 1, colour = str_wrap(esi_field_with_abbr,25))) +
+  facet_rep_wrap(facets = vars(variable), ncol = 2, nrow = 2) +
+  coord_capped_cart(bottom='both', left='both') + # to break axes in bottom left corner
+  geom_point(size = 5, position = position_stack(), alpha = .5) +
+  geom_point(size = 6, position = position_stack(), data = sum_tab_general_overall, colour = 'black', shape = 18) +
+  geom_text_repel(
+    aes(label = ifelse(percent > 0,esi_abbr,''),
+        segment.square = T,
+        segment.inflect = T), #
+    #point.padding = 0.2,
+    force_pull   = 0, # do not pull toward data points
+    nudge_y      = 15,
+    #nudge_x = 5,
+    direction    = "x",
+    angle        = 90,
+    hjust        = 0,
+    segment.size = 0.2,
+    segment.curvature = -1e-20,
+    max.iter = 1e4, max.time = 1,
+    max.overlaps = Inf, # ensure all labels are shown even if overlapping
+    show.legend = FALSE # do not add legend entry
+  ) +
+  scale_y_continuous(limits = c(0,15), name = 'number of scientific disciplines') +
+  xlim(0,100) +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid.major.y = element_blank(),
+        legend.position = 'bottom') +
+  scale_colour_discrete(name = 'Scientific disciplines') +
+  guides(colour = guide_legend(nrow = 6, title.position = 'top'))
+```
+
+```{r fig.cap = "Percentage of journals offering statistical guidance on twenty key topics by scientific discipline.
+Overall proportion shown by black diamond.
+For presentational purposes, disciplines offering no guidance on individual topics are shown, but not labelled.
+Two scientific disciplines in which no journals offered any statistical guidance at all are not shown.
+Graphs ordered from left to right and top to bottom in order of highest proportion overall across scientific disciplines.", fig.topcaption = TRUE, fig.width = 10, fig.height = 28}
+# extract overall data
+sum_tab_topics_overall <- sum_tab_topics %>%
+  filter(esi_field == "OVERALL")
+
+sum_tab_topics %>%
+  filter(esi_field != "OVERALL") %>%
+  ggplot(aes(x=percent, y = 1, colour = str_wrap(esi_field_with_abbr,25))) +
+  facet_rep_wrap(facets = vars(topic), ncol = 2) +
+  coord_capped_cart(bottom='both', left='both') + # to break axes in bottom left corner
+  geom_point(size = 5, position = position_stack(), alpha = .5) +
+  geom_point(size = 6, position = position_stack(), data = sum_tab_topics_overall, colour = 'black', shape = 18) +
+  geom_text_repel(
+    aes(label = ifelse(percent > 0,esi_abbr,''),
+        segment.square = T,
+        segment.inflect = T), #
+    #point.padding = 0.2,
+    force_pull   = 0, # do not pull toward data points
+    nudge_y      = 25,
+    nudge_x = 10,
+    direction    = "x",
+    angle        = 90,
+    hjust        = 0,
+    segment.size = 0.2,
+    segment.curvature = -1e-20,
+    max.iter = 1e4, max.time = 1,
+    max.overlaps = Inf, # ensure all labels are shown even if overlapping
+    show.legend = FALSE # do not add legend entry
+  ) +
+  scale_y_continuous(limits = c(0,25), name = 'number of scientific disciplines') +
+  xlim(0,100) +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid.major.y = element_blank(),
+        legend.position = 'bottom') +
+  scale_colour_discrete(name = 'Scientific disciplines') +
+  guides(colour = guide_legend(nrow = 5, title.position = 'top'))
+```
+
+```{r}
+# try binning method
+# forPlot %>%
+#   #mutate(bin = cut_width(prop, width = 0.2, center = 0.1)) %>%
+#   mutate(bin = cut_width(percent, width = 20, center = 10)) %>%
+#   group_by(esi_field,bin) %>%
+#   summarise(n = n()) %>%
+#   group_by(bin) %>%
+#   summarise(esi_field = esi_field, pos = cumsum(n)) %>%
+#   merge(forPlot, by = "esi_field") %>% # rejoin columns from original dataframe
+#   ggplot(aes(x=percent, y = pos, colour = esi_field)) +
+#   geom_vline(xintercept = sum_tab %>% filter(variable == "guidance", esi_field == "OVERALL") %>% pull(percent)) +
+#   geom_point(size = 5) +
+#   geom_label_repel(aes(label = esi_abbr)) +
+#   theme_minimal() +
+#   theme(axis.title.y = element_blank(),
+#         axis.text.y = element_blank(),
+#         axis.ticks.y = element_blank()) +
+#   scale_colour_discrete()
+```
+
+```{r}
+# # trying out a dot histogram
+# d <- sum_tab %>%
+#   #filter(variable %in% c('guidance', 'internal_guidance', 'internal_guidance_section', 'external_guidance')) %>%
+#   filter(variable %in% c('guidance'))
+# 
+# sum_tab %>%
+#   filter(variable %in% c('guidance', 'internal_guidance', 'internal_guidance_section', 'external_guidance')) %>%
+#   ggplot(aes(x = prop, fill=esi_field, label = esi_abbr, group = esi_field)) + 
+#     #geom_histogram(aes(fill=esi_field), binwidth=0.1, colour="grey20", lwd=0.2) +
+#     geom_dotplot(binwidth=0.1, colour="grey20", lwd=0.2, stackgroups = T, binpositions = 'all', method = 'histodot', dotsize = 0.25, stackdir = 'center') +
+#     facet_grid(rows = vars(variable)) +
+#     theme(axis.title.y = element_blank(), # y axis labels are meaningless on this plot, remove them
+#       axis.text.y = element_blank(),
+#       legend.position = 'bottom')
+# 
+#   # looks ok... but difficult to tell which field is which (too many similar colours, and have to keep looking at legend)
+#   # could we put the field abbreviation text inside the dots? (maybe shorten to two letters)
+#   # also for OVERALL we need something different, e.g., a black vertical line
+```
 
@@
 
-d_internal_guidance <-
-  d_coding %>% 
+d_internal_guidance_noPublisher <-
+  d_all %>%
+  filter(has_internal_guidance == T, # select journals with internal guidance
+         has_publisher_guidance == F | journal == "Scientific Data") # remove journals with shared publisher guidance except for Scientific Data (which shared guidance and its own guidance)
+
+# were going to represent journals with publisher guidance in a single row per publisher. 
+# select exemplar journals for each publisher
+publisher_nature <- d_all %>% filter(journal == "Nature") %>% mutate(journal = paste0("Nature journals"," (n=", publisher_nature_n,')'))
+publisher_cell <- d_all %>% filter(journal == "CELL") %>% mutate(journal = paste0("Cell journals"," (n=", publisher_cell_n,')'))
+publiher_frontiers <- d_all %>% filter(journal == "Frontiers in Microbiology") %>% mutate(journal = paste0("Frontiers journals"," (n=", publisher_frontiers_n,')'))
+
+# combine the journals with no publisher guidance with the publisher guidance rows
+d_internal_guidance <- bind_rows(d_internal_guidance_noPublisher, publisher_nature, publisher_cell, publiher_frontiers)
+
+# for journals with shared publisher guidance we only need only row
+
+plt <- d_internal_guidance %>%
   select(journal, starts_with("has_") & (!contains("guidance"))) %>%
@@
     TRUE ~ "Not applicable (No internal guidance)"
-  ))
-
-# plt <- 
-  d_internal_guidance %>%
-  ggplot(aes(x = topic, y = factor(journal))) +
+  )) %>%
+  ggplot(aes(x = topic, y = str_wrap(journal,25))) +
     geom_tile(aes(fill = outcome), colour = 'black', size = .25) +
@@
     scale_x_discrete(position = 'top') +
-  papaja::theme_apa() +
+    #papaja::theme_apa() +
     xlab("Topics mentioned") +
@@
           axis.title.y =element_blank(),
+          axis.line = element_blank(),
           panel.grid.major=element_line(colour=NA),
@@
 
-Try bar chart
-
-```{r plot_internal_guidance_bar}
-d_internal_guidance %>% 
-  ggplot(aes(y = topic, fill = outcome)) +
-  geom_bar() +
-  scale_fill_manual(values = my_palette)
+```{r fig.height = 40}
+plt
 ```
-
 Look at external guidance
@@
 
-## Issues to address
+# Issues to address
 
@@
 * Coding of external guidance
+
+# Appendix 1
+
+Tabular data (represented in Figures 1 and 2)
+Look at counts and proportions mentioning each topic. Use all journals for denominator. Overall and by field.
+
+```{r tbl_counts_props}
+d_all %>%
+
+  # Recode NAs to FALSE
+  mutate(across(
+    everything()
+    # starts_with("has_") & (!contains("guidance")) # for internal guidance elements
+    , ~replace_na(., FALSE) )) %>%
+
+  select(esi_field, starts_with("has_")) %>% 
+  gtsummary::tbl_summary(
+    by = esi_field,
+      # statistic = list(everything() ~ "{p}% ({n} / {N})")
+  ) %>%
+  add_overall %>% 
+  modify_footnote(everything() ~ NA) %>% 
+  modify_caption("**Number and percentage of journals overall and by scientific discipline providing internal statistical guidance by topic.** Denominator for all proportions is the number of journals in the field, regardless of whether the journal has internal guidance.")
+```

### 2021-08-16 analysis/preliminaryAnalysis.Rmd
major update to preliminary analysis

--- 2021-08-13 analysis/preliminaryAnalysis.Rmd
@@
 date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
-output: html_document
+output:
+  bookdown::html_document2:
+    fig_caption: true
+    number_sections: true
+    global_numbering: true
 editor_options: 
@@
 library(gtsummary) # for summary tables
+library(lemon) # to have axes across all graph facets
+library(ggthemr) # for ggplot theme
+library(ggrepel) # for ggplot text labels
+library(scales) # to wrap text on axis labels
 ```
@@
 * Referred to statistical guidance in external sources (e.g., reporting guidelines or articles)
-* If any statistical guidance was provided by the journal, we coded (YES/NO) whether it mentioned each of 20 statistical topics (e.g., "p-values"), and extracted the verbatim text.
 
-**Note:** When journals referred to guidance in external sources, we have recorded the names of the sources, but *we have currently not extracted guidance from those sources*.
-We return to this point in the *issues to address* section at the end of this document.
+If any statistical guidance was provided by a journal:
+
+* We coded (YES/NO) whether it mentioned each of 20 statistical topics (e.g., "p-values"), and extracted the verbatim text.
+
+**Note:** When journals referred to guidance in external sources, we have recorded the names of the sources, but *we have not currently extracted guidance from those sources*.
+We return to this point in the *issues to address* section (Section \@ref(issues)) at the end of this document.
 
@@
 
-**Note** Some publishers provided statistical guidance that was shared across their portfolio of journals (specifically, 32 Nature journals, 12 Cell journals, and 2 Frontiers journals) .
-Publisher guidance was coded by two team members (as above), but is represented in our data and analysis for each individual journal the guidance applies to.
-For example, the Nature guidance is inherited by 32 journals so appears 32 times in the dataset.
+```{r shared_publisher_journals}
+# count number of journals that have shared publisher guidance
+publisher_nature_n <- d_all %>% filter(publisher_guidance == "Nature") %>% nrow()
+publisher_cell_n <- d_all %>% filter(publisher_guidance == "Cell") %>% nrow()
+publisher_frontiers_n <- d_all %>% filter(publisher_guidance == "Frontiers") %>% nrow()
+```
 
-## Results
+**Note:** We found that some publishers provided statistical guidance that was shared across their portfolio of journals (specifically, `r publisher_nature_n` Nature journals, `r publisher_cell_n` Cell journals, and `r publisher_frontiers_n` Frontiers journals).
+Shared publisher guidance was coded by two team members (as above), but is represented in our data and analysis multiple times - once for each individual journal it applies to.
+For example, the Cell journals guidance is inherited by `r publisher_cell_n` journals so appears `r publisher_cell_n` times in the dataset.
+Unless stated otherwise, publisher guidance is treated as journal-specific guidance (i.e., not external guidance) in the analyses presented below.
 
-Look at counts and proportions mentioning each topic. Use all journals for denominator. Overall and by field.
+# Results
 
-``````{r tbl_counts_props}
-d_journals %>% 
-  select(journal, esi_field) %>% 
-  left_join(d_coding, ., by = "journal") %>%
+```{r summarise_data}
+
+# couldn't figure out to extract summary data from gtsummary::tblsummary so making my own summary dataframe
+sum_tab <- d_all %>%
 
@@
   select(esi_field, starts_with("has_")) %>% 
-  gtsummary::tbl_summary(
-    by = esi_field,
-      # statistic = list(everything() ~ "{p}% ({n} / {N})")
+  group_by(esi_field) %>%
+  summarise(across(starts_with("has_"), list(n = sum, N = ~n()), .names = "{col}__{fn}")) %>%
+  adorn_totals(name = 'OVERALL') %>%
+  pivot_longer(cols = !esi_field, names_to = c('variable', '.value'), names_sep = '__', names_prefix = 'has_') %>%
+  mutate(prop = n/N, percent = round(prop*100, 0)) %>%
+  mutate(esi_abbr = case_when( # create abbreviated field labels
+    esi_field == "OVERALL" ~ "OVERALL",
+    esi_field == "AGRICULTURAL SCIENCES" ~ "AGRI",
+    esi_field == "BIOLOGY & BIOCHEMISTRY" ~ "BIO",       
+    esi_field == "CHEMISTRY" ~ "CHEM" ,                   
+    esi_field == "CLINICAL MEDICINE" ~ "MED",           
+    esi_field == "COMPUTER SCIENCE" ~ "COMSCI",             
+    esi_field == "ECONOMICS & BUSINESS" ~ "ECON",         
+    esi_field == "ENGINEERING" ~ "ENGIN",                  
+    esi_field == "ENVIRONMENT_ECOLOGY" ~ "ECO",         
+    esi_field == "GEOSCIENCES" ~ "GEO",                  
+    esi_field == "IMMUNOLOGY" ~ "IMMUN",                   
+    esi_field == "MATERIALS SCIENCE" ~ "MATSCI",            
+    esi_field == "MATHEMATICS" ~ "MATH",                
+    esi_field == "MICROBIOLOGY" ~ "MICBIO",                 
+    esi_field == "MOLECULAR BIOLOGY & GENETICS" ~ "MOLBIO", 
+    esi_field == "Multidisciplinary" ~ "MULTI",            
+    esi_field == "NEUROSCIENCE & BEHAVIOR" ~ "NEURO",     
+    esi_field == "PHARMACOLOGY & TOXICOLOGY" ~ "PHARM",    
+    esi_field == "PHYSICS" ~ "PHYS",                     
+    esi_field == "PLANT & ANIMAL SCIENCE" ~ "PLANT",       
+    esi_field == "PSYCHIATRY_PSYCHOLOGY" ~ "PSY",       
+    esi_field == "SOCIAL SCIENCES, GENERAL" ~ "SOCSCI",     
+    esi_field == "SPACE SCIENCE" ~ "SPACE"          
+  )) %>%
+  mutate(esi_field_with_abbr = paste0(esi_field,' (', esi_abbr,')')) %>%
+  mutate(var_display = case_when( # create display versions of topic names
+    variable == "guidance" ~ "Provides any statistical guidance",
+    variable == "internal_guidance" ~ "Provides journal-specific statistical guidance",
+    variable == "internal_guidance_section" ~ "Has a dedicated statistical guidance section",
+    variable == "external_guidance" ~ "Refers to external statistical guidance",
+    variable == "p_value" ~ "p values",
+    variable == "significance" ~ "statistical significance",
+    variable == "null_hypo" ~ "null hypotheses",
+    variable == "sample_size" ~ "sample size justification",
+    variable == "conf_int" ~ "confidence intervals",
+    variable == "effect_size" ~ "effect sizes",
+    variable == "multi_compare" ~ "multiple comparisons",
+    variable == "subgroup" ~ "subgroup analyses",
+    variable == "baseline_covar" ~ "baseline covariates",
+    variable == "non_param" ~ "non-parametric tests",
+    variable == "sensitivity" ~ "sensitivity analyses",
+    variable == "model_assume" ~ "checking model assumptions",
+    variable == "exclusion" ~ "data exclusions",
+    variable == "outliers" ~ "handling outliers",
+    variable == "missing" ~ "handling missing data",
+    variable == "one_sided" ~ "one-sided tests",
+    variable == "bayes" ~ "Bayesian statistics",
+    variable == "secondary" ~ "secondary outcomes",
+    variable == "prespecify" ~ "prespecification of analyses",
+    variable == "cat_continuous" ~ "categorisation of continuous data",
+    variable == "publisher_guidance" ~ "Shares publisher guidance"
+  ))
+
+# get fields with no statistical guidance at all
+null_fields <- sum_tab %>% 
+  filter(
+    variable == "guidance",
+    prop == 0
   ) %>%
-  add_overall %>% 
-  modify_footnote(everything() ~ NA) %>% 
-  modify_caption("**Number and proportion of journals overall and by field providing internal statistical guidance by topic.** Denominator for all proportions is the number of journals in the field, regardless of whether the journal has internal guidance.")
+  pull(esi_field)
+
+# select only data on general any/internal/external statistical guidance classifications 
+sum_tab_general <- sum_tab %>%
+  filter(variable %in% c('guidance', 'internal_guidance', 'internal_guidance_section', 'external_guidance')) %>% # select only data on general classifications
+  mutate(variable = factor(variable, levels = c('guidance', 'internal_guidance', 'internal_guidance_section', 'external_guidance')),
+         var_display = factor(var_display, levels = c("Provides any statistical guidance", "Provides journal-specific statistical guidance", "Has a dedicated statistical guidance section", "Refers to external statistical guidance"))) # set variable order
+
+# select only data on topics
+sum_tab_topics <- sum_tab %>%
+  filter(esi_field %notin% null_fields) %>% # remove fields that have no statistical guidance at all
+  filter(variable %notin% c('guidance', 'internal_guidance', 'internal_guidance_section', 'external_guidance', 'publisher_guidance')) %>% # select only data on topics
+  rename(topic = variable, topic_display = var_display)
+
+# get list of statistical topics in order of overall proportions
+ordered_topics <- sum_tab_topics %>%
+  filter(esi_field == "OVERALL") %>%
+  arrange(desc(prop)) %>% # order from highest to lowest
+  pull(topic_display)
+
+# set order of topics
+sum_tab_topics <- sum_tab_topics %>%
+  mutate(topic_display = factor(topic_display, levels = ordered_topics))
 ```
 
-Some initial observations
-- Math and computer science are indeed at 0% across the board.
-- Clinical medicine has 100% guidance. Also high in related field: biology & biochemistry, immunology, molecular bio & genetics. Also psych and neuro fields.
+Figure \@ref(fig:statguidefig) shows the percentage of journals in each scientific discipline (denominator is N = 15) and overall (denominator is N = 330) that (a) offered any statistical guidance; (b) had a dedicated statistical guidance section.
+For tabular data, see Appendix \@ref(app).
 
-Try a mosaic plot for internal guidance...nope!
+```{r crosstab_int_ext}
+crosstab_int_ext <- d_all %>% count(has_internal_guidance, has_external_guidance)
+only_ext <- crosstab_int_ext %>% filter(has_internal_guidance == F, has_external_guidance == T) %>% pull(n)
+```
 
-```{r plot_internal_guidance_mosaic}
-my_palette <- c("lightcoral", "lightgrey", "lightblue")
+You can see that around half (`r sum_tab %>% filter(esi_field == "OVERALL", variable == "guidance") %>% pull(percent)`%) of journals offered any statistical guidance.
+Note that this includes `r only_ext` journals which only referred to statistical guidance in external sources (reporting guidelines or academic papers).
+Just over quarter of journals had a dedicated statistical guidance section of their author instructions (`r sum_tab %>% filter(esi_field == "OVERALL", variable == "internal_guidance_section") %>% pull(percent)`%).
+In two fields (Computer Science and Maths), no journals offered any statistical guidance.
+Journals in health-related fields were more likely to offer statistical guidance and have dedicated statistical guidance sections.
+Notably, 100% of the journals in clinical medicine offered some statistical guidance.
 
-d_internal_guidance <-
-  d_coding %>% 
+<br>
+
+```{r statguidefig, fig.cap = "Percentage of journals offering statistical guidance by scientific discipline (N = 15; represented by coloured dots) and overall (N = 330; represented by black diamond).
+Graphs ordered from left to right and top to bottom in order of highest proportion overall across scientific disciplines.", fig.topcaption = TRUE, fig.width = 10, fig.height = 6}
+# extract overall data
+sum_tab_general_overall <- sum_tab_general %>%
+  filter(variable %in% c('guidance', 'internal_guidance_section')) %>%
+  filter(esi_field == "OVERALL")
+
+sum_tab_general %>%
+  filter(esi_field != "OVERALL") %>%
+  filter(variable %in% c('guidance', 'internal_guidance_section')) %>%
+  ggplot(aes(x=percent, y = 1, colour = str_wrap(esi_field_with_abbr,25))) +
+  facet_rep_wrap(facets = vars(var_display), ncol = 2, nrow = 1) +
+  coord_capped_cart(bottom='both', left='both') + # to break axes in bottom left corner
+  geom_point(size = 5, position = position_stack(), alpha = .5) +
+  geom_point(size = 6, position = position_stack(), data = sum_tab_general_overall, colour = 'black', shape = 18) +
+  geom_text_repel(
+    aes(label = ifelse(percent > 0,esi_abbr,''),
+        segment.square = T,
+        segment.inflect = T), #
+    #point.padding = 0.2,
+    force_pull   = 0, # do not pull toward data points
+    nudge_y      = 15,
+    #nudge_x = 5,
+    direction    = "x",
+    angle        = 90,
+    hjust        = 0,
+    segment.size = 0.2,
+    segment.curvature = -1e-20,
+    max.iter = 1e4, max.time = 1,
+    max.overlaps = Inf, # ensure all labels are shown even if overlapping
+    show.legend = FALSE # do not add legend entry
+  ) +
+  scale_y_continuous(limits = c(0,15), name = '1 dot per scientific discipline') +
+  xlim(0,100) +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid.major.y = element_blank(),
+        legend.position = 'bottom') +
+  scale_colour_discrete(name = 'Scientific disciplines') +
+  guides(colour = guide_legend(nrow = 6, title.position = 'top'))
+```
+
+<br><br>
+
+Figure \@ref(fig:stattopicsfig) shows the percentage of journals in each scientific discipline (denominator is N = 15) and overall (denominator is N = 330) that mentioned each of the twenty pre-specified statistical topics.
+For tabular data, see the Appendix (Section \@ref(app)).
+Please see Section \@ref(issues) (issues) at the end of this document for an important note regarding the topic "prespecification of analyses".
+
+<br>
+
+```{r stattopicsfig, fig.cap = "Percentage of journals offering guidance on twenty statistical topics by scientific discipline (N = 15; represented by coloured dots) and overall (N = 330; represented by black diamond).
+For presentational purposes, disciplines offering no guidance on individual topics are shown, but not labelled.
+Two scientific disciplines in which no journals offered any statistical guidance at all are not shown.
+Graphs ordered from left to right and top to bottom in order of highest proportion overall across scientific disciplines.", fig.topcaption = TRUE, fig.width = 10, fig.height = 28}
+# extract overall data
+sum_tab_topics_overall <- sum_tab_topics %>%
+  filter(esi_field == "OVERALL")
+
+sum_tab_topics %>%
+  filter(esi_field != "OVERALL") %>%
+  ggplot(aes(x=percent, y = 1, colour = str_wrap(esi_field_with_abbr,25))) +
+  facet_rep_wrap(facets = vars(topic_display), ncol = 2) +
+  coord_capped_cart(bottom='both', left='both') + # to break axes in bottom left corner
+  geom_point(size = 5, position = position_stack(), alpha = .5) +
+  geom_point(size = 6, position = position_stack(), data = sum_tab_topics_overall, colour = 'black', shape = 18) +
+  geom_text_repel(
+    aes(label = ifelse(percent > 0,esi_abbr,''),
+        segment.square = T,
+        segment.inflect = T), #
+    #point.padding = 0.2,
+    force_pull   = 0, # do not pull toward data points
+    nudge_y      = 25,
+    nudge_x = 10,
+    direction    = "x",
+    angle        = 90,
+    hjust        = 0,
+    segment.size = 0.2,
+    segment.curvature = -1e-20,
+    max.iter = 1e4, max.time = 1,
+    max.overlaps = Inf, # ensure all labels are shown even if overlapping
+    show.legend = FALSE # do not add legend entry
+  ) +
+  scale_y_continuous(limits = c(0,25), name = '1 dot per scientific discipline') +
+  xlim(0,100) +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        panel.grid.major.y = element_blank(),
+        legend.position = 'bottom') +
+  scale_colour_discrete(name = 'Scientific disciplines') +
+  guides(colour = guide_legend(nrow = 5, title.position = 'top'))
+```
+
+<br>
+
+You can see that some topics (e.g., confidence intervals, p-values) are much more likely to be mentioned than other topics (e.g., handling outliers, categorisation of continuus data).
+Journals in health-related fields, particularly clinical medicine, are often more likely to mention these topics than journals in other fields.
+
+```{r how_much_guidance}
+n_topics <- d_all %>% filter(has_guidance == T, has_internal_guidance == T) %>%
+
+  # Recode NAs to FALSE
+  mutate(across(
+    everything()
+    , ~replace_na(., FALSE) )) %>% 
+  select(journal, starts_with("has_"), -c(has_guidance, has_internal_guidance, has_internal_guidance_section, has_external_guidance, has_publisher_guidance)) %>%
+  pivot_longer(cols = !journal, names_to = 'topic', values_to = 'value') %>%
+  group_by(journal) %>%
+  summarise(topics_mentioned = sum(value))
+```
+
+<br>
+
+For the `r d_all %>% filter(has_guidance == T, has_internal_guidance == T) %>% nrow()` journals that offered some internal guidance (i.e., excluding those that offered no guidance at all and those that only referred to external sources), the histogram in Figure \@ref(fig:topicshist) illustrates that the maximum number of topics mentioned by an individual journal was `r max(n_topics$topics_mentioned)`.
+There were `r n_topics %>% filter(topics_mentioned == 0) %>% nrow()` journals that did not mention any of our prespecified topics (but provided statistical guidance on other topics).
+The median topics mentioned was `r median(n_topics$topics_mentioned)`.
+
+<br>
+
+```{r topicshist, fig.cap = 'Histogram showing how many of the twenty statistical topics were mentioned by each journal'}
+n_topics %>%
+  ggplot(aes(x=topics_mentioned)) +
+  geom_histogram()
+```
+
+<br>
+
+In the grid below (Figure \@ref(fig:guidancemosaic)), you can see which of the twenty topics were mentioned by each journal.
+Journals that shared publisher guidance are represented by one row (per publisher).
+Journals that did not offer any statistical guidance at all, or journals that referred only to guidance in external sources, are not shown.
+
+<br>
+
+```{r}
+my_palette <- c("#56B4E9", "#E69F00")
+
+d_internal_guidance_noPublisher <-
+  d_all %>%
+  filter(has_internal_guidance == T, # select journals with internal guidance
+         has_publisher_guidance == F | journal == "Scientific Data") # remove journals with shared publisher guidance except for Scientific Data (which shared guidance and its own guidance)
+
+# were going to represent journals with publisher guidance in a single row per publisher. 
+# select exemplar journals for each publisher
+publisher_nature <- d_all %>% filter(journal == "Nature") %>% mutate(journal = paste0("Nature journals"," (n=", publisher_nature_n,')'))
+publisher_cell <- d_all %>% filter(journal == "CELL") %>% mutate(journal = paste0("Cell journals"," (n=", publisher_cell_n,')'))
+publisher_frontiers <- d_all %>% filter(journal == "Frontiers in Microbiology") %>% mutate(journal = paste0("Frontiers journals"," (n=", publisher_frontiers_n,')'))
+
+# combine the journals with no publisher guidance with the publisher guidance rows
+d_internal_guidance <- bind_rows(d_internal_guidance_noPublisher, publisher_nature, publisher_cell, publisher_frontiers)
+
+# list of journals in order of most topics mentioned
+journals_topics_mentioned_ordered <- d_internal_guidance %>%
+  # Recode NAs to FALSE
+  mutate(across(
+    everything()
+    , ~replace_na(., FALSE) )) %>% 
+  select(journal, starts_with("has_"), -c(has_guidance, has_internal_guidance, has_internal_guidance_section, has_external_guidance, has_publisher_guidance)) %>%
+  pivot_longer(cols = !journal, names_to = 'topic', values_to = 'value') %>%
+  group_by(journal) %>%
+  summarise(topics_mentioned = sum(value)) %>%
+  arrange(topics_mentioned) %>%
+  pull(journal)
+
+plt <- d_internal_guidance %>%
+  mutate(journal = factor(journal, levels = journals_topics_mentioned_ordered)) %>% # order journals by n topics mentioned
   select(journal, starts_with("has_") & (!contains("guidance"))) %>%
+  mutate(journal = fct_recode(journal,"PNAS" = "PROCEEDINGS OF THE NATIONAL ACADEMY OF SCIENCES OF THE UNITED STATES OF AMERICA")) %>% # shorten a particularly long journal name
   pivot_longer(
@@
     names_to = "topic",
+    names_prefix = 'has_',
     values_to = "outcome",
@@
     TRUE ~ "Not applicable (No internal guidance)"
-  ))
-
-# plt <- 
-  d_internal_guidance %>%
-  ggplot(aes(x = topic, y = factor(journal))) +
-  geom_tile(aes(fill = outcome), colour = 'black', size = .25) +
+  )) %>%
+  mutate(topic_display = case_when( # create display versions of topic names
+    topic == "guidance" ~ "Provides any statistical guidance",
+    topic == "internal_guidance" ~ "Provides journal-specific statistical guidance",
+    topic == "internal_guidance_section" ~ "Has a dedicated statistical guidance section",
+    topic == "external_guidance" ~ "Refers to external statistical guidance",
+    topic == "p_value" ~ "p values",
+    topic == "significance" ~ "statistical significance",
+    topic == "null_hypo" ~ "null hypotheses",
+    topic == "sample_size" ~ "sample size justification",
+    topic == "conf_int" ~ "confidence intervals",
+    topic == "effect_size" ~ "effect sizes",
+    topic == "multi_compare" ~ "multiple comparisons",
+    topic == "subgroup" ~ "subgroup analyses",
+    topic == "baseline_covar" ~ "baseline covariates",
+    topic == "non_param" ~ "non-parametric tests",
+    topic == "sensitivity" ~ "sensitivity analyses",
+    topic == "model_assume" ~ "checking model assumptions",
+    topic == "exclusion" ~ "data exclusions",
+    topic == "outliers" ~ "handling outliers",
+    topic == "missing" ~ "handling missing data",
+    topic == "one_sided" ~ "one-sided tests",
+    topic == "bayes" ~ "Bayesian statistics",
+    topic == "secondary" ~ "secondary outcomes",
+    topic == "prespecify" ~ "prespecification of analyses",
+    topic == "cat_continuous" ~ "categorisation of continuous data",
+    topic == "publisher_guidance" ~ "Shares publisher guidance"
+  )) %>%
+  ggplot(aes(x = topic_display, y = journal)) +
+    geom_tile(aes(fill = outcome), colour = 'black', size = .25, alpha = .6) +
     scale_fill_manual(
@@
     scale_x_discrete(position = 'top') +
-  papaja::theme_apa() +
-  xlab("Topics mentioned") +
+    scale_y_discrete(labels = wrap_format(25)) + # wrap text on y axis
+    xlab("Statistical topics") +
     theme(plot.margin = margin(0, 50, 0, 0),
@@
           axis.title.y =element_blank(),
+          axis.line = element_blank(),
           panel.grid.major=element_line(colour=NA),
           panel.grid.minor=element_line(colour=NA),
-        legend.position = 'bottom',
+          legend.position = 'top',
           axis.text.x  = element_text(angle = 45, hjust = 0),
@@
 
-Try bar chart
-
-```{r plot_internal_guidance_bar}
-d_internal_guidance %>% 
-  ggplot(aes(y = topic, fill = outcome)) +
-  geom_bar() +
-  scale_fill_manual(values = my_palette)
+```{r guidancemosaic, fig.cap = "Grid diagram showing whether each journal provdied guidance on each of twenty prespecified statistical topics", fig.height = 35, fig.width = 10}
+plt
 ```
@@
 
-## Issues to address
+# Issues to address {#issues}
 
-* Pre-registration including general clinical trials advice
-* Coding of external guidance
+* Our coding for the topic "prespecification of analyses" was not consistent across coders.
+Specifically, some coders considered guidance/instructions related to registration of clinical trials to be covered by this topic, whilst others did not.
+I believe the justification of those who did not was that clinical trials registration does not necessarily involve prespecification of a statistical analysis plan.
+That seems a reasonable justification to me, but I'd be interested to hear thoughts from the team.
+If we decide that advice about clinical trials registration alone does not fall into the remit of this topic, then it is relatively straightforward to examine the verbatim text we extracted and re-code the relevant cases.
+However, if we decide that advice about clinical trials registration *does* fall under the remit of this topic, then we will need to re-visit the journal instructions to authors and search for such guidance (as this was not done consistently previously).
+
+* We have recorded whether journals refer to statistical guidance from external sources, such as reporting guidelines or academic papers.
+However, we have not examined those external sources, and we need to decide whether to do so.
+If we perform extraction and coding for all `r lookup_external_guidance %>% nrow()` external sources we have identified, this is clearly a lot of extra work.
+We also perhaps only really interested in the most salient statistical guidance offered by journals.
+On the other hand, if we do not extract information from external sources, then perhaps we are missing an important aspect of statistical guidance i.e., that journals may not provide their own guidance on certain topics because they are aptly covered by external sources.
+An additional complication is that in some cases, the boundary of internal vs external guidance is blurred when the 'external' source is an academic article that appears to have been written by members of the journal's editorial team.
+
+# Appendix {#app}
+
+Tabular data (represented in Figures 1 and 2). There is a scroll bar at the bottom of the table to see all of the scientific fields.
+
+```{r tbl_counts_props}
+d_all %>%
+
+  # Recode NAs to FALSE
+  mutate(across(
+    everything()
+    # starts_with("has_") & (!contains("guidance")) # for internal guidance elements
+    , ~replace_na(., FALSE) )) %>%
+
+  select(esi_field, starts_with("has_")) %>% 
+  gtsummary::tbl_summary(
+    by = esi_field,
+      # statistic = list(everything() ~ "{p}% ({n} / {N})")
+  ) %>%
+  add_overall %>% 
+  modify_footnote(everything() ~ NA) %>% 
+  modify_caption("**Number and percentage of journals overall and by scientific discipline providing internal statistical guidance by topic.** Denominator for all proportions is the number of journals in the field, regardless of whether the journal has internal guidance.")
+```
+
+```{r}
+#### CODE GRAVEYARD - BEWARE ####
+
+# try binning method
+# forPlot %>%
+#   #mutate(bin = cut_width(prop, width = 0.2, center = 0.1)) %>%
+#   mutate(bin = cut_width(percent, width = 20, center = 10)) %>%
+#   group_by(esi_field,bin) %>%
+#   summarise(n = n()) %>%
+#   group_by(bin) %>%
+#   summarise(esi_field = esi_field, pos = cumsum(n)) %>%
+#   merge(forPlot, by = "esi_field") %>% # rejoin columns from original dataframe
+#   ggplot(aes(x=percent, y = pos, colour = esi_field)) +
+#   geom_vline(xintercept = sum_tab %>% filter(variable == "guidance", esi_field == "OVERALL") %>% pull(percent)) +
+#   geom_point(size = 5) +
+#   geom_label_repel(aes(label = esi_abbr)) +
+#   theme_minimal() +
+#   theme(axis.title.y = element_blank(),
+#         axis.text.y = element_blank(),
+#         axis.ticks.y = element_blank()) +
+#   scale_colour_discrete()
+
+# # trying out a dot histogram
+# d <- sum_tab %>%
+#   #filter(variable %in% c('guidance', 'internal_guidance', 'internal_guidance_section', 'external_guidance')) %>%
+#   filter(variable %in% c('guidance'))
+# 
+# sum_tab %>%
+#   filter(variable %in% c('guidance', 'internal_guidance', 'internal_guidance_section', 'external_guidance')) %>%
+#   ggplot(aes(x = prop, fill=esi_field, label = esi_abbr, group = esi_field)) + 
+#     #geom_histogram(aes(fill=esi_field), binwidth=0.1, colour="grey20", lwd=0.2) +
+#     geom_dotplot(binwidth=0.1, colour="grey20", lwd=0.2, stackgroups = T, binpositions = 'all', method = 'histodot', dotsize = 0.25, stackdir = 'center') +
+#     facet_grid(rows = vars(variable)) +
+#     theme(axis.title.y = element_blank(), # y axis labels are meaningless on this plot, remove them
+#       axis.text.y = element_blank(),
+#       legend.position = 'bottom')
+# 
+#   # looks ok... but difficult to tell which field is which (too many similar colours, and have to keep looking at legend)
+#   # could we put the field abbreviation text inside the dots? (maybe shorten to two letters)
+#   # also for OVERALL we need something different, e.g., a black vertical line
+```

### 2021-08-18 analysis/preliminaryAnalysis.Rmd
added some guidance examples

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = F, message = F, warning = T)
+knitr::opts_chunk$set(echo = F, message = F, warning = T, fig.topcaption = TRUE)
 ```
@@
 
-```{r set_style_defaults}
+```{r set_defaults}
 theme_gtsummary_compact() # set table theme
@@
 
+```{r guidance_examples}
+# get the names of the twenty statistical topics
+topic_names <- d_all %>% 
+  select(starts_with("has_"), -c(has_guidance, has_external_guidance, has_internal_guidance, has_internal_guidance_section, has_publisher_guidance)) %>% 
+  rename_all(~str_replace(.,"^has_","")) %>% # remove has_ prefix
+  verify(ncol(.) == 20) %>% # check there are twenty columns (topics) as expected
+  colnames()
+
+set.seed(42)
+guidance_examples_tbl <- d_all %>%
+  select(journal, topic_names) %>% 
+  verify(ncol(.) == 21) %>% # check there are 21 columns as expected (i.e., journal and twenty topics)
+  pivot_longer(cols = !journal, names_to = 'topic', values_to = 'verbatim') %>%
+  distinct(verbatim, .keep_all = T) %>%
+  drop_na(verbatim) %>%
+  filter(nchar(verbatim) < 800) %>% # select shorter guidance only
+  group_by(topic) %>%
+  sample_n(2) %>%
+  ungroup() %>%
+  mutate(verbatim = paste0(verbatim,' (',journal,')')) %>% # add journal name to verbatim column
+  select(-journal) %>% # remove journal column
+  mutate(topic_display = case_when( # create display versions of topic names
+    topic == "p_value" ~ "p values",
+    topic == "significance" ~ "statistical significance",
+    topic == "null_hypo" ~ "null hypotheses",
+    topic == "sample_size" ~ "sample size justification",
+    topic == "conf_int" ~ "confidence intervals",
+    topic == "effect_size" ~ "effect sizes",
+    topic == "multi_compare" ~ "multiple comparisons",
+    topic == "subgroup" ~ "subgroup analyses",
+    topic == "baseline_covar" ~ "baseline covariates",
+    topic == "non_param" ~ "non-parametric tests",
+    topic == "sensitivity" ~ "sensitivity analyses",
+    topic == "model_assume" ~ "checking model assumptions",
+    topic == "exclusion" ~ "data exclusions",
+    topic == "outliers" ~ "handling outliers",
+    topic == "missing" ~ "handling missing data",
+    topic == "one_sided" ~ "one-sided tests",
+    topic == "bayes" ~ "Bayesian statistics",
+    topic == "secondary" ~ "secondary outcomes",
+    topic == "prespecify" ~ "prespecification of analyses",
+    topic == "cat_continuous" ~ "categorisation of continuous data"
+  )) %>%
+  select(topic = topic_display,"example guidance" =verbatim)
+```
+
+To give you a flavour of the kind of guidance provided, I have randomly selected two examples (excluding a few very long examples), for each of the twenty statistical topics, and displayed them in Table \@ref(fig:guidanceexamples) below.
+
+```{r guidanceexamples}
+guidance_examples_tbl %>% kable()
+```
+
 Look at external guidance
@@
 ```
+
+```{r phase_two}
+# export data for phase two
+d_all %>% filter(has_p_value == T | has_significance == T) %>% 
+  mutate(coder = "TEH") %>% 
+  select(journal, esi_field, p_value, significance) %>%
+  mutate(recommended_required = NA,
+         p_how = NA,
+         p_threshold = NA,
+         stat_sig_ok = NA,
+         stat_sig_alpha = NA) %>%
+  write_csv(here('data','raw','templates','phase_two_pval.csv'))
+
+```

### 2021-08-24 analysis/preliminaryAnalysis.Rmd
Create gtsummary demo

@@
 library(here) # for finding files
-library(gtsummary) # for summary tables
+library(gtsummary) # for summary tables note: use dev version `remotes::install_github("ddsjoberg/gtsummary")`
 library(lemon) # to have axes across all graph facets
@@
 ```{r tbl_counts_props}
+tbl_counts_props <-
 d_all %>%
@@
   modify_caption("**Number and percentage of journals overall and by scientific discipline providing internal statistical guidance by topic.** Denominator for all proportions is the number of journals in the field, regardless of whether the journal has internal guidance.")
+
+tbl_counts_props
 ```
+
+## {gtsummary} in-text demo
+
+> Overall stats are in the `stat_0` column, which is not accessible in the current release of {gtsummary}, so hence use the dev version: `remotes::install_github("ddsjoberg/gtsummary")`.
+
+You can see that around half (`r inline_text(tbl_counts_props, variable = has_guidance, column = stat_0, pattern = "{p}")`%) of journals offered any statistical guidance.
+Note that this includes `r only_ext` journals which only referred to statistical guidance in external sources (reporting guidelines or academic papers).
+Just over quarter of journals had a dedicated statistical guidance section of their author instructions (`r inline_text(tbl_counts_props, variable = has_internal_guidance_section, column = stat_0, pattern = "{p}")`%).
+In two fields (Computer Science and Maths), no journals offered any statistical guidance.
+Journals in health-related fields were more likely to offer statistical guidance and have dedicated statistical guidance sections.
+Notably, 100% of the journals in clinical medicine offered some statistical guidance.
 

### 2021-09-01 analysis/preliminaryAnalysis.Rmd
removing gtsummary

--- 2021-08-18 analysis/preliminaryAnalysis.Rmd
@@
 library(here) # for finding files
-library(gtsummary) # for summary tables
 library(lemon) # to have axes across all graph facets
@@
 ```{r set_defaults}
-theme_gtsummary_compact() # set table theme
 ggthemr('fresh') # set ggplot theme
@@
 
-**Note:** When journals referred to guidance in external sources, we have recorded the names of the sources, but *we have not currently extracted guidance from those sources*.
+**Note:** When journals referred to guidance in external sources, we have recorded the names of the sources, but *we have not extracted guidance from those sources*.
 We return to this point in the *issues to address* section (Section \@ref(issues)) at the end of this document.
@@
 Shared publisher guidance was coded by two team members (as above), but is represented in our data and analysis multiple times - once for each individual journal it applies to.
-For example, the Cell journals guidance is inherited by `r publisher_cell_n` journals so appears `r publisher_cell_n` times in the dataset.
+For example, the Cell journals' guidance is inherited by `r publisher_cell_n` journals so appears `r publisher_cell_n` times in the dataset.
 Unless stated otherwise, publisher guidance is treated as journal-specific guidance (i.e., not external guidance) in the analyses presented below.
@@
 ```{r summarise_data}
-
-# couldn't figure out to extract summary data from gtsummary::tblsummary so making my own summary dataframe
 sum_tab <- d_all %>%
@@
   geom_point(size = 5, position = position_stack(), alpha = .5) +
-  geom_point(size = 6, position = position_stack(), data = sum_tab_general_overall, colour = 'black', shape = 18) +
+  geom_point(size = 6, position = position_stack(), data = sum_tab_general_overall, shape = 18, colour = 'black', aes(fill = 'OVERALL')) +
   geom_text_repel(
@@
         legend.position = 'bottom') +
-  scale_colour_discrete(name = 'Scientific disciplines') +
-  guides(colour = guide_legend(nrow = 6, title.position = 'top'))
+  guides(colour = guide_legend(nrow = 8, title.position = 'top', title = '')) +
+  guides(fill = guide_legend(nrow = 1, title.position = 'top', title = "Scientific disciplines"))
 ```
@@
   geom_point(size = 5, position = position_stack(), alpha = .5) +
-  geom_point(size = 6, position = position_stack(), data = sum_tab_topics_overall, colour = 'black', shape = 18) +
+  geom_point(size = 6, position = position_stack(), data = sum_tab_topics_overall, shape = 18, colour = 'black', aes(fill = 'OVERALL')) +
   geom_text_repel(
@@
         legend.position = 'bottom') +
-  scale_colour_discrete(name = 'Scientific disciplines') +
-  guides(colour = guide_legend(nrow = 5, title.position = 'top'))
+  guides(colour = guide_legend(order = 1, nrow = 8, title.position = 'top', title = '')) +
+  guides(fill = guide_legend(order = -1, nrow = 1, title.position = 'top', title = "Scientific disciplines"))
 ```
@@
 ```{r tbl_counts_props}
-d_all %>%
-
-  # Recode NAs to FALSE
-  mutate(across(
-    everything()
-    # starts_with("has_") & (!contains("guidance")) # for internal guidance elements
-    , ~replace_na(., FALSE) )) %>%
-
-  select(esi_field, starts_with("has_")) %>% 
-  gtsummary::tbl_summary(
-    by = esi_field,
-      # statistic = list(everything() ~ "{p}% ({n} / {N})")
-  ) %>%
-  add_overall %>% 
-  modify_footnote(everything() ~ NA) %>% 
-  modify_caption("**Number and percentage of journals overall and by scientific discipline providing internal statistical guidance by topic.** Denominator for all proportions is the number of journals in the field, regardless of whether the journal has internal guidance.")
+sum_tab %>%
+  mutate(field = paste0(esi_field,' N = ',N),
+         table_data = paste0(n,' (',percent,'%)')) %>%
+  select(field,variable,table_data) %>%
+  pivot_wider(id_cols = variable, names_from = field, values_from = table_data) %>%
+  select(variable, `OVERALL N = 330`, everything()) %>%
+  kable(caption = "**Number and percentage of journals overall and by scientific discipline providing internal statistical guidance by topic.** Denominator for all proportions is the number of journals in the field, regardless of whether the journal has internal guidance.")
 ```
@@
   mutate(recommended_required = NA,
-         p_how = NA,
-         p_threshold = NA,
-         stat_sig_ok = NA,
+         p_use = NA,
+         p_reporting = NA,
+         stat_sig_use = NA,
          stat_sig_alpha = NA) %>%
   write_csv(here('data','raw','templates','phase_two_pval.csv'))
-
 ```

### 2021-09-02 analysis/preliminaryAnalysis.Rmd
preliminary analysis 1st full draft

@@
 ---
-title: "An assessment of statistical guidance for authors at influential academic journals across scientific disciplines: Preliminary analysis"
-author: "Analysis performed by: Maia Salholz-Hillel & Tom Hardwicke"
-date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
+title: "Cross-disciplinary assessment of statistical guidance provided to authors by influential academic journals"
+description: |
+  Preliminary analysis for study one and study two.
+author:
+  - name: Maia Salholz-Hillel & Tom Hardwicke
+date: "`r Sys.Date()`"
 output: 
-  bookdown::html_document2:
-    fig_caption: true
-    number_sections: true
-    global_numbering: true
-editor_options: 
-  chunk_output_type: console
+  distill::distill_article:
+    toc: true
 ---
@@
 library(knitr) # for literate programming
-library(papaja) # for article template
 library(tidyverse) # for data munging
-library(tidylog) # inline analysis feedback
+library(janitor) # for data munging
+library(tidylog) # inline code feedback
 library(here) # for finding files
+library(assertr) # for testing
 library(lemon) # to have axes across all graph facets
@@
 library(scales) # to wrap text on axis labels
+library(rmarkdown) # for paged table
 ```
@@
 ```{r load_functions}
-# load custom functions
-source(here('analysis','functions.R'))
+source(here('analysis','functions.R')) # load custom functions
 ```
@@
 ```{r perform_preprocessing}
-# loads raw data, performs preprocessing, saves preprocessed files
-source(here('analysis','preprocessing.R'))
+# loads raw data, performs preprocessing, saves processed data files
+#source(here('analysis','preprocessing.R'))
 ```
@@
 ```{r load_processed_data}
-# loads the processed data files output by the preprocessing performed above
+# loads the processed data files
 load(here('data','processed','d_all.rds'))
@@
 # Introduction
-<br>
-We have now completed data collection for the first phase of this project.
-We extracted author-facing statistical guidance provided by the top-15 journals (by Impact Factor) in each of 22 scientific disciplines (i.e., total N = 330 journals).
+
+We have now completed dual-coded data collection for the first study of this project and single-coded data collection for the second study of this project.
 The preregistered study protocol is available [here](https://osf.io/cz9g3/).
-Below we present a preliminary analysis of the phase one data and highlight some issues to discuss as a team.
+Below we present a preliminary analysis of the data and highlight some issues to discuss as a team.
 
@@
 
-> **Operational definition of statistical guidance**: any advice or instruction related to the appropriate selection, implementation, reporting, or interpretation of statistical analyses.
+## Study one
 
-We recorded whether each journal:
+In November, 2019, we made copies of any documentation (e.g., instructions to authors) that provided author-facing statistical guidance provided by the top-15 journals (ranked by Impact Factor) in each of 22 scientific disciplines (i.e., total N = 330 journals).
+
+**Operational definition of statistical guidance**
+
+> any advice or instruction related to the appropriate selection, implementation, reporting, or interpretation of statistical analyses.
+
+We then examined this documentation and recorded whether each journal:
 
@@
 
-* We coded (YES/NO) whether it mentioned each of 20 statistical topics (e.g., "p-values"), and extracted the verbatim text.
-
-**Note:** When journals referred to guidance in external sources, we have recorded the names of the sources, but *we have not extracted guidance from those sources*.
-We return to this point in the *issues to address* section (Section \@ref(issues)) at the end of this document.
+* We also recorded whether it mentioned each of 20 statistical topics (e.g., "p-values"), and extracted the verbatim text.
 
 All data extraction was performed by a first coder (DS, MM, TB, MSH) and a second coder (TEH) with any coding differences resolved through discussion.
+
+**Note:** When journals referred to guidance in external sources (typically reporting guidelines), we have recorded the names of the sources, but *we have not extracted guidance from those sources*.
 
@@
 **Note:** We found that some publishers provided statistical guidance that was shared across their portfolio of journals (specifically, `r publisher_nature_n` Nature journals, `r publisher_cell_n` Cell journals, and `r publisher_frontiers_n` Frontiers journals).
+An additional journal - Scientific Data - shares guidance with other Nature journals, but also has its own specific guidance.
 Shared publisher guidance was coded by two team members (as above), but is represented in our data and analysis multiple times - once for each individual journal it applies to.
-For example, the Cell journals' guidance is inherited by `r publisher_cell_n` journals so appears `r publisher_cell_n` times in the dataset.
-Unless stated otherwise, publisher guidance is treated as journal-specific guidance (i.e., not external guidance) in the analyses presented below.
+For example, the `r publisher_cell_n` journals published by Cell inherit Cell's "STAR Methods" guidance, so this guidance appears `r publisher_cell_n` times in our dataset.
+
+## Study two
+
+For study two, we planned to more closely examine the guidance provided on five specific topics:
+
+- p-values
+- statistical significance
+- confidence intervals
+- effect sizes
+- sample size justification
+
+We prespecified some information we wanted to extract about these topics; however, when I (Tom) examined the verbatim guidance, I found our prespecified questions and response options to be unsuitable (ambiguous, badly worded, etc).
+So in an entirely data-dependent and exploratory manner, I instead coded the level of endorsement for each topic/method according to four categories:
+
+1. Explicit endorsement - the journal advises or instructs authors to use this method whenever possible/appropriate.
+2. Implicit endorsement - the journal provides advice on the method, implying endorsement, but does not explictly advise that the method should be used.
+3. Implicit opposition - the journal advises that they would prefer the method is not used, but does not explicitly rule it out.
+4. Explicit opposition - the journal advises that the method should not be used.
+
+This was admittedly quite subjective and I had to 'read between the lines' a bit in order to assign guidance to these categories.
+
+I also categorised reporting advice for three topics as below:
+
+p-values:
+
+1. Report exact p-values
+2. Report exact p-values unless very small
+3. Report p-values relative to thresholds (e.g., p < .05)
+4. No guidance
+
+Statistical significance:
+
+1. Report alpha
+2. Use specific alpha (e.g., a = .05)
+3. No guidance
+
+Confidence intervals:
+
+1. Report confidence level
+2. Use specific confidence level (e.g., 95%)
+3. No guidance
+
+So far, the data coding for study two has only been performed by me (Tom), though we originally planned for dual-coding (see issues section at the end of this document).
 
 # Results
+
+## Study one
 
@@
     everything()
-    # starts_with("has_") & (!contains("guidance")) # for internal guidance elements
     , ~replace_na(., FALSE) )) %>% 
-
   select(esi_field, starts_with("has_")) %>% 
@@
   mutate(prop = n/N, percent = round(prop*100, 0)) %>%
-  mutate(esi_abbr = case_when( # create abbreviated field labels
+  mutate(esi_abbr = case_when( # create abbreviated field labels to use in figures
     esi_field == "OVERALL" ~ "OVERALL",
@@
 
-# select only data on general any/internal/external statistical guidance classifications 
+# select only variables that address the general any/internal/external statistical guidance classifications 
 sum_tab_general <- sum_tab %>%
@@
 
-# select only data on topics
+# select only variables on specific topics
 sum_tab_topics <- sum_tab %>%
@@
 
-Figure \@ref(fig:statguidefig) shows the percentage of journals in each scientific discipline (denominator is N = 15) and overall (denominator is N = 330) that (a) offered any statistical guidance; (b) had a dedicated statistical guidance section.
-For tabular data, see Appendix \@ref(app).
+Figure \@ref(fig:statguidefig) shows the percentage of journals in each scientific discipline (denominator is N = 15) and overall (denominator is N = 330) that:
+
+(a) offered any statistical guidance; 
+(b) had a dedicated statistical guidance section.
+
+For tabular data, see the table in the Appendix.
 
@@
 Note that this includes `r only_ext` journals which only referred to statistical guidance in external sources (reporting guidelines or academic papers).
-Just over quarter of journals had a dedicated statistical guidance section of their author instructions (`r sum_tab %>% filter(esi_field == "OVERALL", variable == "internal_guidance_section") %>% pull(percent)`%).
+Just over quarter of journals had a dedicated statistical guidance section in their author instructions (`r sum_tab %>% filter(esi_field == "OVERALL", variable == "internal_guidance_section") %>% pull(percent)`%).
 In two fields (Computer Science and Maths), no journals offered any statistical guidance.
@@
 
-<br>
-
-```{r statguidefig, fig.cap = "Percentage of journals offering statistical guidance by scientific discipline (N = 15; represented by coloured dots) and overall (N = 330; represented by black diamond).
-Graphs ordered from left to right and top to bottom in order of highest proportion overall across scientific disciplines.", fig.topcaption = TRUE, fig.width = 10, fig.height = 6}
+```{r}
 # extract overall data
@@
   filter(esi_field == "OVERALL")
+```
 
+```{r statguidefig, layout="l-page", fig.width = 10, fig.height = 6, fig.cap = "Percentage of journals offering statistical guidance by scientific discipline (N = 15; represented by coloured dots) and overall (N = 330; represented by black diamond)."}
 sum_tab_general %>%
@@
 ```
+Figure \@ref(fig:stattopicsfig) shows the percentage of journals in each scientific discipline (denominator is N = 15) and overall (denominator is N = 330) that mentioned each of the twenty statistical topics we pre-selected.
+For tabular data, see the table in the Appendix.
+Please see the issues section at the end of this document for an important note regarding the topic "prespecification of analyses".
 
-<br><br>
+You can see that some topics (e.g., confidence intervals, p-values) are much more likely to be mentioned than other topics (e.g., handling outliers, categorisation of continuous data).
+Journals in health-related fields, particularly clinical medicine, are often more likely to mention this collection of topics than journals in other fields.
 
-Figure \@ref(fig:stattopicsfig) shows the percentage of journals in each scientific discipline (denominator is N = 15) and overall (denominator is N = 330) that mentioned each of the twenty pre-specified statistical topics.
-For tabular data, see the Appendix (Section \@ref(app)).
-Please see Section \@ref(issues) (issues) at the end of this document for an important note regarding the topic "prespecification of analyses".
-
-<br>
-
-```{r stattopicsfig, fig.cap = "Percentage of journals offering guidance on twenty statistical topics by scientific discipline (N = 15; represented by coloured dots) and overall (N = 330; represented by black diamond).
-For presentational purposes, disciplines offering no guidance on individual topics are shown, but not labelled.
-Two scientific disciplines in which no journals offered any statistical guidance at all are not shown.
-Graphs ordered from left to right and top to bottom in order of highest proportion overall across scientific disciplines.", fig.topcaption = TRUE, fig.width = 10, fig.height = 28}
+```{r stattopicsfig, layout="l-page", fig.width = 10, fig.height = 28, fig.cap = "Percentage of journals offering guidance on twenty statistical topics by scientific discipline (N = 15; represented by coloured dots) and overall (N = 330; represented by black diamond).
+For presentational purposes, two scientific disciplines in which no journals offered any statistical guidance at all are not shown.
+Additionally, disciplines offering no guidance on individual topics are shown, but not labelled.
+Graphs are ordered from left to right and top to bottom in order of highest proportion offering topic guidance overall across scientific disciplines."}
 # extract overall data
@@
 
-<br>
-
-You can see that some topics (e.g., confidence intervals, p-values) are much more likely to be mentioned than other topics (e.g., handling outliers, categorisation of continuus data).
-Journals in health-related fields, particularly clinical medicine, are often more likely to mention these topics than journals in other fields.
-
 ```{r how_much_guidance}
@@
 
-<br>
-
 For the `r d_all %>% filter(has_guidance == T, has_internal_guidance == T) %>% nrow()` journals that offered some internal guidance (i.e., excluding those that offered no guidance at all and those that only referred to external sources), the histogram in Figure \@ref(fig:topicshist) illustrates that the maximum number of topics mentioned by an individual journal was `r max(n_topics$topics_mentioned)`.
@@
 The median topics mentioned was `r median(n_topics$topics_mentioned)`.
-
-<br>
 
@@
 ```
-
-<br>
-
-In the grid below (Figure \@ref(fig:guidancemosaic)), you can see which of the twenty topics were mentioned by each journal.
+In the grid below (Figure \@ref(fig:guidancemosaic)), you can see which of the twenty preselected topics were mentioned by each journal.
 Journals that shared publisher guidance are represented by one row (per publisher).
@@
 
-<br>
-
 ```{r}
-my_palette <- c("#56B4E9", "#E69F00")
+my_palette <- c("#95D840FF", "#404788FF")
 
@@
   )) %>%
+  mutate(topic_display = factor(topic_display, levels = ordered_topics)) %>% # order topics by n topics mentioned
   ggplot(aes(x = topic_display, y = journal)) +
@@
 
-```{r guidancemosaic, fig.cap = "Grid diagram showing whether each journal provdied guidance on each of twenty prespecified statistical topics", fig.height = 35, fig.width = 10}
+```{r guidancemosaic, layout="l-page", fig.cap = "Grid diagram showing whether each journal provided guidance on each of twenty preselected statistical topics.
+Topics are ordered from most (left) to least (right) mentioned.
+Journals are ordered from most (top) to least (bottom) preselected topics mentioned.", fig.height = 35, fig.width = 10}
 plt
@@
 
-To give you a flavour of the kind of guidance provided, I have randomly selected two examples (excluding a few very long examples), for each of the twenty statistical topics, and displayed them in Table \@ref(fig:guidanceexamples) below.
+To give you a flavour of the kind of guidance provided, I have randomly selected two examples (excluding a few very long examples), for each of the twenty statistical topics, and displayed them in Table \@ref(tab:guidanceexamples) below.
 
 ```{r guidanceexamples}
-guidance_examples_tbl %>% kable()
+guidance_examples_tbl %>% kable(caption = "Randomly selected verbatim statistical guidance for each of 20 preselected statistical topics.")
 ```
 
-Look at external guidance
+`r d_all %>% filter(has_external_guidance == T) %>% nrow()` journals referred authors to statistical guidance in an external source (reporting guidelines or other sources like academic papers or websites).
+Of these, `r crosstab_int_ext %>% filter(has_external_guidance == T, has_internal_guidance == F) %>% pull(n)` only referred to statistical guidance in external sources, whereas the other `r crosstab_int_ext %>% filter(has_external_guidance == T, has_internal_guidance == T) %>% pull(n)` also provided their own guidance in additional to referring to external sources.
 
-```{r tbl_external_guidance}
+In the histogram (Figure \@ref(fig:repguidehist)) below, you can see how frequently journals referred to different reporting guidelines.
+Then in Table \@ref(tab:tbl-external-other-guidance) you can see all of the other external guidelines (academic papers and websites) that journals referred to.
+None of these external sources that have been examined by us in detail.
+
+```{r repguidehist, layout="l-page", fig.width = 10, fig.cap = 'Histogram showing how frequently journals referred to different reporting guidelines.'}
+ext_guide_ordered <- lookup_external_guidance %>% arrange(desc(n_journals)) %>% pull(external_guidance)
 lookup_external_guidance %>%
-  knitr::kable()
+  filter(guidance_type == "reporting guideline") %>%
+  mutate(external_guidance = factor(str_to_upper(external_guidance), levels = str_to_upper(ext_guide_ordered)),
+         external_guidance = fct_recode(external_guidance, 
+                                        `CONSORT PRO` = "CONSORT PRO EXTENSION",
+                                        `CONSORT Cluster` = "CONSORT EXTENDED GUIDELINES",
+                                        `HuGENet` = "HUMAN GENOME EPIDEMIOLOGY NETWORK (HUGENET) GUIDELINES",
+                                        `FDA Diagnostic Tests` = "FDA GUIDELINES (HTTPS://PERMA.CC/W2EY-MSTA)",
+                                        `CONSORT abstract` = "CONSORT ABSTRACT EXTENSION",
+         NIH = "NIH PRINCIPLES AND GUIDELINES FOR REPORTING PRECLINICAL RESEARCH",
+         NLM = "NLM RESEARCH REPORTING GUIDELINES AND INITIATIVES (HTTPS://WWW.NLM.NIH.GOV/SERVICES/RESEARCH_REPORT_GUIDE.HTML)")) %>%
+  filter(guidance_type == "reporting guideline") %>%
+  ggplot(aes(x=external_guidance, y = n_journals)) +
+  geom_col() +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  ylim(0,100) +
+  ylab('number of journals') +
+  xlab('reporting guideline')
+```
+Other external sources (not reporting guidelines):
+
+```{r tbl-external-other-guidance}
+lookup_external_guidance %>%
+  filter(guidance_type != "reporting guideline") %>%
+  select(-guidance_type) %>%
+  kable(caption = "Frequency of other external sources of statistical guidance (not reporting guidelines)")
 ```
 
-External guidance types
+## Study Two
 
-```{r tbl_external_guidance_types}
-lookup_external_guidance %>% 
-  count(guidance_type) %>% 
-  kable()
+Study two involved a closer look at five preselected statistical topics.
+Table \@ref(tab:s2table) shows how many journals offerred different levels of endorsement for these statistical methods.
+You can see that it was fairly rare to oppose the use of any of these methods.
+A few journals are taking a stand against the use of "statistical significance", but this is rare.
+Few journals are strongly advocating the use of p-values, but the vast majority find them acceptable.
+It was quite common to recommend reporting of effect sizes and confidence intervals.
+
+```{r}
+d_s2 <- read_csv(here('data','primary','data_phase_two.csv'))
 ```
+
+```{r s2table}
+d_s2 %>% select(journal, esi_field, p_use, stat_sig_use, conf_int_use, effect_size_use, sample_size_justification) %>%
+  pivot_longer(cols = p_use:sample_size_justification, names_to = "topic") %>%
+  filter(value != "No guidance") %>%
+  count(topic, value) %>%
+  mutate(value = fct_recode(value, `Implicit opposition` = 'Implicit opposition (for low sample sizes)',
+         `Explicit opposition` = 'Explicit opposition (without prespecified multiplicity corrections)')) %>%
+  pivot_wider(id_cols = topic, names_from = value, values_from = n) %>%
+  mutate(across(everything(), .fns = ~replace_na(.,0))) %>% 
+  mutate(topic = fct_recode(topic,
+                            `confidence intervals` = "conf_int_use",
+                            `effect sizes` = "effect_size_use",
+                            `p-values` = "p_use",
+                            `sample size justification` = "sample_size_justification",
+                            `statistical significance` = "stat_sig_use")) %>%
+  kable(caption = "Number of journals offering degrees of endorsement for five specific statistical methods")
+```
+
+Note that the one journal with implicit opposition to confidence intervals said this was only in the case of small sample sizes.
+And the one journal with explicit opposition to p-values said this was only when there were no pre-specified multiplicity corrections.
+
+Of the `r d_s2 %>% filter(!is.na(p_value)) %>% nrow()` journals offering guidance on p-values, `r d_s2 %>% filter(!is.na(p_value)) %>% count(p_reporting) %>% filter(p_reporting == "Exact") %>% pull(n)` advised reporting of exact p-values, `r d_s2 %>% filter(!is.na(p_value)) %>% count(p_reporting) %>% filter(p_reporting == "Exact unless very small") %>% pull(n)` advised reporting of exact p-values unless they were very small (e.g., < .001), `r d_s2 %>% filter(!is.na(p_value)) %>% count(p_reporting) %>% filter(p_reporting == "Thresholds [0.05, 0.01, 0.001]") %>% pull(n)` advised to use reporting thresholds (e.g., <.05, <.01, <.001), and `r d_s2 %>% filter(!is.na(p_value)) %>% count(p_reporting) %>% filter(p_reporting == "No guidance") %>% pull(n)` offered no guidance on how to report p-values.
+
+Of the `r d_s2 %>% filter(!is.na(significance)) %>% nrow()` journals offering guidance on statistical significance, `r d_s2 %>% filter(!is.na(significance)) %>% count(stat_sig_reporting) %>% filter(stat_sig_reporting == "Report alpha") %>% pull(n)` advised authors report the alpha level they used, `r d_s2 %>% filter(!is.na(significance)) %>% count(stat_sig_reporting) %>% filter(stat_sig_reporting == "0.05") %>% pull(n)` advised using an alpha level of .05, `r d_s2 %>% filter(!is.na(significance)) %>% count(stat_sig_reporting) %>% filter(stat_sig_reporting == "0.01") %>% pull(n)` advised using an alpha level of .01, and `r d_s2 %>% filter(!is.na(significance)) %>% count(stat_sig_reporting) %>% filter(stat_sig_reporting == "No guidance") %>% pull(n)` offered no guidance on reporting statistical significance.
+
+Of the `r d_s2 %>% filter(!is.na(conf_int)) %>% nrow()` journals offering guidance on confidence intervals, `r d_s2 %>% filter(!is.na(conf_int)) %>% count(conf_int_reporting) %>% filter(conf_int_reporting == "Report 95% Cis") %>% pull(n)` advised authors to report 95% confidence intervals, `r d_s2 %>% filter(!is.na(conf_int)) %>% count(conf_int_reporting) %>% filter(conf_int_reporting == "Report confidence level") %>% pull(n)` advised authors report their chosen confidence level, and `r d_s2 %>% filter(!is.na(conf_int)) %>% count(conf_int_reporting) %>% filter(conf_int_reporting == "No guidance") %>% pull(n)` offered no guidance on reporting confidence intervals.
 
@@
 
+* In two fields (Computer Science and Maths), no journals offered any statistical guidance - I do not know these fields well, but I wonder if this indicates that statistical tools are rarely used, in which case one might reasonably suggest we should not even have bothered to examine these fields.
+Any thoughts on this?
+
 * Our coding for the topic "prespecification of analyses" was not consistent across coders.
-Specifically, some coders considered guidance/instructions related to registration of clinical trials to be covered by this topic, whilst others did not.
+Specifically, some coders considered that this topic also covered any guidance/instructions related to registration of clinical trials, whilst others did not.
 I believe the justification of those who did not was that clinical trials registration does not necessarily involve prespecification of a statistical analysis plan.
@@
 If we perform extraction and coding for all `r lookup_external_guidance %>% nrow()` external sources we have identified, this is clearly a lot of extra work.
-We also perhaps only really interested in the most salient statistical guidance offered by journals.
+A counter-argument is that we are only really interested in the most salient statistical guidance offered directly by journals.
 On the other hand, if we do not extract information from external sources, then perhaps we are missing an important aspect of statistical guidance i.e., that journals may not provide their own guidance on certain topics because they are aptly covered by external sources.
@@
 
+* In the protocol we say that for Study 2, data extraction and coding will be in duplicate.
+However, we do caveat this by saying it might be necessary to deviate from the plan depending on our resources.
+I have done all of the first coding, do we have the capacity to do duplicate coding?
+My worry is that it will slow us down considerably, and I am keen to get these results published before they are outdated (the statistical guidance was originally extracted in November, 2019).
+
 # Appendix {#app}
 
-Tabular data (represented in Figures 1 and 2). There is a scroll bar at the bottom of the table to see all of the scientific fields.
+Tabular data (represented in Figures 1 and 2).
+Number and percentage of journals overall and by scientific field providing internal statistical guidance by topic (click the black arrow to scroll through the data for each field).
+The denominator for proportions is the number of journals in the field (N = 15) or overall (N = 330).
 
-```{r tbl_counts_props}
+```{r tbl-counts-props}
 sum_tab %>%
+  filter(variable != "publisher_guidance") %>%
   mutate(field = paste0(esi_field,' N = ',N),
@@
   select(variable, `OVERALL N = 330`, everything()) %>%
-  kable(caption = "**Number and percentage of journals overall and by scientific discipline providing internal statistical guidance by topic.** Denominator for all proportions is the number of journals in the field, regardless of whether the journal has internal guidance.")
-```
-
-```{r}
-#### CODE GRAVEYARD - BEWARE ####
-
-# try binning method
-# forPlot %>%
-#   #mutate(bin = cut_width(prop, width = 0.2, center = 0.1)) %>%
-#   mutate(bin = cut_width(percent, width = 20, center = 10)) %>%
-#   group_by(esi_field,bin) %>%
-#   summarise(n = n()) %>%
-#   group_by(bin) %>%
-#   summarise(esi_field = esi_field, pos = cumsum(n)) %>%
-#   merge(forPlot, by = "esi_field") %>% # rejoin columns from original dataframe
-#   ggplot(aes(x=percent, y = pos, colour = esi_field)) +
-#   geom_vline(xintercept = sum_tab %>% filter(variable == "guidance", esi_field == "OVERALL") %>% pull(percent)) +
-#   geom_point(size = 5) +
-#   geom_label_repel(aes(label = esi_abbr)) +
-#   theme_minimal() +
-#   theme(axis.title.y = element_blank(),
-#         axis.text.y = element_blank(),
-#         axis.ticks.y = element_blank()) +
-#   scale_colour_discrete()
-
-# # trying out a dot histogram
-# d <- sum_tab %>%
-#   #filter(variable %in% c('guidance', 'internal_guidance', 'internal_guidance_section', 'external_guidance')) %>%
-#   filter(variable %in% c('guidance'))
-# 
-# sum_tab %>%
-#   filter(variable %in% c('guidance', 'internal_guidance', 'internal_guidance_section', 'external_guidance')) %>%
-#   ggplot(aes(x = prop, fill=esi_field, label = esi_abbr, group = esi_field)) + 
-#     #geom_histogram(aes(fill=esi_field), binwidth=0.1, colour="grey20", lwd=0.2) +
-#     geom_dotplot(binwidth=0.1, colour="grey20", lwd=0.2, stackgroups = T, binpositions = 'all', method = 'histodot', dotsize = 0.25, stackdir = 'center') +
-#     facet_grid(rows = vars(variable)) +
-#     theme(axis.title.y = element_blank(), # y axis labels are meaningless on this plot, remove them
-#       axis.text.y = element_blank(),
-#       legend.position = 'bottom')
-# 
-#   # looks ok... but difficult to tell which field is which (too many similar colours, and have to keep looking at legend)
-#   # could we put the field abbreviation text inside the dots? (maybe shorten to two letters)
-#   # also for OVERALL we need something different, e.g., a black vertical line
+  paged_table(options = list(rows.print = 25, cols.print = 5))
 ```
@@
 ```{r phase_two}
-# export data for phase two
-d_all %>% filter(has_p_value == T | has_significance == T) %>% 
-  mutate(coder = "TEH") %>% 
-  select(journal, esi_field, p_value, significance) %>%
-  mutate(recommended_required = NA,
-         p_use = NA,
-         p_reporting = NA,
-         stat_sig_use = NA,
-         stat_sig_alpha = NA) %>%
-  write_csv(here('data','raw','templates','phase_two_pval.csv'))
+# p2_done <- read_csv(here('data','raw','phase_two_pval.csv'))
+# # export data for phase two
+# p2 <- d_all %>% filter(has_p_value == T | has_significance == T | has_effect_size == T | has_conf_int == T | has_sample_size == T) %>% 
+#   mutate(coder = "TEH") %>% 
+#   select(journal, esi_field, p_value, significance, conf_int, effect_size, sample_size) %>%
+#   mutate(conf_int_use = NA,
+#          effect_size_use = NA,
+#          effect_size_reporting = NA,
+#          sample_size_justification = NA,
+#          sample_size_power = NA) 
+# 
+# full_join(p2, p2_done, 
+#           by = c('journal','esi_field', 'p_value', 'significance')) %>%
+#   arrange(p_value, significance) %>%
+#   write_csv(here('data','raw','templates','phase_two.csv'))
+# 
+# tmp %>% filter(!is.na(significance))
+# d_all %>% filter(has_significance == T)
 ```

### 2021-09-02 analysis/preprocessing.R

@@ -2,9 +2,6 @@
 # load data from 'primary' folder
 # do any necessary munging
 # save processed data (in csv format and R format)
-library(tidyverse)
-library(assertr)
-library(janitor)
 
 # Load data
 d_coding <- read_csv(here::here("data", "primary", "data_coding.csv"))
@@ -420,15 +417,16 @@ lookup_external_guidance <-
   mutate(guidance_type = case_when(
     
     # Papers have doi (mostly)
-    str_detect(external_guidance, "10\\.\\d{4,9}/[-.;()/:\\w\\d]+$") ~ "paper",
-    str_detect(external_guidance, "statistics for biologists|ncbi.nlm.nih.gov/books/nbk153593|biostathandbook|biology_statistics_made_simple_using_excel") ~ "paper", #unsure
+    str_detect(external_guidance, "10\\.\\d{4,9}/[-.;()/:\\w\\d]+$") ~ "other",
+    str_detect(external_guidance, "statistics for biologists|ncbi.nlm.nih.gov/books/nbk153593|biostathandbook|biology_statistics_made_simple_using_excel") ~ "other", #unsure
     
     # Reporting guidelines are mostly a single word, possibly with a dash
     str_detect(external_guidance, "^[\\w-]+$") ~ "reporting guideline",
     str_detect(external_guidance, "consort|apa jars") ~ "reporting guideline",
     str_detect(external_guidance, "^nih|^nlm|^fda") ~ "reporting guideline",
-    str_detect(external_guidance, "biosharing|hugenet") ~ "reporting guideline", #unsure
-    str_detect(external_guidance, "apa|asa|cell star|frontiersin") ~ "publisher",
+    str_detect(external_guidance, "hugenet") ~ "reporting guideline", #unsure
+    str_detect(external_guidance, "cell star|frontiersin") ~ "publisher",
+    str_detect(external_guidance, "apa|asa") ~ "other",
     TRUE ~ NA_character_
   ))
 
@@ -436,7 +434,7 @@ lookup_external_guidance <-
 d_all <- inner_join(d_coding,d_journals, by = 'journal')
 
 readr::write_csv(d_all, file =  here::here("data", "processed", "d_all.csv"))
-save(d_coding, file =  here::here("data", "processed", "d_all.rds"))
+save(d_all, file =  here::here("data", "processed", "d_all.rds"))
 
[…]

### 2021-09-06 analysis/preliminaryAnalysis.Rmd
Update preliminaryAnalysis.Rmd

@@
 library(assertr) # for testing
+library(pointblank) # for testing
 library(lemon) # to have axes across all graph facets

### 2022-01-31 analysis/preprocessing.R
uploading data codebooks

@@ -427,6 +427,8 @@ lookup_external_guidance <-
   )) %>%
   filter(external_guidance != "equator") # as per protocol - we do not count general references to EQUATOR guidelines, so remove
 
+d_coding <- d_coding %>% select(-external_guidance_clean)
+
 # bind d_coding and d_journals
 d_all <- inner_join(d_coding,d_journals, by = 'journal')
 

### 2022-01-31 analysis/preprocessing.R
adding d_endorse to processed folder

@@ -6,6 +6,7 @@
 # Load data
 d_coding <- read_csv(here("data", "primary", "data_coding.csv"))
 d_journals <- read_csv(here("data", "primary", "data_journals.csv"))
+d_endorse <- read_csv(here('data','primary','data_endorsements.csv'))
 
 # Clean up column names
 d_journals <-
@@ -434,6 +435,9 @@ d_all <- inner_join(d_coding,d_journals, by = 'journal')
 
 write_csv(d_all, file =  here("data", "processed", "d_all.csv"))
 save(d_all, file =  here("data", "processed", "d_all.rds"))
+
+write_csv(d_endorse, file =  here("data", "processed", "d_endorse.csv"))
+save(d_endorse, file =  here("data", "processed", "d_endorse.rds"))
 
 write_csv(lookup_external_guidance, file =  here("data", "processed", "lookup_external_guidance.csv"))
 save(lookup_external_guidance, file =  here("data", "processed", "lookup_external_guidance.rds"))

### 2022-01-31 manuscript/manuscript.Rmd
manuscript

---
title             : 'Statistical guidance to authors at top-ranked scientific journals:A cross-disciplinary assessment'
shorttitle        : "Statistical guidance"

author:
  - name: Tom E. Hardwicke
    affiliation: '1'
    corresponding: yes
    email: tom.hardwicke@uva.nl
    address: 'Tom E. Hardwicke, Department of Psychology, University of Amsterdam, Nieuwe Achtergracht 129-B, 1018 WT Amsterdam, Netherlands.'
  - name: Maia Salholz-Hillel
    affiliation: '2'
  - name: 'Mario Malicki'
    affiliation: '3'
  - name: 'Denes Szűcs'
    affiliation: '4'
  - name: Theiss Bendixen
    affiliation: '5'
  - name: John P. A. Ioannidis
    affiliation: '3,6,7'

affiliation:
  - id: '1'
    institution: 'Department of Psychology, University of Amsterdam'
  - id: '2'
    institution: 'QUEST Center for Responsible Research, Berlin Institute of Health (BIH) at Charité Universitätsmedizin Berlin'
  - id: '3'
    institution: 'Meta-Research Innovation Center at Stanford (METRICS), Stanford University'
  - id: '4'
    institution: Department of Psychology, University of Cambridge
  - id: '5'
    institution: Department of the Study of Religion, Aarhus University
  - id: '6'
    institution: 'Meta-Research Innovation Center Berlin (METRIC-B), QUEST Center for Responsible Research, Berlin Institute of Health (BIH) at Charité Universitätsmedizin Berlin'
  - id: '7'
    institution: 'Departments of Medicine, Epidemiology and Population Health, Biomedical Data Science, and Statistics, Stanford University'

authornote: ''

abstract: |
  Scientific journals may counter the misuse and misinterpretation of statistical methods by providing statistical guidance to authors. Here, we sought to assess the nature and prevalence of statistical guidance offered to authors by 15 journals (top-ranked by Impact Factor) in each of 22 scientific disciplines (N = 330 journals). For each journal, two investigators manually extracted and classified statistical guidance provided on the journal website. 160 (48%) journals offered some statistical guidance and 93 (28%) had a dedicated statistical guidance section in their author instructions. Statistical guidance was most common in biomedical and health-related disciplines. Among twenty prespecified statistical topics, only two were mentioned in more than a quarter of the journals: confidence intervals (n = 90, 27% of journals) and p-values (n = 88, 27% of journals). Guidance on these topics was inconsistent across journals. For six ‘hotly debated’ topics, we found only three journals that explicitly opposed the use of “statistical significance”; more commonly, journals implicitly endorsed the use of p-values (n = 77), statistical significance (n = 35), and Bayesian statistics (n = 39) and explicitly endorsed reporting of effect sizes (n = 62), confidence intervals (n = 85), and sample size justifications (n = 67). Our investigation shows large gaps and inconsistent coverage in statistical guidance provided by top-ranked journals across scientific disciplines. These data may help to inform debates about the appropriate role of journal policy in addressing statistical issues.

keywords          : "statistics, statistical guidelines, author instructions, meta-research, journal policy"

bibliography      : ["references.bib"]

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
always_allow_html: true
appendix: "supplementary_information.Rmd"

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r load_packages}
library(papaja) # for article template
library(knitr) # for literate programming
library(tidyverse) # for data munging
library(janitor) # for data munging
#library(tidylog) # inline code feedback
library(here) # for finding files
library(assertr) # for testing
library(scales) # to colour palette
library(patchwork) # for multipanel figures
library(kableExtra) # for latex tables
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r load_functions}
source(here('analysis','functions.R')) # load custom functions
```

```{r perform_preprocessing}
# loads raw data, performs preprocessing, saves processed data files
source(here('analysis','preprocessing.R'))
```

```{r load_processed_data}
# loads the processed data files
load(here('data','processed','d_all.rds'))
load(here('data','processed','lookup_external_guidance.rds'))
load(here('data','processed','d_endorse.rds'))
```

```{r set_defaults}
# define a custom colour palette that links specific colours to specific scientific disciplines across graphs
scale_fill_domains <- function(...){
    ggplot2:::manual_scale(
        'fill', 
        values = setNames(c(hue_pal()(22),'#000000'), c(levels(factor(sum_tab_topics$esi_abbr[! sum_tab_topics$esi_abbr == 'OVERALL'])), 'OVERALL')), 
        ...
    )
}
```

# Introduction
The validity of scientific claims often depends on the appropriate selection, implementation, reporting, and interpretation of statistical analyses. However, serious concerns have been raised about the frequent misinterpretation and misuse of statistical methods in several scientific disciplines (Altman, 1982; Gigerenzer, 2018; Nieuwenhuis et al., 2011; Sedlmeier & Gigerenzer, 1989; Strasak et al., 2007; Wasserstein & Lazar, 2016). Academic journals may attempt to counter these problems by providing guidance on statistical issues in their instructions to authors (Altman et al., 1983; Bailar & Mosteller, 1988; Smith, 2005).

It is unclear whether journal-based statistical guidance, advocacy, or regulation is helpful or detrimental, especially in the context of controversial statistical topics (Mayo, 2021). There continues to be heated debate among statisticians and methodologists about fundamental statistical issues, as exemplified by the diverse commentaries in a 2019 special issue of The American Statistician (Wasserstein et al., 2019). P-values continue to be vigorously attacked (Wagenmakers, 2007) and staunchly defended (Mayo, 2018), and recent years have seen competing calls to either abandon (McShane et al., 2019), redefine (Benjamin et al., 2018), or justify (Lakens et al., 2018) statistical significance. Others have advocated for a so-called “New Statistics” centered on effect size estimation and confidence intervals (Cumming, 2014) or promoted wider adoption of Bayesian statistics (Goodman, 1999; Kruschke & Liddell, 2018).

In some cases, these debates have influenced journal policy and drastic changes in policy have sometimes backfired. For example, in 2005, in an attempt to avoid the “pitfalls of traditional statistical inference”, the journal Psychological Science endorsed a novel statistic prep (Killeen, 2005) which was subsequently adopted by the majority of publishing authors, and then abandoned under heavy criticism a few years later (Iverson et al., 2009). In 2015, the journal Basic and Applied Social Psychology introduced an outright ban on p-values and confidence intervals “because the state of the art remains uncertain” (Trafimow & Marks, 2015). An assessment of articles published in the first year after the ban argued that the absence of statistical safeguards led many authors to overstate their findings (Fricker et al., 2019). 
Although particularly controversial journal policies easily catch our attention, they may not be representative. Currently there is limited empirical data documenting the prevalence and content of statistical guidance for authors issued by scientific journals. Such data may help to inform debates about the appropriate role of journal policy in addressing statistical issues (Mayo, 2021) as well as identify gaps in journal-based statistical guidance within and across scientific disciplines. Some previous research has addressed this topic and suggested that there is probably large room for improvement. For example, Schriger et al. (2006) reported that 39% of 166 leading medical journals provided statistical guidance to authors. Topics addressed included p-values (20%), confidence intervals (14%), statistical power (10%), multiple testing (5%), modeling (4%), sensitivity analysis (4%), and Bayesian methods (1%). Recently, Malicki et al. (2019) used a text-mining approach to search for a limited set of terms related to Bayes factors, confidence intervals, effect sizes, and sample size calculations in instructions to authors provided by 835 journals across a range of scientific disciplines. The prevalence of any statistical guidance, defined by these terms, was 6% across all journals. Another empirical evaluation (Malicki et al., 2020) found that none of 57 preprint servers provided any statistical advice. These three studies provide some insight into the prevalence and content of statistical guidance offered to authors; however, Schriger et al. (2006) was limited to medical journals, Malicki et al. (2020) was limited to preprint servers, and Malicki et al. (2019) focused on only a few statistical topics.

In the present study, we assessed the nature and prevalence of statistical guidance offered to authors by 15 journals (top-ranked by Impact Factor) in each of 22 scientific disciplines (N = 330 journals). We manually inspected journal websites to identify any author-facing guidance related to statistics. We noted how often guidance pertained to twenty prespecified statistical issues and assessed whether journals endorsed or opposed statistical methods related to six topics that we considered in advance to be ‘hotly’ debated  (p-values, statistical significance, effect sizes, confidence intervals, sample size justification, and Bayesian statistics).

# Materials and methods
The study protocol (rationale, methods, and analysis plan) was pre-registered on November 23rd, 2019 (https://osf.io/cz9g3/). All departures from this protocol are explicitly acknowledged in Supplementary Information A. All data exclusions and measures conducted during this study are reported in this manuscript.

## Design
We examined whether each journal provided any statistical guidance, included a dedicated statistical guidance section in its instructions to authors, or referred to any external statistical guidance (e.g., reporting guideline). For journals that provided their own statistical guidance, we recorded whether the guidance addressed each of twenty prespecified (https://osf.io/cz9g3/) topics (see Figure 1B).

## Sample
The sample consisted of the top 15 journals ranked by 2017 Journal Impact Factor (JIF) in each of 22 high-level scientific disciplines (N = 330 journals). JIFs were obtained from Journal Citation Reports (https://jcr.clarivate.com) and scientific disciplines were defined by Essential Science Indicators (https://perma.cc/MD4V-A5X5). Journals that only publish reviews were not included.

## Procedure

1. Two investigators (TB & TEH) manually searched each journal’s website as of November 2019 and made copies of any statistical guidance intended for authors (for details see Supplementary Information B). Our operational definition of statistical guidance was: any advice or instruction related to the appropriate selection, implementation, reporting, or interpretation of statistical analyses.

2. A group of investigators (MM, DS, MSH) then manually examined and classified the extracted statistical guidance. All classifications were verified by a second investigator (TEH) and any differences were resolved through discussion. Investigators recorded whether journals provided a dedicated statistical guidance section in their author instructions and whether references were provided to external statistical guidance, such as reporting guidelines[^footnoteOne].

[^footnoteOne]: General instructions to follow external guidelines (e.g., “follow EQUATOR reporting guidelines”) that did specify specific guidelines, were not counted.

3. For guidance provided directly by journals, investigators recorded whether each of twenty prespecified topics were mentioned and extracted all relevant verbatim guidance pertaining to those topics[^footnoteTwo]. Investigators were each randomly assigned (using the R function ‘sample’) to five ranks and coded journals at that rank in each of the 22 scientific disciplines (i.e., 110 journals per coder).

[^footnoteTwo]: Note that for the topic “prespecification of analyses”, we did not count general instructions to register clinical trials unless they specifically referred to statistical analyses in some way, for example, a requirement to share the original statistical analysis plan.

4. For six topics that we considered in advance to be ‘hotly’ debated in the statistical literature (p-values, statistical significance, confidence intervals, effect sizes, sample size justification, and Bayesian statistics), one investigator (TEH) examined the extracted verbatim guidance and classified the level of endorsement offered by the journal according to four categories:

Explicit endorsement - the journal advises or instructs authors to use this method whenever possible/appropriate (e.g., “Authors should report p-values”).

Implicit endorsement - the journal provides advice on the method, implying endorsement, but does not explicitly advise that the method should be used (e.g., “Articles that use significance testing should report the alpha level for all tests”).

Implicit opposition - the journal advises that they would prefer the method is not used, but does not explicitly rule it out (e.g., “articles should not usually contain statistical significance testing”).

Explicit opposition - the journal advises that the method should not be used (e.g., “Authors should replace p-values with effect size estimates and 95% confidence intervals”).

5. For guidance pertaining to p-values, statistical significance, and confidence intervals, one investigator (TEH) recorded whether there were instructions to report exact p-values / alpha levels / confidence levels or use specific p-value thresholds / alpha / confidence levels (e.g., p < .05, a = .05, 95% confidence interval).

## Data analysis
We report descriptive statistics for all journals overall (N = 330 journals) and stratified by scientific discipline (n = 15 journals per discipline). Analyses were performed using R code. All data, materials, and analysis scripts related to this study are publicly available on the Open Science Framework (https://dx.doi.org/10.17605/OSF.IO/JRX7D). To facilitate reproducibility this manuscript was written by interleaving regular prose and analysis code using knitr (Xie, 2017) and papaja (Aust & Barth, 2020), and is available in a Code Ocean container (https://doi.org/10.24433/CO.6131540.v1) which re-creates the software environment in which the original analyses were performed.

# Results

## Journal Impact Factors

```{r journal-characteristics}
journalJIF <- d_journals %>% 
  rename(jif = journal_impact_factor) %>%
  group_by(esi_field) %>% 
  summarise(Median_jif = round(median(jif),2),
            IQR_jif = round(IQR(jif),2),
            Min_jif = round(min(jif),2),
            Max_jif = round(max(jif),2)) %>%
  ungroup() %>%
  bind_rows( # add overall summary
    d_journals %>% 
      rename(jif = journal_impact_factor) %>%
      summarise(Median_jif = round(median(jif),2),
                IQR_jif = round(IQR(jif),2),
                Min_jif = round(min(jif),2),
                Max_jif = round(max(jif),2)) %>%
      mutate(esi_field = 'OVERALL')
  ) %>%
  mutate(esi_field = str_to_title(esi_field)) %>%
  rename("Scientific discipline" = esi_field)

# for in text reporting
allFieldChar <- journalJIF %>% filter(`Scientific discipline` == 'Overall')
```

Across all 22 scientific fields (n = 330 journals), the median 2017 Journal Impact Factor was `r allFieldChar %>% pull(Median_jif)` (IQR = `r allFieldChar %>% pull(IQR_jif)`; range `r allFieldChar %>% pull(Min_jif)` - `r allFieldChar %>% pull(Max_jif)`). Journal Impact Factors for each scientific discipline are available in Supplementary Table F1.

```{r shared_publisher_journals}
# count number of journals that have shared publisher guidance
publisher_nature_n <- d_all %>% filter(publisher_guidance == "Nature") %>% nrow()
publisher_cell_n <- d_all %>% filter(publisher_guidance == "Cell") %>% nrow()
publisher_frontiers_n <- d_all %>% filter(publisher_guidance == "Frontiers") %>% nrow()
```

```{r summarise_data}
sum_tab <- d_all %>%

  # Recode NAs to FALSE
  mutate(across(
    everything()
    , ~replace_na(., FALSE) )) %>% 
  select(esi_field, starts_with("has_")) %>% 
  group_by(esi_field) %>%
  summarise(across(starts_with("has_"), list(n = sum, N = ~n()), .names = "{col}__{fn}")) %>%
  adorn_totals(name = 'OVERALL') %>%
  pivot_longer(cols = !esi_field, names_to = c('variable', '.value'), names_sep = '__', names_prefix = 'has_') %>%
  mutate(prop = n/N, percent = prop*100) %>%
  mutate(esi_abbr = case_when( # create abbreviated field labels to use in figures
    esi_field == "OVERALL" ~ "OVERALL",
    esi_field == "AGRICULTURAL SCIENCES" ~ "AGRI",
    esi_field == "BIOLOGY & BIOCHEMISTRY" ~ "BIO",       
    esi_field == "CHEMISTRY" ~ "CHEM" ,                   
    esi_field == "CLINICAL MEDICINE" ~ "MED",           
    esi_field == "COMPUTER SCIENCE" ~ "COM",             
    esi_field == "ECONOMICS & BUSINESS" ~ "ECON",         
    esi_field == "ENGINEERING" ~ "ENG",                  
    esi_field == "ENVIRONMENT_ECOLOGY" ~ "ECO",         
    esi_field == "GEOSCIENCES" ~ "GEO",                  
    esi_field == "IMMUNOLOGY" ~ "IMM",                   
    esi_field == "MATERIALS SCIENCE" ~ "MATSCI",            
    esi_field == "MATHEMATICS" ~ "MATH",                
    esi_field == "MICROBIOLOGY" ~ "MICBIO",                 
    esi_field == "MOLECULAR BIOLOGY & GENETICS" ~ "MOLBIO", 
    esi_field == "Multidisciplinary" ~ "MULTI",            
    esi_field == "NEUROSCIENCE & BEHAVIOR" ~ "NEURO",     
    esi_field == "PHARMACOLOGY & TOXICOLOGY" ~ "PHARM",    
    esi_field == "PHYSICS" ~ "PHYS",                     
    esi_field == "PLANT & ANIMAL SCIENCE" ~ "PLANT",       
    esi_field == "PSYCHIATRY_PSYCHOLOGY" ~ "PSY",       
    esi_field == "SOCIAL SCIENCES, GENERAL" ~ "SOC",     
    esi_field == "SPACE SCIENCE" ~ "SPACE"          
  )) %>%
  mutate(esi_field_with_abbr = paste0(esi_field,' (', esi_abbr,')')) %>%
  mutate(var_display = case_when( # create display versions of topic names
    variable == "guidance" ~ "Provides any statistical guidance",
    variable == "internal_guidance" ~ "Provides journal-specific statistical guidance",
    variable == "internal_guidance_section" ~ "Dedicated statistical guidance section",
    variable == "external_guidance" ~ "External statistical guidance",
    variable == "p_value" ~ "p values",
    variable == "significance" ~ "statistical significance",
    variable == "null_hypo" ~ "null hypotheses",
    variable == "sample_size" ~ "sample size",
    variable == "conf_int" ~ "confidence intervals",
    variable == "effect_size" ~ "effect sizes",
    variable == "multi_compare" ~ "multiple comparisons",
    variable == "subgroup" ~ "subgroup analyses",
    variable == "baseline_covar" ~ "baseline covariates",
    variable == "non_param" ~ "non-parametric tests",
    variable == "sensitivity" ~ "sensitivity analyses",
    variable == "model_assume" ~ "model assumptions",
    variable == "exclusion" ~ "data exclusions",
    variable == "outliers" ~ "outliers",
    variable == "missing" ~ "missing data",
    variable == "one_sided" ~ "one-tailed tests",
    variable == "bayes" ~ "Bayesian statistics",
    variable == "secondary" ~ "secondary outcomes",
    variable == "prespecify" ~ "prespecify analyses",
    variable == "cat_continuous" ~ "CCV",
    variable == "publisher_guidance" ~ "Shares publisher guidance"
  ))

# get fields with no statistical guidance at all
null_fields <- sum_tab %>% 
  filter(
    variable == "guidance",
    prop == 0
  ) %>%
  pull(esi_field)

# select only variables that address the general any/internal/external statistical guidance classifications 
sum_tab_general <- sum_tab %>%
  filter(variable %in% c('guidance', 'internal_guidance', 'internal_guidance_section', 'external_guidance')) %>% # select only data on general classifications
  mutate(variable = factor(variable, levels = c('guidance', 'internal_guidance', 'internal_guidance_section', 'external_guidance')),
         var_display = factor(var_display, levels = c("Provides any statistical guidance", "Provides journal-specific statistical guidance", "Dedicated statistical guidance section", "External statistical guidance"))) # set variable order

# select only variables on specific topics
sum_tab_topics <- sum_tab %>%
  #filter(esi_field %notin% null_fields) %>% # remove fields that have no statistical guidance at all
  filter(variable %notin% c('guidance', 'internal_guidance', 'internal_guidance_section', 'external_guidance', 'publisher_guidance')) %>% # select only data on topics
  rename(topic = variable, topic_display = var_display)

# get list of statistical topics in order of overall proportions
ordered_topics <- sum_tab_topics %>%
  filter(esi_field == "OVERALL") %>%
  arrange(desc(prop)) %>% # order from highest to lowest
  pull(topic_display)

# set order of topics
sum_tab_topics <- sum_tab_topics %>%
  mutate(topic_display = factor(topic_display, levels = ordered_topics))
```

```{r crosstab_int_ext}
crosstab_int_ext <- d_all %>% count(has_internal_guidance, has_external_guidance)
only_ext <- crosstab_int_ext %>% filter(has_internal_guidance == F, has_external_guidance == T) %>% pull(n)
```

## Frequency of statistical guidance in general

Out of `r sum_tab %>% filter(esi_field == "OVERALL", variable == "guidance") %>% pull(N)` journals we found that `r sum_tab %>% filter(esi_field == "OVERALL", variable == "guidance") %>% pull(n)` (`r sum_tab %>% filter(esi_field == "OVERALL", variable == "guidance") %>% pull(percent)`%) offered statistical guidance (Figure 1A; for equivalent tabular data, see Supplementary Information C). This includes `r only_ext` journals which only referred to statistical guidance in external sources (reporting guidelines or academic papers) and `r sum_tab %>% filter(esi_field == "OVERALL", variable == "internal_guidance") %>% pull(n)` journals that provided their own guidance (of which `r crosstab_int_ext %>% filter(has_external_guidance == T, has_internal_guidance == T) %>% pull(n)` also referred to external sources). `r sum_tab %>% filter(esi_field == "OVERALL", variable == "internal_guidance_section") %>% pull(n)` (`r sum_tab %>% filter(esi_field == "OVERALL", variable == "internal_guidance_section") %>% pull(percent)`%) of `r sum_tab %>% filter(esi_field == "OVERALL", variable == "internal_guidance_section") %>% pull(N)` journals had a dedicated statistical guidance section in their author instructions (Figure 1B). Journals offered statistical guidance more often in biomedical and health-related domains compared to other domains. Notably, all 15 Clinical Medicine journals offered some statistical guidance. In two domains (Computer Science and Mathematics), no journals offered any statistical guidance. Some journals shared the same publisher-level guidance, including `r publisher_nature_n` Nature journals, `r publisher_cell_n` Cell journals, and `r publisher_frontiers_n` Frontiers journals.

```{r}
# build plot for any statistical guidance
plot_any <- sum_tab_general %>%
  filter(variable == 'guidance') %>%
  arrange(percent) %>%
  mutate(esi_abbr = factor(esi_abbr, levels = c(esi_abbr[! esi_abbr == 'OVERALL'], 'OVERALL'), ordered = T)) %>%
  ggplot(aes(x=esi_abbr, y = percent, fill = esi_abbr)) +
  ylim(0,100) +
  geom_segment(aes(xend=esi_abbr, y=0, yend=percent)) +
  geom_point(size = 5, shape = 21, colour = 'black', data = . %>% filter(esi_abbr != 'OVERALL')) + 
  geom_point(size = 5, shape = 23, colour = 'black', data = . %>% filter(esi_abbr == 'OVERALL')) + 
  ggtitle('Any statistical guidance') +
  ylab('journals (%)') +
  scale_fill_domains() +
  theme_apa() +
  theme(
    panel.grid.major.x = element_line(),
    axis.title.y = element_blank(),
    legend.position = 'none') + 
  coord_flip()

# build plot for dedicated statistical guidance section
plot_dedicated <- sum_tab_general %>%
  filter(variable == 'internal_guidance_section') %>%
  arrange(percent) %>%
  mutate(esi_abbr = factor(esi_abbr, levels = c(esi_abbr[! esi_abbr == 'OVERALL'], 'OVERALL'), ordered = T)) %>%
  ggplot(aes(x=esi_abbr, y = percent, fill = esi_abbr)) +
  ylim(0,100) +
  geom_segment(aes(xend=esi_abbr, y=0, yend=percent)) +
  geom_point(size = 5, shape = 21, colour = 'black', data = . %>% filter(esi_abbr != 'OVERALL')) + 
  geom_point(size = 5, shape = 23, colour = 'black', data = . %>% filter(esi_abbr == 'OVERALL')) + 
  ggtitle('Dedicated statistical guidance section') +
  ylab('journals (%)') +
  scale_fill_domains() +
  theme_apa() +
  theme(
    panel.grid.major.x = element_line(),
    axis.title.y = element_blank(),
    legend.position = 'none') + 
  coord_flip()

# build combined plot for any statistical guidance and dedicated statistical guidance section

domainOrder <- sum_tab_general %>%
  filter(variable == 'guidance') %>%
  arrange(percent) %>% 
  pull(esi_abbr)

domainOrder <- c(domainOrder[! domainOrder == 'OVERALL'], 'OVERALL')

plot_any_dedicated_joint <- sum_tab_general %>%
  filter(variable %in% c('guidance', 'internal_guidance_section')) %>%
  mutate(esi_abbr = factor(esi_abbr, levels = domainOrder, ordered = T)) %>%
  ggplot(aes(x=esi_abbr, y = percent, fill = esi_abbr)) +
  ylim(0,100) +
  geom_segment(aes(xend=esi_abbr, y=0, yend=percent), data =. %>% filter(variable == 'guidance')) +
  geom_point(size = 5, shape = 21, colour = 'black', data = . %>% filter(variable == 'guidance', esi_abbr != 'OVERALL')) + 
  geom_point(size = 5, shape = 21, colour = 'black', data = . %>% filter(variable == 'guidance', esi_abbr == 'OVERALL')) + 
  geom_point(size = 2, shape = 23, colour = 'black', data = . %>% filter(variable == 'internal_guidance_section', esi_abbr != 'OVERALL')) + 
  geom_point(size = 2, shape = 23, colour = 'black', data = . %>% filter(variable == 'internal_guidance_section', esi_abbr == 'OVERALL')) + 
  ggtitle('General statistical guidance') +
  ylab('journals (%)') +
  scale_fill_domains() +
  theme_apa() +
  theme(
    panel.grid.major.x = element_line(),
    axis.title.y = element_blank(),
    legend.position = 'none') + 
  coord_flip()

```

## Frequency of statistical guidance on specific topics

```{r how_much_guidance}
n_topics <- d_all %>% filter(has_guidance == T, has_internal_guidance == T) %>%
  mutate(across(everything(), ~replace_na(., FALSE) )) %>% # Recode NAs to FALSE
  select(journal, starts_with("has_"), -c(has_guidance, has_internal_guidance, has_internal_guidance_section, has_external_guidance, has_publisher_guidance)) %>%
  pivot_longer(cols = !journal, names_to = 'topic', values_to = 'value') %>%
  group_by(journal) %>%
  summarise(topics_mentioned = sum(value))
```

Of the `r d_all %>% filter(has_guidance == T, has_internal_guidance == T) %>% nrow()` journals that provided their own guidance, `r n_topics %>% filter(topics_mentioned > 0) %>% nrow()` mentioned at least one of the twenty prespecified statistical topics (Supplementary Figures D2 and D3) and `r n_topics %>% filter(topics_mentioned == 0) %>% nrow()` did not mention any of these topics (but did provide guidance on other miscellaneous statistical issues). The maximum number of topics mentioned by an individual journal was `r max(n_topics$topics_mentioned)`. The median topics mentioned was `r median(n_topics$topics_mentioned)` (IQR = `r IQR(n_topics$topics_mentioned)`; Supplementary Figure D1).

The frequency of guidance pertaining to each of twenty prespecified statistical topics is shown in detail per discipline in Supplementary Information C. Guidance was most common for confidence intervals (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "conf_int") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "conf_int") %>% pull(percent) %>% round(0)`%, Figure 2A), p-values (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "p_value") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "p_value") %>% pull(percent) %>% round(0)`%, Figure 2B), sample size justification (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "sample_size") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "sample_size") %>% pull(percent) %>% round(0)`%, Figure 2C), effect sizes (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "effect_size") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "effect_size") %>% pull(percent) %>% round(0)`%, Figure 2D), prespecification of analyses (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "prespecify") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "prespecify") %>% pull(percent) %>% round(0)`%, Figure 2E), and data exclusions (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "exclusion") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "exclusion") %>% pull(percent) %>% round(0)`%, Figure 2F). Guidance was less common for checking model assumptions (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "model_assume") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "model_assume") %>% pull(percent) %>% round(0)`%, Supplementary Figure C1, Panel A), multiple comparisons (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "multi_compare") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "multi_compare") %>% pull(percent) %>% round(0)`%, Supplementary Figure C1, Panel B), statistical significance (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "significance") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "significance") %>% pull(percent) %>% round(0)`%, Supplementary Figure C1, Panel C), one-tailed hypothesis tests (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "one_sided") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "one_sided") %>% pull(percent) %>% round(0)`%, Supplementary Figure C1, Panel D), Bayesian statistics (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "bayes") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "bayes") %>% pull(percent) %>% round(0)`%, Supplementary Figure C1, Panel E), null hypotheses (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "null_hypo") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "null_hypo") %>% pull(percent) %>% round(0)`%, Supplementary Figure C1, Panel F), subgroup analyses (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "subgroup") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "subgroup") %>% pull(percent) %>% round(0)`%, Supplementary Figure C2, Panel A), secondary outcomes (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "secondary") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "secondary") %>% pull(percent) %>% round(0)`%, Supplementary Figure C2, Panel B), missing data (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "missing") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "missing") %>% pull(percent) %>% round(0)`%, Supplementary Figure C2, Panel C), sensitivity analyses (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "sensitivity") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "sensitivity") %>% pull(percent) %>% round(0)`%, Supplementary Figure C2, Panel D), non-parametric tests (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "non_param") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "non_param") %>% pull(percent) %>% round(0)`%, Supplementary Figure C3, Panel A), baseline covariates (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "baseline_covar") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "baseline_covar") %>% pull(percent) %>% round(0)`%, Supplementary Figure C3, Panel B), outliers (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "outliers") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "outliers") %>% pull(percent) %>% round(0)`%, Supplementary Figure C3, Panel C), and categorisation of continuous outcomes (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "cat_continuous") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "cat_continuous") %>% pull(percent) %>% round(0)`%, Supplementary Figure C3, Panel D). Guidance on the twenty topics tended to be more common amongst journals in biomedical and health-related domains, especially Clinical Medicine, compared to other domains. Examples of verbatim guidance are provided in Supplementary Table D1.

```{r}
plot_all_topics <- sum_tab_topics %>% 
  filter(esi_field == "OVERALL") %>% 
  arrange(percent) %>%
  mutate(topic_display = factor(topic_display, levels = topic_display, ordered = T)) %>%
  ggplot(aes(x=topic_display, y = percent)) +
    ylim(0,30) +
    geom_segment(aes(xend=topic_display, y=0, yend=percent)) +
    geom_point(size = 5, shape = 21, fill = 'black', colour = 'black') + 
    ggtitle('Specific statistical guidance') +
    ylab('journals (%)') +
    theme_apa() +
    theme(
      panel.grid.major.x = element_line(),
      axis.title.y = element_blank(),
      axis.text.y = element_text(size = 14),
      legend.position = 'none') + 
    coord_flip()
```

```{r generaltopicsfig, fig.path='figs/', dev=c('png', 'pdf'), fig.width = 8, fig.height = 6.5, fig.cap = "Percentage of journals offering some statistical guidance (Panel A circles), having a dedicated statistical guidance section (Panel A diamonds), or offering guidance on each of twenty prespecified statistical topics (Panel B).  Black circles represent data for all journals overall (N = 330 journals). Coloured circles and diamonds in Panel A represent individual scientific disciplines (n = 15 journals). Data is in descending order by percentage (after overall data in Panel A). CCV in Panel B refers to categorizing continuous outcomes. Discipline abbreviations: Agricultural Sciences (AGRI), Biology & Biochemistry (BIO), Chemistry (CHEM), Clinical Medicine (MED), Computer Science (COMSCI), Economics & Business (ECON), Engineering (ENGIN), Environment & Ecology (ECO), Geosciences (GEO), Immunology (IMMUN), Materials Science (MATSCI), Mathematics (MATH), Microbiology (MICBIO), Molecular Biology & Genetics (MOLBIO), Multidisciplinary (MULTI), Neuroscience & Behavior (NEURO), Pharmacology & Toxicology (PHARM), Physics (PHYS), Plant & Animal Science (PLANT), Psychiatry & Psychology (PSY), Social Sciences (SOCSCI), Space Science (SPACE)."}
plot_any_dedicated_joint + plot_all_topics + plot_annotation(tag_levels = 'A') &
  theme(plot.tag.position = c(0, 1),
        plot.tag = element_text(size = 14, hjust = 0, vjust = 0))
```

```{r}
# build plots for discipline-level data for each of the twenty topics
plotDomains <- function(thisTopic){
  thisData <- sum_tab_topics %>%
    filter(topic_display == thisTopic) %>%
    arrange(percent) %>%
    mutate(esi_abbr = factor(esi_abbr, levels = c(esi_abbr[! esi_abbr == 'OVERALL'], 'OVERALL'), ordered = T))

  ggplot(data = thisData, aes(x=esi_abbr, y = percent, fill = esi_abbr)) +
    ylim(0,100) +
    geom_segment(aes(xend=esi_abbr, y=0, yend=percent)) +
    geom_point(size = 5, shape = 21, colour = 'black', data = thisData %>% filter(esi_abbr != 'OVERALL')) + 
    geom_point(size = 5, shape = 21, colour = 'black', data = thisData %>% filter(esi_abbr == 'OVERALL')) + 
    ggtitle(thisTopic) +
    ylab('journals (%)') +
    scale_fill_domains() +
    theme_apa() +
    theme(
      panel.grid.major.x = element_line(),
      axis.title.y = element_blank(),
      legend.position = 'none') + 
    coord_flip()
}

topic_order <- sum_tab_topics %>% 
  filter(esi_field == "OVERALL") %>%
  arrange(desc(percent)) %>%
  pull(topic_display)

plot_list <- map(topic_order, plotDomains) # create list of plots for the 20 statistical topics
```

```{r stattopicsfig1, fig.path='figs/', fig.width = 8, fig.height = 10.5, fig.cap = "Percentage of journals offering guidance on six most commonly mentioned statistical topics (Panels A-F, ordered by frequency). Data is shown for all journals overall (N = 330 journals; black circles) and then in descending order by percentage for each scientific discipline (n = 15 journals; coloured circles). For discipline abbreviations see Figure 1."}
wrap_plots(plot_list[1:6], ncol = 3) + plot_annotation(tag_levels = 'A') &
  theme(plot.tag.position = c(0, 1),
        plot.tag = element_text(size = 14, hjust = 0, vjust = 0))
```

## External statistical guidance

```{r}
ext_guide_ordered <- lookup_external_guidance %>% arrange(n_journals) %>% pull(external_guidance)

reporting_guidelines <- lookup_external_guidance %>%
  filter(guidance_type == "reporting guideline") %>%
  mutate(external_guidance = factor(str_to_upper(external_guidance), levels = str_to_upper(ext_guide_ordered)),
         external_guidance = fct_recode(external_guidance, 
                                        `CONSORT PRO` = "CONSORT PRO EXTENSION",
                                        `CONSORT Cluster` = "CONSORT EXTENDED GUIDELINES",
                                        `HuGENet` = "HUMAN GENOME EPIDEMIOLOGY NETWORK (HUGENET) GUIDELINES",
                                        `FDA Diagnostic Tests` = "FDA GUIDELINES (HTTPS://PERMA.CC/W2EY-MSTA)",
                                        `CONSORT abstract` = "CONSORT ABSTRACT EXTENSION",
         NIH = "NIH PRINCIPLES AND GUIDELINES FOR REPORTING PRECLINICAL RESEARCH",
         NLM = "NLM RESEARCH REPORTING GUIDELINES AND INITIATIVES (HTTPS://WWW.NLM.NIH.GOV/SERVICES/RESEARCH_REPORT_GUIDE.HTML)"))
```

`r d_all %>% filter(has_external_guidance == T) %>% nrow()` journals referred authors to statistical guidance in an external source (Supplementary Table E1). Of these, `r crosstab_int_ext %>% filter(has_external_guidance == T, has_internal_guidance == F) %>% pull(n)` only referred to statistical guidance in external sources, whereas the other `r crosstab_int_ext %>% filter(has_external_guidance == T, has_internal_guidance == T) %>% pull(n)` also provided their own additional guidance. `r lookup_external_guidance %>% nrow()` different external sources were referred to, of which `r lookup_external_guidance %>% filter(guidance_type == 'reporting guideline') %>% nrow()` were reporting guidelines and `r lookup_external_guidance %>% filter(guidance_type != 'reporting guideline') %>% nrow()` were other types of guidance, such as academic papers, textbooks, or organizational guidelines. The most commonly referred to reporting guidelines were CONSORT (n = `r reporting_guidelines %>% filter(external_guidance == "CONSORT") %>% pull(n_journals)` journals), ARRIVE (n = `r reporting_guidelines %>% filter(external_guidance == "ARRIVE") %>% pull(n_journals)` journals), PRISMA (n = `r reporting_guidelines %>% filter(external_guidance == "PRISMA") %>% pull(n_journals)` journals), REMARK (n = `r reporting_guidelines %>% filter(external_guidance == "REMARK") %>% pull(n_journals)` journals), and STARD (n = `r reporting_guidelines %>% filter(external_guidance == "STARD") %>% pull(n_journals)` journals), with all other reporting guidelines being referred to by `r reporting_guidelines %>% filter(external_guidance == "ICMJE") %>% pull(n_journals)` or fewer journals. All other types of external guidance were referred to by `r lookup_external_guidance %>% filter(guidance_type != "reporting guideline") %>% pull(n_journals) %>% max()` or fewer journals.

## Statistical guidance on ‘hot’ topics

```{r s2table}
d_endorse_tab <- d_endorse %>%
  select(c(journal,esi_field), ends_with("_use")) %>%
  pivot_longer(cols = ends_with("_use"), names_to = 'topic', values_to = 'endorsement') %>%
  mutate(endorsement = fct_recode(
    endorsement,
    `Implicit opposition` = 'Implicit opposition (for low sample sizes)',
    `Explicit opposition` = 'Explicit opposition (without prespecified multiplicity corrections)')) %>%
  filter(endorsement != "No guidance")

d_endorse_tab %>%
  count(topic, endorsement, .drop = F) %>%
  mutate(topic = fct_recode(topic,
                            `confidence intervals` = "conf_int_use",
                            `effect sizes` = "effect_size_use",
                            `p-values` = "p_use",
                            `sample size justification` = "sample_size_use",
                            `statistical significance` = "stat_sig_use",
                            `Bayesian statistics` = "bayes_use")) %>%
  pivot_wider(id_cols = topic, names_from = endorsement, values_from = n) %>%
  select(topic, 'Explicit endorsement', 'Implicit endorsement', 'Implicit opposition', 'Explicit opposition') %>%
  kable(caption = "Number of journals offering various levels of endorsement for six ‘hot’ statistical topics. Explicit endorsement = the journal advises or instructs authors to use this method whenever possible/appropriate. Implicit endorsement = the journal provides advice on the method, implying endorsement, but does not explicitly advise that the method should be used. Implicit opposition = the journal advises that they would prefer the method is not used, but does not explicitly rule it out. Explicit opposition = the journal advises that the method should not be used.", booktabs = T, format = 'latex') %>%
  column_spec(2:5, width = "6em")
```

Table 1 shows how many journals offered different levels of endorsement for six statistical topics that we considered in advance to be hotly debated in the statistical literature (note that these only partly overlap with the most commonly mentioned statistical topics in journal guidance, as shown in Figure 2). `r d_endorse %>% filter(!is.na(conf_int)) %>% nrow()` journals provided guidance on confidence intervals, of which the vast majority (n = `r d_endorse %>% filter(conf_int_use == "Explicit endorsement") %>% nrow()`) offered explicit endorsement. Most of these journals (n = `r d_endorse %>% filter(!is.na(conf_int), conf_int_reporting == "No guidance") %>% nrow()`) did not provide any guidance on confidence levels, but `r d_endorse %>% filter(!is.na(conf_int)) %>% count(conf_int_reporting) %>% filter(conf_int_reporting == "Report 95% Cis") %>% pull(n)` advised authors to use a 95% confidence level and `r d_endorse %>% filter(!is.na(conf_int)) %>% count(conf_int_reporting) %>% filter(conf_int_reporting == "Report confidence level") %>% pull(n)` advised authors to report their chosen confidence level.

`r d_endorse %>% filter(!is.na(p_value)) %>% nrow()` journals provided guidance on p-values, of which the vast majority offered implicit endorsement (n = `r d_endorse %>% filter(p_use == "Implicit endorsement") %>% nrow()`). `r d_endorse %>% filter(p_use == "Explicit opposition (without prespecified multiplicity corrections)") %>% nrow()` journal explicitly opposed the use of p-values, but only in the absence of prespecified multiplicity corrections. `r d_endorse %>% filter(!is.na(p_value)) %>% count(p_reporting) %>% filter(p_reporting == "Exact") %>% pull(n)` journals advised reporting exact p-values, `r d_endorse %>% filter(!is.na(p_value)) %>% count(p_reporting) %>% filter(p_reporting == "Exact unless very small") %>% pull(n)` advised reporting of exact p-values unless they were very small (e.g., p < .001), `r d_endorse %>% filter(!is.na(p_value)) %>% count(p_reporting) %>% filter(p_reporting == "Thresholds [0.05, 0.01, 0.001]") %>% pull(n)` advised authors to use reporting thresholds (e.g., <.05, <.01, <.001), and `r d_endorse %>% filter(!is.na(p_value)) %>% count(p_reporting) %>% filter(p_reporting == "No guidance") %>% pull(n)` offered no guidance on on whether to report exact p-values or use thresholds.

`r d_endorse %>% filter(!is.na(significance)) %>% nrow()` journals provided guidance on statistical significance, of which the majority (n = `r d_endorse %>% filter(stat_sig_use == "Implicit endorsement") %>% nrow()`) offered implicit endorsement. `r d_endorse %>% filter(stat_sig_use == "Explicit opposition") %>% nrow()` journals explicitly opposed the use of statistical significance, instructing authors to report exact p-values and/or standard errors/confidence intervals instead. `r d_endorse %>% filter(!is.na(significance)) %>% count(stat_sig_reporting) %>% filter(stat_sig_reporting == "Report alpha") %>% pull(n)` journals advised authors to report the alpha level they used, `r d_endorse %>% filter(!is.na(significance)) %>% count(stat_sig_reporting) %>% filter(stat_sig_reporting == "0.05") %>% pull(n)` advised using an alpha level of .05, `r d_endorse %>% filter(!is.na(significance)) %>% count(stat_sig_reporting) %>% filter(stat_sig_reporting == "0.01") %>% pull(n)` advised using an alpha level of .01, and `r d_endorse %>% filter(!is.na(significance)) %>% count(stat_sig_reporting) %>% filter(stat_sig_reporting == "No guidance") %>% pull(n)` offered no guidance on how on alpha levels.

`r d_endorse %>% filter(!is.na(sample_size)) %>% nrow()` journals provided guidance on sample size justification, of which the vast majority (n = `r d_endorse %>% filter(sample_size_use == "Explicit endorsement") %>% nrow()`) offered explicit endorsement. `r d_endorse %>% filter(!is.na(effect_size)) %>% nrow()` journals offered guidance on effect sizes, of which the vast majority (n = `r d_endorse %>% filter(effect_size_use == "Explicit endorsement") %>% nrow()`) offered explicit endorsement. All `r d_endorse %>% filter(!is.na(bayes)) %>% nrow()` journals that provided guidance about Bayesian statistics provided implicit endorsement.

# Discussion
Our investigation found that 48% of 330 top-ranked journals across 22 scientific disciplines provided some statistical guidance to authors, with 28% of journals having a dedicated statistical guidance section in their instructions to authors. It was far more common for journals to provide statistical guidance in biomedical and health-related disciplines relative to non-health related disciplines like the social sciences and physical sciences. Statistical guidance addressed a range of topics, and was most common for hotly debated topics: p-values, confidence intervals, sample size justification, and effect sizes. However, very few journals appeared to take particularly controversial positions, such as banning p-values or the use of statistical significance.

Our finding that 48% of journals provide statistical guidance is substantially higher than observed in previous research. Malicki et al. (2019) found that statistical guidance was provided by only 6% of 835 journals across a similarly broad range of scientific disciplines; however, this estimate was based on a text-mining approach sensitive to only a narrow range of statistical topics (specifically, Bayes factors, confidence intervals, effect sizes, and power / sample size calculations). Additionally, our sample only included journals with high Impact Factors whereas Malicki et al. examined a representative (randomly selected) sample of journals across a range of citation impact. It is possible that journals with lower citation impact are less likely to provide statistical guidance (as is the case for authorship and conflict of interest guidance according to a recent systematic review, Malicki et al. 2021). Our methodology is more comparable to the study by Schriger et al. (2006) who reported that 39% of 166 major medical journals provided statistical guidance. By contrast, all of the medical journals we assessed provided statistical guidance; however, we only assessed the 15 journals with the highest Impact Factor. It is possible that there has been an increase in the provision of statistical guidance over time (as has been shown for guidance related to authorship, conflicts of interest, and data sharing, Malicki et al. 2021), but empirical comparisons are not possible because journal-level information is not provided in Schriger et al. Our study contributes to this literature by establishing a benchmark against which future studies of journal-based statistical guidance can be compared.

There was substantial variation across disciplines in the provision of statistical guidance. Our study was not designed to detect reasons for these differences. It is plausible that journals in biomedical and health-related disciplines are more likely to provide statistical guidance because (1) it seems especially important to avoid errors with highly proximal repercussions for human health; (2) biomedical and health-related research is often subject to stricter regulatory oversight; (3) there is a plethora of empirical data highlighting the severity and prevalence of statistical misuse (Altman, 1994; Strasak et al., 2007); (4) journals may feel they need to compensate for lower statistical competency among researchers due to inadequate educational programs (Aiken et al., 2008; Windish et al., 2007). Differences in the use of statistics between disciplines is also likely to be a major factor; for example, statistics are likely to be less useful in mathematical contexts relative to more applied sciences and may explain the absence of statistical guidance in the assessed Computer Science and Mathematics journals. Cultural or historic factors related to the role that journals traditionally play in the oversight of research may also be important, though we are not aware of relevant empirical evidence that documents this.

It is unclear whether journal-based statistical guidance is an effective means to address misuse of statistics. Indeed, one may be skeptical about whether instructions to authors are even commonly read. Some evidence does suggest that, at least in some cases, editorial advocacy for specific statistical practices has been effective. For example, several journals have successfully encouraged authors to abandon null-hypothesis significance testing in favour of an estimation approach centered on reporting effect sizes and confidence intervals (Fidler et al., 2004; Finch et al., 2004; Giofrè et al., 2017). However, these effects seem at least partly dependent on the active campaigning and oversight of specific editors, rather than written guidance to authors alone. Thus, a coordinated approach in which the journal’s expectations about statistical practice are made explicit in author instructions, then reinforced by editors and reviewers is likely important. However, even this may be insufficient. In one study, Fidler et al. (2004) observed that although authors tended to follow instructions to report confidence intervals, they often showed little evidence of actually using them to aid their interpretation of the research results. Soliciting dedicated statistical reviewers may be more effective at addressing statistical issues beyond simple reporting (Hardwicke et al., 2019; Hardwicke & Goodman, 2020). Ultimately, any journal-based intervention intended to improve statistical practice is likely to be limited by the knowledge, motivation, and understanding of authors, the improvement of which will require non-trivial changes to reward structures and educational practices deeply rooted in the scientific ecosystem (Goodman, 2019). 
As noted in the introduction, even well-intentioned statistical guidance offered by journals may be detrimental. This was clear in the case of the flawed statistic prep introduced at the journal Psychological Science (Iverson et al., 2009; Killeen, 2005) and arguably the case with the ongoing p-value ban at the journal Basic and Applied Social Psychology, depending on one’s statistical philosophy (Fricker et al., 2019; Trafimow & Marks, 2015). Mayo (2021) has recently argued that journal editors should enforce proper use of statistical methods but avoid ‘taking sides’ on particularly controversial matters, or advocating for specific statistical philosophies. However, the definition of ‘proper use’ may be in the eye of the beholder and some experimentation with different approaches could be informative, if ethically appropriate and properly monitored with meta-research (Hardwicke et al., 2020). Nevertheless, we found that very few journals in our sample appeared to be outwardly and decisively ‘taking sides’ in the sense of explicitly opposing the use of hotly debated statistical tools like p-values and statistical significance; however, the choice of wording often implicitly endorsed or opposed certain practices, which suggests journals may be “taking sides” albeit without clear language

It was common for journals to provide their own idiosyncratic statistical guidance. However, some publishers had developed guidance that was applied across multiple journals, often in the form of checklists (e.g., the Nature Life Sciences Reporting Checklist). Moreover, many journals referred to the same reporting guidelines, particularly the most established ones available on the EQUATOR website, like CONSORT, ARRIVE and PRISMA. It is unclear whether authors refer to and editors and reviewers enforce external reporting guidelines comparably with internal ones. Standardization could improve efficiency and consistency for authors, reviewers, and readers; but some degree of variability and experimentation may also be desirable. It is unclear where the balance lies.

Our study has some important limitations. (1) We focused only on statistical guidance provided on journal websites. We do not know how many authors actually read this information, or the extent to which journals enforce their use through editorial oversight or statistical review (Hardwicke et al., 2019; Hardwicke & Goodman, 2020). (2) We deliberately based our sample on 22 broadly defined disciplines to ensure our findings are representative across the sciences, but this does not necessarily imply that the findings will straightforwardly generalize to lower-level sub-fields within those disciplines. Moreover, we focused only on journals with the highest Impact Factor because they are likely to have the clearest potential to modify author behaviour. Lower-ranked journals may differ in many important respects; for example, they may have fewer resources to dedicate to developing statistical guidance. It is unclear if they might be more or less willing than top-ranked journals to introduce policy changes that could cause controversy, such as p-value bans. (3) As noted (Footnote 2), for the topic ‘preregistration of analyses’ we did not count general instructions to register clinical trials unless they specifically referred to the prespecification of statistical analyses. It is possible that some journals assumed a general instruction to register trials also constituted a specific instruction to prespecify statistical analyses; however we did not consider that to be sufficiently direct. Similarly, we did not consider a general instruction to authors that they should follow EQUATOR reporting guidelines (without specifically mentioning individual guidelines) to be sufficiently direct, and did not count it as statistical guidance (Footnote 1).

In conclusion, statistics are often a vital pillar in the edifice of scientific claims, but they are widely misused and misinterpreted across several scientific disciplines (Altman, 1982; Gigerenzer, 2018; Nieuwenhuis et al., 2011; Sedlmeier & Gigerenzer, 1989; Strasak et al., 2007; Wasserstein & Lazar, 2016). Journals could play a key role in alleviating these statistical ailments before they plague the academic literature. However, our investigation shows that even at top-ranked journals, only a minority provide any statistical guidance. Moreover, there are inconsistencies in the guidance provided across journals. Our investigation provides a layer of empirical evidence upon which to inform ongoing debates about the appropriate role of journals in the statistical practice of scientists.

# Funding statement
The authors received no specific funding for this work and no funder had any role in the research. Tom E. Hardwicke receives funding from the European Union’s Horizon 2020 research and innovation programme under the Marie Skłodowska-Curie grant agreement No. 841188. Maia Salholz-Hillel is employed as a researcher under research grants from the German Bundesministerium für Bildung und Forschung (BMBF). The Meta-Research Innovation Center at Stanford (METRICS) is supported by a grant from the Laura and John Arnold Foundation. The Meta-Research Innovation Center Berlin (METRIC-B) is supported by a grant from the Einstein Foundation and Stiftung Charite.

# Conflict of interest statement
Mario Malicki is a Co-Editor-in-Chief of the journal Research Integrity and Peer Review. All other authors declare no conflicts of interest.

# Author contributions
Conceptualization: T.E.H., J.P.A.I., M.M., and D.S. Data curation: T.E.H. and M.S.H. Formal analysis: T.E.H. and M.S.H. Investigation: T.E.H., M.S.H., M.M., and T.B. Methodology: T.E.H., J.P.A.I., M.M., and D.S. Project administration: T.E.H. Resources: T.E.H. Software: T.E.H. Supervision: T.E.H. and J.P.A.I. Validation: T.E.H. Visualization: T.E.H. Writing - original draft: T.E.H. Writing - review & editing: T.E.H., J.P.A.I., M.S.H., M.M., D.S., and T.B.

\newpage

# References

Aiken, L. S., West, S. G., & Millsap, R. E. (2008). Doctoral training in statistics, measurement, and methodology in psychology: Replication and extension of Aiken, West, Sechrest, and Reno’s (1990) survey of PhD programs in North America. The American Psychologist, 63(1), 32–50. https://doi.org/10.1037/0003-066X.63.1.32
Altman, D. G. (1982). Statistics in medical journals. Statistics in Medicine, 1(1), 59–71. https://doi.org/10.1002/sim.4780010109
Altman, D. G. (1994). The scandal of poor medical research. BMJ, 308(6924), 283–284. https://doi.org/10.1136/bmj.308.6924.283
Altman, D. G., Gore, S. M., Gardner, M. J., & Pocock, S. J. (1983). Statistical guidelines for contributors to medical journals. British Medical Journal (Clinical Research Ed.), 286(6376), 1489–1493.
Aust, F., & Barth, M. (2020). Papaja: Create APA manuscripts with rmarkdown. https://github.com/crsh/papaja
Bailar, J. C., & Mosteller, F. (1988). Guidelines for statistical reporting in articles for medical journals. Annals of Internal Medicine, 108(2), 266–273. https://doi.org/10.7326/0003-4819-108-2-266
Benjamin, D. J., Berger, J. O., Johannesson, M., Nosek, B. A., Wagenmakers, E.-J., Berk, R., Bollen, K. A., Brembs, B., Brown, L., Camerer, C., Cesarini, D., Chambers, C. D., Clyde, M., Cook, T. D., De Boeck, P., Dienes, Z., Dreber, A., Easwaran, K., Efferson, C., … Johnson, V. E. (2018). Redefine statistical significance. Nature Human Behaviour, 2(1), 6–10. https://doi.org/10.1038/s41562-017-0189-z
Cumming, G. (2014). The New Statistics: Why and how. Psychological Science, 25(1), 7–29. https://doi.org/10.1177/0956797613504966
Fidler, F., Thomason, N., Cumming, G., Finch, S., & Leeman, J. (2004). Editors can lead researchers to confidence intervals, but can’t make them think: Statistical reform lessons from medicine. Psychological Science, 15(2), 119–126. https://doi.org/10.1111/j.0963-7214.2004.01502008.x
Finch, S., Cumming, G., Williams, J., Palmer, L., Griffith, E., Alders, C., Anderson, J., & Goodman, O. (2004). Reform of statistical inference in psychology: The case of Memory & Cognition. Behavior Research Methods, Instruments, & Computers, 36(2), 312–324. https://doi.org/10.3758/BF03195577
Fricker, R. D., Burke, K., Han, X., & Woodall, W. H. (2019). Assessing the statistical analyses used in Basic and Applied Social Psychology after their p-value ban. The American Statistician, 73(sup1), 374–384. https://doi.org/10.1080/00031305.2018.1537892
Gigerenzer, G. (2018). Statistical rituals: The replication delusion and how we got there. Advances in Methods and Practices in Psychological Science, 1(2), 198–218. https://doi.org/10.1177/2515245918771329
Giofrè, D., Cumming, G., Fresc, L., Boedker, I., & Tressoldi, P. (2017). The influence of journal submission guidelines on authors’ reporting of statistics and use of open research practices. PLOS ONE, 12(4), e0175583. https://doi.org/10/f93jf8
Goodman, S. N. (1999). Toward evidence-based medical statistics. 2: The Bayes Factor. Annals of Internal Medicine, 130(12), 1005. https://doi.org/10.7326/0003-4819-130-12-199906150-00019
Goodman, S. N. (2019). Why is Getting Rid of P-Values So Hard? Musings on Science and Statistics. The American Statistician, 73(sup1), 26–30. https://doi.org/10.1080/00031305.2018.1558111
Hardwicke, T. E., Frank, M. C., Vazire, S., & Goodman, S. N. (2019). Should psychology journals adopt specialized statistical review? Advances in Methods and Practices in Psychological Science, 2(3), 240–249. https://doi.org/10.1177/2515245919858428
Hardwicke, T. E., & Goodman, S. N. (2020). How often do leading biomedical journals use statistical experts to evaluate statistical methods? The results of a survey. PLOS ONE, 15(10), e0239598. https://doi.org/10.1371/journal.pone.0239598
Hardwicke, T. E., Serghiou, S., Janiaud, P., Danchev, V., Crüwell, S., Goodman, S. N., & Ioannidis, J. P. A. (2020). Calibrating the scientific ecosystem through meta-research. Annual Review of Statistics and Its Application, 7(1), 11–37. https://doi.org/10.1146/annurev-statistics-031219-041104
Iverson, G. J., Lee, M. D., Zhang, S., & Wagenmakers, E.-J. (2009). prep: An agony in five Fits. Journal of Mathematical Psychology, 53(4), 195–202. https://doi.org/10.1016/j.jmp.2008.09.004
Killeen, P. R. (2005). An alternative to null-hypothesis significance tests. Psychological Science, 16(5), 345–353. https://doi.org/10.1111/j.0956-7976.2005.01538.x
Kruschke, J. K., & Liddell, T. M. (2018). Bayesian data analysis for newcomers. Psychonomic Bulletin & Review, 25(1), 155–177. https://doi.org/10.3758/s13423-017-1272-1
Lakens, D., Adolfi, F. G., Albers, C. J., Anvari, F., Apps, M. A. J., Argamon, S. E., Baguley, T., Becker, R. B., Benning, S. D., Bradford, D. E., Buchanan, E. M., Caldwell, A. R., Van Calster, B., Carlsson, R., Chen, S.-C., Chung, B., Colling, L. J., Collins, G. S., Crook, Z., … Zwaan, R. A. (2018). Justify your alpha. Nature Human Behaviour, 2(3), 168–171. https://doi.org/10.1038/s41562-018-0311-x
Malicki, M., Aalbersberg, Ij. J., Bouter, L., & ter Riet, G. (2019). Journals’ instructions to authors: A cross-sectional study across scientific disciplines. PLOS ONE, 14(9), e0222157. https://doi.org/10.1371/journal.pone.0222157
Malicki, M., Jeroncic, A., Aalbersberg, Ij. J., Bouter, L., & ter Riet, G. (2021). Systematic review and meta-analyses of studies analysing instructions to authors from 1987 to 2017. Nature Communications, 12(1), 5840. https://doi.org/10.1038/s41467-021-26027-y
Malicki, M., Jeroncic, A., ter Riet, G., Bouter, L. M., Ioannidis, J. P. A., Goodman, S. N., & Aalbersberg, Ij. J. (2020). Preprint servers’ policies, submission requirements, and transparency in reporting and research integrity recommendations. JAMA, 324(18), 1901–1903. https://doi.org/10.1001/jama.2020.17195
Mayo, D. G. (2018). Statistical Inference as Severe Testing: How to Get Beyond the Statistics Wars (1st ed.). Cambridge University Press. https://www.cambridge.org/core/product/identifier/9781107286184/type/book
Mayo, D. G. (2021). The statistics wars and intellectual conflicts of interest. Conservation Biology. https://doi.org/10.1111/cobi.13861
McShane, B. B., Gal, D., Gelman, A., Robert, C., & Tackett, J. L. (2019). Abandon Statistical Significance. The American Statistician, 73(sup1), 235–245. https://doi.org/10.1080/00031305.2018.1527253
Nieuwenhuis, S., Forstmann, B. U., & Wagenmakers, E.-J. (2011). Erroneous analyses of interactions in neuroscience: A problem of significance. Nature Neuroscience, 14(9), 1105–1107. https://doi.org/10.1038/nn.2886
Schriger, D. L., Arora, S., & Altman, D. G. (2006). The content of medical journal instructions for authors. Annals of Emergency Medicine, 48(6), 743-749.e4. https://doi.org/10.1016/j.annemergmed.2006.03.028
Sedlmeier, P., & Gigerenzer, G. (1989). Do studies of statistical power have an effect on the power of studies? Psychological Bulletin, 105(2), 309–316.
Smith, R. (2005). Statistical review for medical journals, journal’s perspective. In Encyclopedia of Biostatistics. John Wiley & Sons, Ltd. https://doi.org/10.1002/0470011815.b2a17141
Strasak, A. M., Zaman, Q., Marinell, G., Pfeiffer, K. P., & Ulmer, H. (2007). The use of statistics in medical research: A comparison of the New England Journal of Medicine and Nature Medicine. The American Statistician, 61(1), 47–55. https://doi.org/10.1198/000313007X170242
Trafimow, D., & Marks, M. (2015). Editorial. Basic and Applied Social Psychology, 37(1), 1–2. https://doi.org/10.1080/01973533.2015.1012991
Wagenmakers, E.-J. (2007). A practical solution to the pervasive problems ofp values. Psychonomic Bulletin & Review, 14(5), 779–804. https://doi.org/10.3758/BF03194105
Wasserstein, R. L., & Lazar, N. A. (2016). The ASA’s statement on p-values: Context, process, and purpose. The American Statistician, 70(2), 129–133. https://doi.org/10.1080/00031305.2016.1154108
Wasserstein, R. L., Schirm, A. L., & Lazar, N. A. (2019). Moving to a World Beyond “p < 0.05”. The American Statistician, 73(sup1), 1–19. https://doi.org/10.1080/00031305.2019.1583913
Windish, D. M., Huot, S. J., & Green, M. L. (2007). Medicine Residents’ Understanding of the Biostatistics and Results in the Medical Literature. JAMA, 298(9), 1010–1022. https://doi.org/10.1001/jama.298.9.1010
Xie, Y. (2017). Dynamic documents with R and knitr. CRC Press.

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup

### 2022-01-31 manuscript/supplementary_information.Rmd
manuscript

# Supplementary Information A. Protocol amendments.
The study protocol (rationale, methods, and analysis plan) was pre-registered on November 23rd, 2019 (https://osf.io/cz9g3/). Supplementary Table A1 outlines amendments between the pre-registered protocol and final report.

See manuscript for Supplementary Table A1.

# Supplementary Information B. Preserving statistical guidance.
We used three methods to preserve verbatim copies of statistical guidance we identified on journal webpages. During our initial examination of webpages in November, 2019, we firstly attempted to preserve webpages using the service perma.cc (https://perma.cc). Occasionally, when this tool failed, we made a PDF copy of the webpage which we stored on the OSF (https://osf.io/tzjc6/). Occasionally, we retroactively realized we needed to preserve a journal webpage, in which case we used the Internet Archive’s Wayback Machine service (https://web.archive.org/) to obtain a copy of the webpage as it existed in November, 2019. Links to the preserved documents are available in a data file on the OSF (https://osf.io/aj6wu/). All coding was performed using the copied information to avoid reproducibility problems if journals updated their websites during coding.

# Supplementary Information C. Study One additional figures.

```{r stattopicsfig2, fig.path='figs/', fig.width = 8, fig.height = 10.5, fig.cap = "Percentage of journals offering guidance on specific statistical topics, topics seven to twelve (Panels A-F, ordered by frequency). Data is shown for all journals overall (N = 330 journals; black circles) and then in descending order by percentage for each scientific discipline (n = 15 journals; coloured circles). For discipline abbreviations see Figure 1. For equivalent tabular data see https://osf.io/pkjhf/"}
wrap_plots(plot_list[7:12], ncol = 3) + plot_annotation(tag_levels = 'A') &
  theme(plot.tag.position = c(0, 1),
        plot.tag = element_text(size = 14, hjust = 0, vjust = 0))
```

```{r stattopicsfig3, fig.path='figs/', fig.width = 8, fig.height = 9, fig.cap = "Percentage of journals offering guidance on specific statistical topics, topics thirteen to sixteen (Panels A-D, ordered by frequency). Data is shown for all journals overall (N = 330 journals; black circles) and then in descending order by percentage for each scientific discipline (n = 15 journals; coloured circles). For discipline abbreviations see Figure 1. For equivalent tabular data see https://osf.io/pkjhf/"}
wrap_plots(plot_list[13:16], ncol = 2) + plot_annotation(tag_levels = 'A') &
  theme(plot.tag.position = c(0, 1),
        plot.tag = element_text(size = 14, hjust = 0, vjust = 0))
```

```{r stattopicsfig4, fig.path='figs/', fig.width = 8, fig.height = 9, fig.cap = "Percentage of journals offering guidance on specific statistical topics, topics seventeen to twenty (Panels A-D, ordered by frequency). Data is shown for all journals overall (N = 330 journals; black circles) and then in descending order by percentage for each scientific discipline (n = 15 journals; coloured circles). CCV = categorizing continuous variables. For discipline abbreviations see Figure 1. For equivalent tabular data see https://osf.io/pkjhf/"}
wrap_plots(plot_list[17:20], ncol = 2) + plot_annotation(tag_levels = 'A') &
  theme(plot.tag.position = c(0, 1),
        plot.tag = element_text(size = 14, hjust = 0, vjust = 0))
```

# Supplementary Information D: Additional data on twenty prespecified statistical topics.

```{r topicsHist, fig.path='figs/', fig.cap = 'Number of journals providing guidance on twenty prespecified statistical topics.'}
n_topics %>%
  ggplot(aes(x=topics_mentioned)) +
    geom_histogram() +
    ylim(0,40) +
    xlab('topics mentioned (n)') +
    ylab('journals (n)') +
    theme_apa() +
    theme(
      panel.grid.major.y = element_line()
    )
```

```{r}
my_palette <- c("#95D840FF", "#404788FF")

d_internal_guidance_noPublisher <-
  d_all %>%
  filter(has_internal_guidance == T, # select journals with internal guidance
         has_publisher_guidance == F | journal == "Scientific Data") # remove journals with shared publisher guidance except for Scientific Data (which shared guidance and its own guidance)

# were going to represent journals with publisher guidance in a single row per publisher. 
# select exemplar journals for each publisher
publisher_nature <- d_all %>% filter(journal == "Nature") %>% mutate(journal = paste0("Nature journals"," (n=", publisher_nature_n,')'))
publisher_cell <- d_all %>% filter(journal == "CELL") %>% mutate(journal = paste0("Cell journals"," (n=", publisher_cell_n,')'))
publisher_frontiers <- d_all %>% filter(journal == "Frontiers in Microbiology") %>% mutate(journal = paste0("Frontiers journals"," (n=", publisher_frontiers_n,')'))

# combine the journals with no publisher guidance with the publisher guidance rows
d_internal_guidance <- bind_rows(d_internal_guidance_noPublisher, publisher_nature, publisher_cell, publisher_frontiers)

# list of journals in order of most topics mentioned
journals_topics_mentioned_ordered <- d_internal_guidance %>%
  # Recode NAs to FALSE
  mutate(across(
    everything()
    , ~replace_na(., FALSE) )) %>% 
  mutate(journal = toupper(journal)) %>%
  select(journal, starts_with("has_"), -c(has_guidance, has_internal_guidance, has_internal_guidance_section, has_external_guidance, has_publisher_guidance)) %>%
  pivot_longer(cols = !journal, names_to = 'topic', values_to = 'value') %>%
  group_by(journal) %>%
  summarise(topics_mentioned = sum(value)) %>%
  arrange(topics_mentioned) %>%
  pull(journal)

plt_data <- d_internal_guidance %>%
  select(-has_guidance,-has_external_guidance,-has_internal_guidance,-has_internal_guidance_section,-has_publisher_guidance) %>%
  filter(if_any(starts_with("has"), ~ .x == TRUE)) %>% # remove the nine journals that had no guidance on any of the 20 topics
  mutate(journal = factor(toupper(journal)),
         journal = factor(journal, levels = journals_topics_mentioned_ordered)) %>% # order journals by n topics mentioned
  select(journal, starts_with("has_") & (!contains("guidance"))) %>%
  mutate(journal = fct_recode(journal, # shorten particularly long journal names well know by abbreviation
                              "PNAS" = "PROCEEDINGS OF THE NATIONAL ACADEMY OF SCIENCES OF THE UNITED STATES OF AMERICA",
                              "JAMA" = "JAMA-JOURNAL OF THE AMERICAN MEDICAL ASSOCIATION",
                              "THE BMJ" = "BMJ-BRITISH MEDICAL JOURNAL",
                              "JOURNAL OF ALLERGY & CLINICAL IMMUNOLOGY" = "JOURNAL OF ALLERGY AND CLINICAL IMMUNOLOGY",
                              "JOURNAL OF ALLERGY & CLINICAL IMMUNOLOGY-IN PRACTICE" = "JOURNAL OF ALLERGY AND CLINICAL IMMUNOLOGY-IN PRACTICE",
                              "INTERNATIONAL JOURNAL OF BEHAVIORAL NUTRITION & PHYSICAL ACTIVITY" = "INTERNATIONAL JOURNAL OF BEHAVIORAL NUTRITION AND PHYSICAL ACTIVITY")) %>% 
  pivot_longer(
    -journal,
    names_to = "topic",
    names_prefix = 'has_',
    values_to = "outcome",
    values_drop_na = FALSE
  ) %>%
  mutate(outcome = case_when(
    outcome ~ "Mentioned",
    !outcome ~ "Not mentioned",
    TRUE ~ "Not applicable (No internal guidance)"
  )) %>%
  mutate(topic_display = case_when( # create display versions of topic names
    topic == "guidance" ~ "Provides any statistical guidance",
    topic == "internal_guidance" ~ "Provides journal-specific statistical guidance",
    topic == "internal_guidance_section" ~ "Dedicated statistical guidance section",
    topic == "external_guidance" ~ "Refers to external statistical guidance",
    topic == "p_value" ~ "p values",
    topic == "significance" ~ "statistical significance",
    topic == "null_hypo" ~ "null hypotheses",
    topic == "sample_size" ~ "sample size",
    topic == "conf_int" ~ "confidence intervals",
    topic == "effect_size" ~ "effect sizes",
    topic == "multi_compare" ~ "multiple comparisons",
    topic == "subgroup" ~ "subgroup analyses",
    topic == "baseline_covar" ~ "baseline covariates",
    topic == "non_param" ~ "non-parametric tests",
    topic == "sensitivity" ~ "sensitivity analyses",
    topic == "model_assume" ~ "model assumptions",
    topic == "exclusion" ~ "data exclusions",
    topic == "outliers" ~ "outliers",
    topic == "missing" ~ "missing data",
    topic == "one_sided" ~ "one-tailed tests",
    topic == "bayes" ~ "Bayesian statistics",
    topic == "secondary" ~ "secondary outcomes",
    topic == "prespecify" ~ "prespecify analyses",
    topic == "cat_continuous" ~ "CCV",
    topic == "publisher_guidance" ~ "Shares publisher guidance"
  )) %>%
  mutate(topic_display = factor(topic_display, levels = ordered_topics)) # order topics by n topics mentioned

plt_one <- plt_data %>% filter(journal %in% levels(plt_data$journal)[44:86]) %>% # plot first half
  ggplot(aes(x = topic_display, y = journal)) +
    geom_tile(aes(fill = outcome), colour = 'black', size = .25, alpha = .6) +
    scale_fill_manual(
      values = my_palette,
      # labels = c('Not mentioned', 'Unclear', 'Mentioned')
    ) +
    scale_x_discrete(position = 'top') +
    scale_y_discrete(labels = wrap_format(40)) + # wrap text on y axis
    theme(plot.margin = margin(0, 50, 0, 0),
          axis.ticks = element_blank(),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.line = element_blank(),
          panel.grid.major=element_line(colour=NA),
          panel.grid.minor=element_line(colour=NA),
          legend.position = 'top',
          axis.text.x  = element_text(angle = 45, hjust = 0),
          axis.text.y  = element_text(size = 8),
          panel.background = element_blank())

plt_two <- plt_data %>% filter(journal %in% levels(plt_data$journal)[1:43]) %>% # plot second half
  ggplot(aes(x = topic_display, y = journal)) +
    geom_tile(aes(fill = outcome), colour = 'black', size = .25, alpha = .6) +
    scale_fill_manual(
      values = my_palette,
      # labels = c('Not mentioned', 'Unclear', 'Mentioned')
    ) +
    scale_x_discrete(position = 'top') +
    scale_y_discrete(labels = wrap_format(40)) + # wrap text on y axis
    theme(plot.margin = margin(0, 50, 0, 0),
          axis.ticks = element_blank(),
          axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.line = element_blank(),
          panel.grid.major=element_line(colour=NA),
          panel.grid.minor=element_line(colour=NA),
          legend.position = 'top',
          axis.text.x  = element_text(angle = 45, hjust = 0),
          axis.text.y  = element_text(size = 8),
          panel.background = element_blank())
```

```{r gridOne, fig.path='figs/', fig.cap = "Whether each journal provided guidance on each of twenty statistical topics. Journals that shared publisher guidance are represented by one row (per publisher). Journals that did not offer any statistical guidance at all, or journals that referred only to guidance in external sources, are not shown. CCV = categorizing continuous variables.", fig.height = 11, fig.width = 8}
plt_one
```

```{r gridTwo, fig.path='figs/', fig.cap = "Continuation of Supplementary Figure D2.", fig.height = 11, fig.width = 8}
plt_two
```

# Supplementary Information E. External sources of statistical guidance.

```{r}
reporting_guidelines %>%
  select(-guidance_type) %>%
  kable(caption = "Number of journals referring to reporting guidelines.", booktabs = T, longtable = T)
```

```{r}
lookup_external_guidance %>% 
  filter(guidance_type != 'reporting guideline') %>%
  select(-guidance_type) %>%
  kable(caption = "Number of journals referring to other external statistical guidance.", booktabs = T, longtable = T) %>%
  column_spec(1, width = "30em")
```

# Supplementary Information F. Journal Impact Factors stratified by scientific discipline.

```{r}
journalJIF  %>%
  kable(caption = "Journal Impact Factors stratified by each of 22 scientific disciplines")
```

### 2022-02-01 manuscript/manuscript.Rmd
Update manuscript.Rmd

@@
 plot_list <- map(topic_order, plotDomains) # create list of plots for the 20 statistical topics
+
+# to save individual plots for twitter
+# for(i in 1:20){
+#    ggsave(plot = plot_list[[i]], file = here('manuscript','figs','twitter',paste("topic",i,".png",sep="")), width = 6, height = 8, units = "in", dpi = 300)
+#  }
 ```

### 2022-02-01 manuscript/manuscript.Rmd
Update manuscript.Rmd

@@
         plot.tag = element_text(size = 14, hjust = 0, vjust = 0))
+
+# to save individual plots for twitter
+# ggsave(plot = plot_any_dedicated_joint, file = here('manuscript','figs','twitter','fig1A.png'), width = 6, height = 6.5, units = "in", dpi = 300)
+# ggsave(plot = plot_all_topics, file = here('manuscript','figs','twitter','fig1B.png'), width = 6, height = 6.5, units = "in", dpi = 300)
 ```

### 2022-04-28 analysis/preprocessing.R
added domain classifications

@@ -224,8 +224,6 @@ d_coding <- bind_rows(d_coding,allPublishers_shared)
 # Check that there is one row per journal in d_coding
 d_coding <- d_coding %>% verify(nrow(.) == nrow(d_journals))
   
-
-
 # EXTERNAL GUIDANCE -------------------------------------------------------
 
 # Tidy external guidance
@@ -432,6 +430,23 @@ d_coding <- d_coding %>% select(-external_guidance_clean)
 
 # bind d_coding and d_journals
 d_all <- inner_join(d_coding,d_journals, by = 'journal')
+
+# add higher-level domain classifications
+d_all <- d_all %>%
+  mutate(domain = case_when(
+    esi_field %in% c('AGRICULTURAL SCIENCES','BIOLOGY & BIOCHEMISTRY','ENVIRONMENT_ECOLOGY','MICROBIOLOGY','MOLECULAR BIOLOGY & GENETICS','NEUROSCIENCE & BEHAVIOR','PLANT & ANIMAL SCIENCE') 
+    ~'Life sciences',
+    esi_field %in% c('CLINICAL MEDICINE','IMMUNOLOGY','PHARMACOLOGY & TOXICOLOGY') 
+    ~'Health sciences',
+    esi_field %in% c('CHEMISTRY','ENGINEERING','GEOSCIENCES','MATERIALS SCIENCE','SPACE SCIENCE') 
+    ~'Physical sciences',
+    esi_field %in% c('ECONOMICS & BUSINESS','PSYCHIATRY_PSYCHOLOGY','SOCIAL SCIENCES, GENERAL') 
+    ~'Social sciences',
+    esi_field %in% c('COMPUTER SCIENCE','MATHEMATICS','PHYSICS') 
+    ~'Formal sciences',
+    esi_field %in% c('Multidisciplinary')
+    ~'Multidisciplinary'
+  ))
 
 write_csv(d_all, file =  here("data", "processed", "d_all.csv"))
 save(d_all, file =  here("data", "processed", "d_all.rds"))

### 2022-04-28 manuscript/manuscript.Rmd
fixed bug arising from using everything with across

@@
   mutate(across(
-    everything()
-    , ~replace_na(., FALSE) )) %>% 
+    starts_with('has_'), ~replace_na(., FALSE) )) %>% 
   select(esi_field, starts_with("has_")) %>% 
@@
 n_topics <- d_all %>% filter(has_guidance == T, has_internal_guidance == T) %>%
-  mutate(across(everything(), ~replace_na(., FALSE) )) %>% # Recode NAs to FALSE
+  mutate(across(starts_with('has_'), ~replace_na(., FALSE) )) %>% # Recode NAs to FALSE
   select(journal, starts_with("has_"), -c(has_guidance, has_internal_guidance, has_internal_guidance_section, has_external_guidance, has_publisher_guidance)) %>%
@@
 \endgroup
+
+\newpage
+
+# (APPENDIX) Appendix {-}
+
+```{r child = "supplementary_information.Rmd"}
+```

### 2022-04-28 manuscript/supplementary_information.Rmd
fixed bug arising from using everything with across

@@
   mutate(across(
-    everything()
-    , ~replace_na(., FALSE) )) %>% 
+    starts_with('has_'), ~replace_na(., FALSE) )) %>% 
   mutate(journal = toupper(journal)) %>%

### 2022-04-28 analysis/preprocessing.R
putting physics in correct domain

@@ -438,11 +438,11 @@ d_all <- d_all %>%
     ~'Life sciences',
     esi_field %in% c('CLINICAL MEDICINE','IMMUNOLOGY','PHARMACOLOGY & TOXICOLOGY') 
     ~'Health sciences',
-    esi_field %in% c('CHEMISTRY','ENGINEERING','GEOSCIENCES','MATERIALS SCIENCE','SPACE SCIENCE') 
+    esi_field %in% c('CHEMISTRY','ENGINEERING','GEOSCIENCES','MATERIALS SCIENCE','SPACE SCIENCE','PHYSICS') 
     ~'Physical sciences',
     esi_field %in% c('ECONOMICS & BUSINESS','PSYCHIATRY_PSYCHOLOGY','SOCIAL SCIENCES, GENERAL') 
     ~'Social sciences',
-    esi_field %in% c('COMPUTER SCIENCE','MATHEMATICS','PHYSICS') 
+    esi_field %in% c('COMPUTER SCIENCE','MATHEMATICS') 
     ~'Formal sciences',
     esi_field %in% c('Multidisciplinary')
     ~'Multidisciplinary'

### 2022-05-15 manuscript/manuscript.Rmd
AmStat round 2 revisions

@@
 library(scales) # to colour palette
+library(ggforce) # helpful plotting functions
 library(patchwork) # for multipanel figures
@@
 # define a custom colour palette that links specific colours to specific scientific disciplines across graphs
+
+scale_fill_disciplines <- function(...){
+    ggplot2:::manual_scale(
+        'fill',
+        values = setNames(c(hue_pal()(22),'#000000'), c(levels(factor(sum_tab_topics$esi_abbr[! sum_tab_topics$esi_abbr == 'OVERALL'])), 'OVERALL')),
+        ...
+    )
+}
+
 scale_fill_domains <- function(...){
@@
         'fill', 
-        values = setNames(c(hue_pal()(22),'#000000'), c(levels(factor(sum_tab_topics$esi_abbr[! sum_tab_topics$esi_abbr == 'OVERALL'])), 'OVERALL')), 
+        values = setNames(c(hue_pal()(5),'#000000'), c(levels(factor(sum_tab_topics$domain[! sum_tab_topics$domain == 'OVERALL'])), 'OVERALL')), 
         ...
@@
 Others have advocated for a so-called “New Statistics” centered on effect size estimation and confidence intervals (Cumming, 2014) or promoted wider adoption of Bayesian statistics (Goodman, 1999; Kruschke & Liddell, 2018).
-
 In some cases, these debates have influenced journal policy and drastic changes in policy have sometimes backfired.
@@
 An assessment of articles published in the first year after the ban argued that the absence of statistical safeguards led many authors to overstate their findings (Fricker et al., 2019).
+
 Although particularly controversial journal policies easily catch our attention, they may not be representative.
@@
 
-Across all 22 scientific fields (n = 330 journals), the median 2017 Journal Impact Factor was `r allFieldChar %>% pull(Median_jif)` (IQR = `r allFieldChar %>% pull(IQR_jif)`; range `r allFieldChar %>% pull(Min_jif)` - `r allFieldChar %>% pull(Max_jif)`).
+Across all 22 scientific disciplines (n = 330 journals), the median 2017 Journal Impact Factor was `r allFieldChar %>% pull(Median_jif)` (IQR = `r allFieldChar %>% pull(IQR_jif)`; range `r allFieldChar %>% pull(Min_jif)` - `r allFieldChar %>% pull(Max_jif)`).
 Journal Impact Factors for each scientific discipline are available in Supplementary Table F1.
@@
     starts_with('has_'), ~replace_na(., FALSE) )) %>% 
-  select(esi_field, starts_with("has_")) %>% 
-  group_by(esi_field) %>%
+  select(esi_field, domain, starts_with("has_")) %>% 
+  group_by(esi_field, domain) %>%
   summarise(across(starts_with("has_"), list(n = sum, N = ~n()), .names = "{col}__{fn}")) %>%
   adorn_totals(name = 'OVERALL') %>%
-  pivot_longer(cols = !esi_field, names_to = c('variable', '.value'), names_sep = '__', names_prefix = 'has_') %>%
+  pivot_longer(cols = !c(esi_field,domain), names_to = c('variable', '.value'), names_sep = '__', names_prefix = 'has_') %>%
   mutate(prop = n/N, percent = prop*100) %>%
+  mutate(domain = ifelse(esi_field == "OVERALL", "OVERALL", domain)) %>%
   mutate(esi_abbr = case_when( # create abbreviated field labels to use in figures
@@
     esi_field == "ENGINEERING" ~ "ENG",                  
-    esi_field == "ENVIRONMENT_ECOLOGY" ~ "ECO",         
+    esi_field == "ENVIRONMENT & ECOLOGY" ~ "ECO",         
     esi_field == "GEOSCIENCES" ~ "GEO",                  
@@
     esi_field == "MOLECULAR BIOLOGY & GENETICS" ~ "MOLBIO", 
-    esi_field == "Multidisciplinary" ~ "MULTI",            
+    esi_field == "MULTIDISCIPLINARY" ~ "MULTI",            
     esi_field == "NEUROSCIENCE & BEHAVIOR" ~ "NEURO",     
@@
     esi_field == "PLANT & ANIMAL SCIENCE" ~ "PLANT",       
-    esi_field == "PSYCHIATRY_PSYCHOLOGY" ~ "PSY",       
-    esi_field == "SOCIAL SCIENCES, GENERAL" ~ "SOC",     
+    esi_field == "PSYCHIATRY & PSYCHOLOGY" ~ "PSY",       
+    esi_field == "SOCIAL SCIENCES (GENERAL)" ~ "SOC",     
     esi_field == "SPACE SCIENCE" ~ "SPACE"          
@@
 
+```{r}
+# domain-level summary
+
+sum_tab_domain <- sum_tab %>% 
+  group_by(domain, variable) %>%
+  summarise(n = sum(n), N = sum(N)) %>%
+  mutate(prop = n/N, percent = n/N*100)
+
+any_guidance_healthlife <- sum_tab_domain %>% filter(domain == 'HEALTH & LIFE SCIENCES', variable == 'guidance')
+any_guidance_multi <- sum_tab_domain %>% filter(domain == 'MULTIDISCIPLINARY', variable == 'guidance')
+any_guidance_social <- sum_tab_domain %>% filter(domain == 'SOCIAL SCIENCES', variable == 'guidance')
+any_guidance_physical <- sum_tab_domain %>% filter(domain == 'PHYSICAL SCIENCES', variable == 'guidance')
+any_guidance_formal <- sum_tab_domain %>% filter(domain == 'FORMAL SCIENCES', variable == 'guidance')
+```
+
+The frequency of journal author instructions providing statistical guidance varied considerably across domains (Health and Life Sciences: `r any_guidance_healthlife %>% pull(n)` of `r any_guidance_healthlife %>% pull(N)` journals, `r any_guidance_healthlife %>% pull(percent) %>% round()`%; Multidisciplinary: `r any_guidance_multi %>% pull(n)` of `r any_guidance_multi %>% pull(N)` journals, `r any_guidance_multi %>% pull(percent) %>% round()`%; Social Sciences: `r any_guidance_social %>% pull(n)` of `r any_guidance_social %>% pull(N)` journals, `r any_guidance_social %>% pull(percent) %>% round()`%; Physical Sciences: `r any_guidance_physical %>% pull(n)` of `r any_guidance_physical %>% pull(N)` journals, `r any_guidance_physical %>% pull(percent) %>% round()`; Formal Sciences: `r any_guidance_formal %>% pull(n)` of `r any_guidance_formal %>% pull(N)` journals, `r any_guidance_formal %>% pull(percent) %>% round()`.
+
 Out of `r sum_tab %>% filter(esi_field == "OVERALL", variable == "guidance") %>% pull(N)` journals we found that `r sum_tab %>% filter(esi_field == "OVERALL", variable == "guidance") %>% pull(n)` (`r sum_tab %>% filter(esi_field == "OVERALL", variable == "guidance") %>% pull(percent)`%) offered statistical guidance (Figure 1A; for equivalent tabular data, see Supplementary Information C).
@@
 `r sum_tab %>% filter(esi_field == "OVERALL", variable == "internal_guidance_section") %>% pull(n)` (`r sum_tab %>% filter(esi_field == "OVERALL", variable == "internal_guidance_section") %>% pull(percent)`%) of `r sum_tab %>% filter(esi_field == "OVERALL", variable == "internal_guidance_section") %>% pull(N)` journals had a dedicated statistical guidance section in their author instructions (Figure 1B).
-Journals offered statistical guidance more often in biomedical and health-related domains compared to other domains.
+Journals offered statistical guidance more often in Health and Life Sciences compared to other domains.
 Notably, all 15 Clinical Medicine journals offered some statistical guidance.
@@
   ylab('journals (%)') +
-  scale_fill_domains() +
+  scale_fill_disciplines() +
   theme_apa() +
@@
 
-domainOrder <- sum_tab_general %>%
+disciplineOrder <- sum_tab_general %>%
   filter(variable == 'guidance') %>%
@@
 
-domainOrder <- c(domainOrder[! domainOrder == 'OVERALL'], 'OVERALL')
+disciplineOrder <- c(disciplineOrder[! disciplineOrder == 'OVERALL'], 'OVERALL')
 
@@
   filter(variable %in% c('guidance', 'internal_guidance_section')) %>%
-  mutate(esi_abbr = factor(esi_abbr, levels = domainOrder, ordered = T)) %>%
+  mutate(esi_abbr = factor(esi_abbr, levels = disciplineOrder, ordered = T)) %>%
   ggplot(aes(x=esi_abbr, y = percent, fill = esi_abbr)) +
@@
   coord_flip()
+
+```
+
+```{r}
+### REVISED PLOTS ###
+
+# first make a data frame with overall data for each domain
+
+domain_overalls <- sum_tab_general %>% 
+  filter(esi_field != "OVERALL") %>% # remove previous overall summary
+  group_by(domain, variable, var_display) %>%
+  summarise(n = sum(n), N = sum(N)) %>%
+  mutate(prop = n/N, percent = n/N*100, esi_abbr = paste(domain,'(ALL)'), esi_field = paste(domain,'(ALL)'), esi_field_with_abbr = paste(domain,'(ALL)'), priority = 2)
+
+forPlot <- sum_tab_general %>% 
+  filter(esi_field != "OVERALL") %>% # remove previous overall summary 
+  mutate(priority = 1) %>%
+  rbind(domain_overalls)
+
+# set ordering of disciplines and domains
+domainOrder <- domain_overalls %>%
+  filter(variable == 'guidance') %>%
+  arrange(desc(percent)) %>% 
+  pull(domain)
+
+axesOrder <- forPlot %>%
+  filter(variable == 'guidance', esi_field != "MULTIDISCIPLINARY") %>%
+  mutate(domain = factor(domain, levels = domainOrder, ordered = T)) %>%
+  arrange(domain, priority, percent) %>% 
+  pull(esi_field)
+
+ready <- forPlot %>% 
+  filter(esi_field != "MULTIDISCIPLINARY",
+    variable %in% c('guidance', 'internal_guidance_section')) %>%
+  mutate(esi_field = factor(esi_field, levels = axesOrder, ordered = T),
+         domain = factor(domain, levels = domainOrder, ordered = T))
+
+plot_any_dedicated_joint <- ready %>%
+  ggplot(aes(y=esi_field, x = percent, fill = domain)) +
+  xlim(0,100) +
+  ggforce::facet_col(facets = vars(domain), 
+                     scales = "free_y", 
+                     space = "free") +
+  geom_segment(aes(yend=esi_field, x=0, xend=percent), data =. %>% filter(variable == 'guidance')) +
+  geom_point(size = 5, shape = 21, colour = 'black', data = . %>% filter(variable == 'guidance')) + 
+  geom_point(size = 2, shape = 23, colour = 'black', data = . %>% filter(variable == 'internal_guidance_section')) + 
+  geom_col(data = . %>% filter(variable == 'guidance', grepl('ALL', esi_field)), alpha = .2) + 
+  xlab('journals (%)') +
+  scale_fill_domains() +
+  theme_apa() +
+  theme(
+    panel.grid.major.x = element_line(),
+    axis.title.y = element_blank(),
+    legend.position = 'none',
+    #strip.text = element_text(hjust = 0),
+    strip.text.x = element_blank())
 
@@
 Guidance was less common for checking model assumptions (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "model_assume") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "model_assume") %>% pull(percent) %>% round(0)`%, Supplementary Figure C1, Panel A), multiple comparisons (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "multi_compare") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "multi_compare") %>% pull(percent) %>% round(0)`%, Supplementary Figure C1, Panel B), statistical significance (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "significance") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "significance") %>% pull(percent) %>% round(0)`%, Supplementary Figure C1, Panel C), one-tailed hypothesis tests (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "one_sided") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "one_sided") %>% pull(percent) %>% round(0)`%, Supplementary Figure C1, Panel D), Bayesian statistics (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "bayes") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "bayes") %>% pull(percent) %>% round(0)`%, Supplementary Figure C1, Panel E), null hypotheses (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "null_hypo") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "null_hypo") %>% pull(percent) %>% round(0)`%, Supplementary Figure C1, Panel F), subgroup analyses (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "subgroup") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "subgroup") %>% pull(percent) %>% round(0)`%, Supplementary Figure C2, Panel A), secondary outcomes (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "secondary") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "secondary") %>% pull(percent) %>% round(0)`%, Supplementary Figure C2, Panel B), missing data (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "missing") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "missing") %>% pull(percent) %>% round(0)`%, Supplementary Figure C2, Panel C), sensitivity analyses (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "sensitivity") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "sensitivity") %>% pull(percent) %>% round(0)`%, Supplementary Figure C2, Panel D), non-parametric tests (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "non_param") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "non_param") %>% pull(percent) %>% round(0)`%, Supplementary Figure C3, Panel A), baseline covariates (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "baseline_covar") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "baseline_covar") %>% pull(percent) %>% round(0)`%, Supplementary Figure C3, Panel B), outliers (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "outliers") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "outliers") %>% pull(percent) %>% round(0)`%, Supplementary Figure C3, Panel C), and categorisation of continuous outcomes (n = `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "cat_continuous") %>% pull(n)`, `r sum_tab_topics %>% filter(esi_field == "OVERALL", topic == "cat_continuous") %>% pull(percent) %>% round(0)`%, Supplementary Figure C3, Panel D).
-Guidance on the twenty topics tended to be more common amongst journals in biomedical and health-related domains, especially Clinical Medicine, compared to other domains.
+Guidance on the twenty topics tended to be more common amongst journals in Health and Life Sciences, especially Clinical Medicine, compared to other domains.
 Examples of verbatim guidance are provided in Supplementary Table D1.
@@
   mutate(topic_display = factor(topic_display, levels = topic_display, ordered = T)) %>%
-  ggplot(aes(x=topic_display, y = percent)) +
-    ylim(0,30) +
-    geom_segment(aes(xend=topic_display, y=0, yend=percent)) +
+  ggplot(aes(x=topic_display, y = n)) +
+    ylim(0,100) +
+    geom_segment(aes(xend=topic_display, y=0, yend=n)) +
     geom_point(size = 5, shape = 21, fill = 'black', colour = 'black') + 
-    ggtitle('Specific statistical guidance') +
-    ylab('journals (%)') +
+    ggtitle('B. Specific statistical guidance') +
+    ylab('journals (n)') +
     theme_apa() +
@@
 
-```{r generaltopicsfig, fig.path='figs/', dev=c('png', 'pdf'), fig.width = 8, fig.height = 6.5, fig.cap = "Percentage of journals offering some statistical guidance (Panel A circles), having a dedicated statistical guidance section (Panel A diamonds), or offering guidance on each of twenty prespecified statistical topics (Panel B).
+```{r generaltopicsfig, fig.path='figs/', dev=c('png', 'pdf'), fig.width = 8, fig.height = 9, fig.cap = "Percentage of journals offering some statistical guidance (Panel A circles), having a dedicated statistical guidance section (Panel A diamonds), or offering guidance on each of twenty prespecified statistical topics (Panel B).
 Black circles represent data for all journals overall (N = 330 journals).
@@
 Discipline abbreviations: Agricultural Sciences (AGRI), Biology & Biochemistry (BIO), Chemistry (CHEM), Clinical Medicine (MED), Computer Science (COMSCI), Economics & Business (ECON), Engineering (ENGIN), Environment & Ecology (ECO), Geosciences (GEO), Immunology (IMMUN), Materials Science (MATSCI), Mathematics (MATH), Microbiology (MICBIO), Molecular Biology & Genetics (MOLBIO), Multidisciplinary (MULTI), Neuroscience & Behavior (NEURO), Pharmacology & Toxicology (PHARM), Physics (PHYS), Plant & Animal Science (PLANT), Psychiatry & Psychology (PSY), Social Sciences (SOCSCI), Space Science (SPACE)."}
-plot_any_dedicated_joint + plot_all_topics + plot_annotation(tag_levels = 'A') &
-  theme(plot.tag.position = c(0, 1),
+
+# plot_any_dedicated_joint + plot_all_topics + plot_annotation(tag_levels = 'A') & theme(plot.tag.position = c(0, 1),
+#        plot.tag = element_text(size = 14, hjust = 0, vjust = 0))
+
+plot_any_dedicated_joint + plot_all_topics & theme(plot.tag.position = c(0, 1),
         plot.tag = element_text(size = 14, hjust = 0, vjust = 0))
@@
 # ggsave(plot = plot_all_topics, file = here('manuscript','figs','twitter','fig1B.png'), width = 6, height = 6.5, units = "in", dpi = 300)
+```
+```{r generaltopicsfignew, fig.path='figs/', dev=c('png', 'pdf'), fig.width = 8, fig.height = 9, fig.cap = "Percentage of journals offering some statistical guidance (Panel A circles), having a dedicated statistical guidance section (Panel A diamonds), or offering guidance on each of twenty prespecified statistical topics (Panel B).
+Black circles represent data for all journals overall (N = 330 journals).
+Coloured circles and diamonds in Panel A represent individual scientific disciplines (n = 15 journals).
+Data is in descending order by percentage (after overall data in Panel A).
+CCV in Panel B refers to categorizing continuous outcomes.
+Discipline abbreviations: Agricultural Sciences (AGRI), Biology & Biochemistry (BIO), Chemistry (CHEM), Clinical Medicine (MED), Computer Science (COMSCI), Economics & Business (ECON), Engineering (ENGIN), Environment & Ecology (ECO), Geosciences (GEO), Immunology (IMMUN), Materials Science (MATSCI), Mathematics (MATH), Microbiology (MICBIO), Molecular Biology & Genetics (MOLBIO), Multidisciplinary (MULTI), Neuroscience & Behavior (NEURO), Pharmacology & Toxicology (PHARM), Physics (PHYS), Plant & Animal Science (PLANT), Psychiatry & Psychology (PSY), Social Sciences (SOCSCI), Space Science (SPACE)."}
+plot_any_dedicated_joint & theme(plot.tag.position = c(0, 1),
+        plot.tag = element_text(size = 14, hjust = 0, vjust = 0))
 ```
@@
 
-  ggplot(data = thisData, aes(x=esi_abbr, y = percent, fill = esi_abbr)) +
+  ggplot(data = thisData, aes(x=esi_abbr, y = percent, fill = domain)) +
     ylim(0,100) +
@@
 
+```{r}
+# REVISED PLOTS
+
+sum_tab_topics_domains <- sum_tab_topics %>% 
+  group_by(domain, topic_display) %>% 
+  summarise(n = sum(n), N = sum(N)) %>% 
+  mutate(prop = n/N, percent = n/N*100)
+
+# get order of topics in terms of most frequently mentioned
+topics_ordered <- sum_tab_topics_domains %>%
+  filter(domain == 'OVERALL') %>%
+  arrange(percent) %>%
+  pull(topic_display)
+
+sum_tab_topics_domains <- sum_tab_topics_domains %>%
+  mutate(topic_display = factor(topic_display, levels = topics_ordered, ordered = T))
+
+customDomains <- c("HEALTH & LIFE SCIENCES"="HEALTH & LIFE SCIENCES", 
+                      "MULTIDISCIPLINARY"="MULTI- DISCIPLINARY", 
+                      "PHYSICAL SCIENCES"="PHYSICAL SCIENCES", 
+                      "SOCIAL SCIENCES"="SOCIAL SCIENCES")
+
+specific_fig_new <- sum_tab_topics_domains %>%
+  filter(domain %notin% c('FORMAL SCIENCES','OVERALL')) %>% 
+  ggplot(aes(y = topic_display, x = percent, fill = domain)) +
+  facet_wrap(.~domain, nrow = 1, labeller = as_labeller(customDomains, multi_line = T, default=label_wrap_gen(10))) +
+  #facet_wrap(.~domain, nrow = 1, labeller = label_wrap_gen(width = 2, multi_line = TRUE)) +
+  xlim(0,60) +
+  xlab('journals (%)') +
+  geom_segment(aes(yend=topic_display, x=0, xend=percent)) +
+  geom_point(size = 5, shape = 21, colour = 'black') +
+  scale_fill_domains() +
+  theme_apa() +
+  theme(
+    panel.grid.major.x = element_line(),
+    axis.title.y = element_blank(),
+    legend.position = 'none',
+    strip.background = 
+  )
+
+```
+
+```{r specifictopicsfignew, fig.path='figs/', dev=c('png', 'pdf'), fig.width = 8, fig.height = 7, fig.cap = "Caption here"}
+specific_fig_new
+```
+
 ## External statistical guidance
@@
          NLM = "NLM RESEARCH REPORTING GUIDELINES AND INITIATIVES (HTTPS://WWW.NLM.NIH.GOV/SERVICES/RESEARCH_REPORT_GUIDE.HTML)"))
+```
+
+```{r}
+sum_tab_domain %>% filter(variable == 'external_guidance') %>% arrange(desc(percent))
 ```
@@
 Our investigation found that 48% of 330 top-ranked journals across 22 scientific disciplines provided some statistical guidance to authors, with 28% of journals having a dedicated statistical guidance section in their instructions to authors.
-It was far more common for journals to provide statistical guidance in biomedical and health-related disciplines relative to non-health related disciplines like the social sciences and physical sciences.
+It was far more common for journals to provide statistical guidance in Health and Life Sciences relative to non-health related disciplines like the social sciences and physical sciences.
 Statistical guidance addressed a range of topics, and was most common for hotly debated topics: p-values, confidence intervals, sample size justification, and effect sizes.

### 2022-05-15 manuscript/supplementary_information (conflicted).Rmd
AmStat round 2 revisions

--- 2022-04-28 manuscript/supplementary_information.Rmd
@@
 ```{r}
+# ORIGINAL GRID PLOTS 
 my_palette <- c("#95D840FF", "#404788FF")
@@
 
-```{r gridOne, fig.path='figs/', fig.cap = "Whether each journal provided guidance on each of twenty statistical topics.
-Journals that shared publisher guidance are represented by one row (per publisher).
-Journals that did not offer any statistical guidance at all, or journals that referred only to guidance in external sources, are not shown.
-CCV = categorizing continuous variables.", fig.height = 11, fig.width = 8}
-plt_one
+```{r}
+# REVISED GRID PLOTS 
+my_palette <- c("#95D840FF", "#404788FF")
+
+disciplineOrder <- sum_tab_general %>%
+  filter(variable == 'guidance') %>%
+  arrange(desc(percent)) %>%
+  pull(esi_field)
+
+plt_data <- d_all %>%
+  #filter(has_internal_guidance == T) %>% # select journals with internal guidance
+  # Recode NAs to FALSE
+  mutate(across(
+    starts_with('has_'), ~replace_na(., FALSE) )) %>% 
+  mutate(journal = toupper(journal),
+         journal = ifelse(has_publisher_guidance == T, paste0(journal,'*'), journal)) %>% # add asterisk to journals that share publisher guidance
+  select(journal,esi_field,domain, starts_with("has_"), -c(has_guidance, has_internal_guidance, has_internal_guidance_section, has_external_guidance, has_publisher_guidance)) %>%
+  #filter(if_any(starts_with("has"), ~ .x == TRUE)) # remove the nine journals that had no guidance on any of the 20 topics
+  mutate(journal = fct_recode(journal, # shorten particularly long journal names well know by abbreviation
+                              "PNAS" = "PROCEEDINGS OF THE NATIONAL ACADEMY OF SCIENCES OF THE UNITED STATES OF AMERICA",
+                              "JAMA" = "JAMA-JOURNAL OF THE AMERICAN MEDICAL ASSOCIATION",
+                              "THE BMJ" = "BMJ-BRITISH MEDICAL JOURNAL",
+                              "JOURNAL OF ALLERGY & CLINICAL IMMUNOLOGY" = "JOURNAL OF ALLERGY AND CLINICAL IMMUNOLOGY",
+                              "JOURNAL OF ALLERGY & CLINICAL IMMUNOLOGY-IN PRACTICE" = "JOURNAL OF ALLERGY AND CLINICAL IMMUNOLOGY-IN PRACTICE",
+                              "INTERNATIONAL JOURNAL OF BEHAVIORAL NUTRITION & PHYSICAL ACTIVITY" = "INTERNATIONAL JOURNAL OF BEHAVIORAL NUTRITION AND PHYSICAL ACTIVITY")) %>% 
+  pivot_longer(
+    -c(journal,esi_field,domain),
+    names_to = "topic",
+    names_prefix = 'has_',
+    values_to = "outcome",
+    values_drop_na = FALSE
+  ) %>%
+  mutate(outcome = case_when(
+    outcome ~ "Mentioned",
+    !outcome ~ "Not mentioned",
+    TRUE ~ "Not applicable (No internal guidance)"
+  )) %>%
+  mutate(topic_display = case_when( # create display versions of topic names
+    topic == "guidance" ~ "Provides any statistical guidance",
+    topic == "internal_guidance" ~ "Provides journal-specific statistical guidance",
+    topic == "internal_guidance_section" ~ "Dedicated statistical guidance section",
+    topic == "external_guidance" ~ "Refers to external statistical guidance",
+    topic == "p_value" ~ "p values",
+    topic == "significance" ~ "statistical significance",
+    topic == "null_hypo" ~ "null hypotheses",
+    topic == "sample_size" ~ "sample size",
+    topic == "conf_int" ~ "confidence intervals",
+    topic == "effect_size" ~ "effect sizes",
+    topic == "multi_compare" ~ "multiple comparisons",
+    topic == "subgroup" ~ "subgroup analyses",
+    topic == "baseline_covar" ~ "baseline covariates",
+    topic == "non_param" ~ "non-parametric tests",
+    topic == "sensitivity" ~ "sensitivity analyses",
+    topic == "model_assume" ~ "model assumptions",
+    topic == "exclusion" ~ "data exclusions",
+    topic == "outliers" ~ "outliers",
+    topic == "missing" ~ "missing data",
+    topic == "one_sided" ~ "one-tailed tests",
+    topic == "bayes" ~ "Bayesian statistics",
+    topic == "secondary" ~ "secondary outcomes",
+    topic == "prespecify" ~ "prespecify analyses",
+    topic == "cat_continuous" ~ "CCV",
+    topic == "publisher_guidance" ~ "Shares publisher guidance"
+  )) %>%
+  mutate(topic_display = factor(topic_display, levels = ordered_topics), # order by n topics mentioned
+         esi_field = factor(esi_field, levels = disciplineOrder), # order by discipline
+         domain = factor(domain, levels = domainOrder[1:5])) # order domain
+
+gridPlotter <- function(thisDiscipline){
+  thisData <- plt_data %>% 
+    filter(esi_field == thisDiscipline)
+
+  journalOrder <- thisData %>% 
+    mutate(outcome = factor(outcome),
+           journal = as.character(journal),
+           journal = factor(journal)) %>%
+    count(journal,outcome, .drop = F) %>%
+    filter(outcome == 'Mentioned') %>%
+    arrange(n) %>%
+    pull(journal)
+
+  thisData %>%
+    mutate(journal = factor(journal, levels = journalOrder)) %>%
+    ggplot(aes(x = topic_display, y = journal)) +
+      ggforce::facet_col(facets = vars(esi_field), scales = 'free_y', space = 'free') +
+      geom_tile(aes(fill = outcome), colour = 'black', size = .25, alpha = .6) +
+      scale_fill_manual(
+        values = my_palette,
+        # labels = c('Not mentioned', 'Unclear', 'Mentioned')
+      ) +
+      scale_x_discrete(position = 'top') +
+      scale_y_discrete(labels = wrap_format(40)) + # wrap text on y axis
+      theme(plot.margin = margin(0, 50, 0, 0),
+            axis.ticks = element_blank(),
+            axis.title.y = element_blank(),
+            axis.title.x = element_blank(),
+            axis.line = element_blank(),
+            panel.grid.major=element_line(colour=NA),
+            panel.grid.minor=element_line(colour=NA),
+            legend.position = 'top',
+            axis.text.x  = element_text(angle = 45, hjust = 0),
+            axis.text.y  = element_text(size = 8),
+            panel.background = element_blank())
+}
 ```
 
-```{r gridTwo, fig.path='figs/', fig.cap = "Continuation of Supplementary Figure D2.", fig.height = 11, fig.width = 8}
-plt_two
+```{r gridMed, fig.path='figs/', fig.cap = "Whether each Clinical Medicine journals provided guidance on each of twenty statistical topics.
+Topics are ordered left to right by frequency of mentions across disciplines.
+Disciplines, followed by journals, are ordered top to bottom by frequency of statistical guidance.
+CCV = categorizing continuous variables.", fig.height = 5, fig.width = 8}
+gridPlotter('CLINICAL MEDICINE')
+```
+
+```{r gridMed, fig.path='figs/', fig.cap = "Continuation of Supplementary Figure D2 for the discipline of Immunology", fig.height = 5, fig.width = 8}
+gridPlotter('IMMUNOLOGY')
+```
+
+```{r gridMed, fig.path='figs/', fig.cap = "Continuation of Supplementary Figure D2 for the discipline of Immunology", fig.height = 5, fig.width = 8}
+gridPlotter('PSYCHIATRY & PSYCHOLOGY')
 ```

### 2022-05-15 analysis/preprocessing.R

@@ -431,21 +431,29 @@ d_coding <- d_coding %>% select(-external_guidance_clean)
 # bind d_coding and d_journals
 d_all <- inner_join(d_coding,d_journals, by = 'journal')
 
+# rename some disciplines for display purposes
+d_all <- d_all %>%
+  mutate(esi_field = case_when(
+    esi_field == "ENVIRONMENT_ECOLOGY" ~ "ENVIRONMENT & ECOLOGY",        
+    esi_field == "Multidisciplinary" ~ "MULTIDISCIPLINARY",          
+    esi_field == "PSYCHIATRY_PSYCHOLOGY" ~ "PSYCHIATRY & PSYCHOLOGY",      
+    esi_field == "SOCIAL SCIENCES, GENERAL" ~ "SOCIAL SCIENCES (GENERAL)",
+    TRUE ~ esi_field
+  ))
+  
 # add higher-level domain classifications
 d_all <- d_all %>%
   mutate(domain = case_when(
-    esi_field %in% c('AGRICULTURAL SCIENCES','BIOLOGY & BIOCHEMISTRY','ENVIRONMENT_ECOLOGY','MICROBIOLOGY','MOLECULAR BIOLOGY & GENETICS','NEUROSCIENCE & BEHAVIOR','PLANT & ANIMAL SCIENCE') 
-    ~'Life sciences',
-    esi_field %in% c('CLINICAL MEDICINE','IMMUNOLOGY','PHARMACOLOGY & TOXICOLOGY') 
-    ~'Health sciences',
+    esi_field %in% c('AGRICULTURAL SCIENCES','PSYCHIATRY & PSYCHOLOGY','BIOLOGY & BIOCHEMISTRY','ENVIRONMENT & ECOLOGY','MICROBIOLOGY','MOLECULAR BIOLOGY & GENETICS','NEUROSCIENCE & BEHAVIOR','PLANT & ANIMAL SCIENCE', 'CLINICAL MEDICINE','IMMUNOLOGY','PHARMACOLOGY & TOXICOLOGY') 
+    ~'HEALTH & LIFE SCIENCES',
     esi_field %in% c('CHEMISTRY','ENGINEERING','GEOSCIENCES','MATERIALS SCIENCE','SPACE SCIENCE','PHYSICS') 
-    ~'Physical sciences',
[…]

### 2022-05-24 manuscript/manuscript.Rmd
edits to PRC figure

@@
 
-```{r generaltopicsfig, fig.path='figs/', dev=c('png', 'pdf'), fig.width = 8, fig.height = 9, fig.cap = "Percentage of journals offering some statistical guidance (Panel A circles), having a dedicated statistical guidance section (Panel A diamonds), or offering guidance on each of twenty prespecified statistical topics (Panel B).
+```{r prcfig, fig.path='figs/', dev=c('png', 'pdf'), fig.width = 7, fig.height = 8, fig.cap = "Percentage of journals offering some statistical guidance (Panel A circles), having a dedicated statistical guidance section (Panel A diamonds), or offering guidance on each of twenty prespecified statistical topics (Panel B).
 Black circles represent data for all journals overall (N = 330 journals).
@@
 
-plot_any_dedicated_joint + plot_all_topics & theme(plot.tag.position = c(0, 1),
-        plot.tag = element_text(size = 14, hjust = 0, vjust = 0))
+# PEER REVIEW CONGRESS PLOT
+plot_any_dedicated_joint + xlim(0,105) +ggtitle('A.
+General statistical guidance') + theme(axis.text.y = element_text(size = 10), plot.title.position = "plot", plot.title = element_text(hjust = 0)) + plot_all_topics + ggtitle('B.
+Specific statistical guidance') + theme(axis.text.y = element_text(size = 12), plot.title.position = "plot", plot.title = element_text(hjust = 0))
 
@@
 ```
+
 ```{r generaltopicsfignew, fig.path='figs/', dev=c('png', 'pdf'), fig.width = 8, fig.height = 9, fig.cap = "Percentage of journals offering some statistical guidance (Panel A circles), having a dedicated statistical guidance section (Panel A diamonds), or offering guidance on each of twenty prespecified statistical topics (Panel B).
